### 2017-12-20 ARTICLE.tex
First commit
First commit

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}

%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}

%%%% To use the SIGCHI extended abstract template, please visit
% https://www.overleaf.com/read/zzzfqvkmrfzn

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{booktabs} % For formal tables
%% perso packages:
\usepackage{epstopdf}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{multirow}
%\usepackage{arydshln}
\usepackage{subfig}
\usepackage{balance}
\usepackage{caption}
\usepackage{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT's Identification
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CONFERENCE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Conference
\acmConference[ICMR 2018]{ACM Yokohama conference}{June 2018}{Yokohama, Japan} 
\acmYear{2018}
\copyrightyear{2018}

\acmArticle{4}%??
\acmPrice{15.00}%??

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Annotating, understanding, and predicting long-term video memorability}
\titlenote{Produces the permission block, and copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
\texttt{acmart.pdf} document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Romain Cohendet}
%\authornote{}
%\orcid{1234-5678-9012}
\affiliation{%
	\institution{Technicolor}
	\streetaddress{}
	\city{Rennes} 
	\state{France} 
	\postcode{}
}
\email{romain.cohendet@technicolor.com}

\author{Karthik Yadati}
\affiliation{%
	\institution{University of Delft}
	\streetaddress{}
	\city{Delft} 
	\state{} 
	\postcode{}
}
\email{N.K.Yadati@tudelft.nl}

\author{Ngoc Khanh Duong}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes} 
	\country{France}}
\email{Quang-Khanh-Ngoc.Duong@technicolor.com}

\author{Claire-Helene Demarty}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes}
	\country{France}
}
\email{Claire-Helene.Demarty@technicolor.com}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{R. Cohendet et al.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Following the interest of these last years for image memorability prediction, video memorability has recently attracted attention of researchers.
The growing number of video shared everyday forces us to find new way to deal with them to make the most relevant their occurrence in our everyday lives.

Recently, two studies have tried to predict video memorability ; however, the immaturity of this work is obvious.
In particular, none of these studies on images nor videos tried to measure 'very' long-term memory: they just measured the memory of items some minutes after the memory encoding.
In this paper, we propose a dataset of 700 videos annotated with memorability scores corresponding to real long-term memory.
We discuss in details the method and the proponents.
We then propose a model to predict memorability of videos and test it on the available dataset of images and videos.
% Our approach improves over current memorability methods in this task.
The material -- videos plus theirs memorability scores -- is made accessible for researchers.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% NEW ACM TAXONOMY to categorise ACM article
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010520.10010553.10010562</concept_id>
	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010520.10010575.10010755</concept_id>
	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10010520.10010553.10010554</concept_id>
	<concept_desc>Computer systems organization~Robotics</concept_desc>
	<concept_significance>100</concept_significance>
	</concept>
	<concept>
	<concept_id>10003033.10003083.10003095</concept_id>
	<concept_desc>Networks~Network reliability</concept_desc>
	<concept_significance>100</concept_significance>
	</concept>
	</ccs2012>  
\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% KEYWORDS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\keywords{Video memorability, Long-term memory, Measurement protocol, Memorability scores, Deep learning, Multimedia information retrieval}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ABOUT ICMR'2018 article...
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
%if we want the text be called from there
 %\input{samplebody-conf}

%% About the ICMR article on Media retrieval with video memorability
	% focus on "Multimedia content-based retrieval"
	% Length: from 6 to 8 pages + 1 additional page for references (no longer distinction between long and short papers)
	% Submission deadline: 2018/01/20
	% Contributions addressing the challenges of large-scale search and user behavior analysis are especially welcome

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}%nb of sections %2 PAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%citer Qomex
%citer Deep learning emotional images bias
\paragraph{Opening} % oder Ok
To deal with the constant increase in shared videos, we must imagine new ways to organize -- in particular, to retrieve -- videos with the aim of improving the relevance of their occurrences in our every day lives.
Contrary to other metrics used to quantify image of video importance, such as aesthetics or interestingness, whose corresponding ground truths corresponds to subjective judgments, memorability can be objectively measured by a memory test.
As such, memorability can be regarded as a particularly relevant element to help us picking a video among several ones.

\paragraph{The field of video memorability is very immature}
%from images to video memorability
The image memorability prediction challenge has attracted increasing attention from the seminal work of Isola \textit{et al.} released in 2001 \cite{isola_2011_makes}.
Recently, models achieved very good results with the introduction of deep learning to address the challenge \cite{khosla_2015_understanding,baveye_2016_deep}% plus papier Hammad
As a result of this success, the search field has been extended to videos.
Video memorability prediction is however a very new field, and there are only two studies, to our knowledge, that address this issue \cite{han_2015_learning,shekhar_2017_show}.

At least two important problems could explained this scarcity of studies.
Firstly, protocols used in the existing studies are not smart enough, but are easily critisizeable.
Secondly, authors didn't make their data accessible for community, so no database exist.

\subsection{Problem 1: The problem of the weakness of the studies}
%%% Critiques of previous studies %%%

\paragraph{Critics of Han et al.}
The previous attempt at creating memorability video dataset first approach to model video memorability had been propose by \cite{han_2015_learning}. 
This first attempt is a "strange" adaptation of \cite{isola_2011_makes} protocols.

% A memory game of 12 hours
- Firstly, the authors call their task a "memory game" as Isola \textit{et al.}. If it were justified for Isola et al., 2011, la tâche dure pour 1 unique participant environ 12 heures. The authors do not provide this time, but it cans be deduced from the by the following calculation:

Thenks to this maneer, participants obtained a dataset of videos with only 20 participants (we don\'t know if they are student or payed).

répatit sur 10 jours diffférens (5 pour les phases dapprentissage + 5 pour les phase de mesure de la mémoire).

uen autre aberration e

%critique of Shekhar et al.
In \cite{shekhar_2017_show}, 
- des questions pas forcément aussi faciles/diffilec entre les vidéos, qui peuvent expliquer leur

Enfin, dans les deux études précédentes, la congruence inter-humaine, minimum nécessaire pour s'assurer de la cohérence des données et avoir une base de comparaison pour les modèles, n'a pas été fournie.

%Lasting long-term memory for IMAGES - critic of Isola et al.
on comprend les auteur d'avoir agit ainsi pour consituter une base de grand ampleur (Khosla et al.)
It is long-term memory a proprement parlé because.
but memories evolove in l-g mem
=> Ebbinghaus curve
So memorabilities of souvenirs can change after a certain delays
%as proposed in [ma thèse] 
%Memorability continue to evolve in long-term memory --> Ebbinghaus -- or more recent similar -- curve
%Memorability image/videos in computer vision never interested in REAL long-term memory
%We want to correct this problem by proposing a new way to measure memorability.
% saut de la tâche d'apprentissage (remplacée par le questionnaire)
% To sum up, we want to obtain the "long-lasting memorability" of the videos

\paragraph{Problem 2: No existing database}
%there is no database
=> Dataset non disponible pour les deux études.

The fact that there exist no dataset to modeling video memorability is the most serious obstacle which prevent video memorability prediction to take off.
This should be the very first goal of pioneers of this new research field: to provide data to community.
For images, the pioneers created and made downloadable such a database \cite{isola_2011_makes}, which enable to made image memorability research flourishing.
The second released database, far larger, launched new possibility for deep learning and enable to obtained the aforementioned really good results \cite{khosla_2015_understanding}.
Nous pouvons parier qu'il en sera de même pour le cham de la mémorabilité des vidéos.

%it's not easy to construct the first dataset
The establishment of a dataset, in particular of the first dataset of this kind, may not be self-evident.

Elle nécessite de définir les points cardinaux qui vont guider – et contraindre – les chercheurs qui utiliseront ces premières données disponibles.

Pour preuve, l'influence considérable qu'Isola et al. (puis Khosla et al., de la même équipe, avec leur base de données beaucoup plus conséquente rendue disponible en 2015) ont eu sur les études réalisées dans leur sillage : jusqu'à aujourd'hui, la quasi-totalité des études ayant porté sur la mémorabilité des images ont adopté – avec les bases de données de ces auteurs – leurs définitions de la mémorabilité.

Or, la manière dont les auteurs ont mesuré la mémorabilité des images, qui n'est pratiquement jamais rediscutée, n'est qu'une des manières de procéder.

Aussi, si les modèles de prédiction de la mémorabilité des images ont obtenu d'excellents résultats (très proches de ceux qu'on obtiendrait en inférant des résultats d'un groupe d'observateur les résultats d'un autre groupe d'observateurs), qu'est-ce que ces modèles prédisent si bien, en fait ?

	%Difficulty: define video unity; complexity of the videos compare to images
	Videos are more complex than images. Contrary to images, videos do not constitute clearly defined units, but have supplementary dimensions -- sound and movement -- that makes difficult the definition of what a video is.
	%from ahn et al. Videos contain a synthesis of visual, aural, textual, and 	motion information and generally convey relatively richer 	media content than images.

%Conclusion paragraph: we want to contribute to the protocol creation by proposing one method
Aussi cette étude veut participer à réfléchir aux protocoles qui pourraient être employée pour contituer à l'avenir à un protocole de plus grande envergure.

\paragraph{Goals of our study}
%link the two reasons precebndtky evoque
These two reasons are linked. They give us our priority: build a dataset and define a protocol to do that.
This dataset -- and a fortiori the protocol -- should avoid the drawbacks of previous work in video and image memorability.
%in image memorability, difficultys come from que'elle ont justement utilisés ces bases de données.

%2 resaons => our goal of our study => correct these drawbacks
=> To propose a new protocol to collect data/for annotation
=> To propose a dataset
=> The dataset will have "lasting" long-term memorability annotations
We want our work would benefit to video but also to image memorability.

%Ainsi, puisque les limitations d'un protocole destiné à collecter des données tendent à se répercuter dans les travaux qui utilisent celles-ci, un protocole destiné à constituer la base nécessaire pour répondre au défi de la prédiction de la mémorabilité des vidéos devra répondre à de strictes exigences de qualité. La qualité des données obtenues, primordiale, nécessite une connaissance profonde des phénomènes que ces données sont censées refléter. La mémoire étant un phénomène étudié (depuis plus d'un siècle) par la psychologie, et les modèles computationnels permettant de l'inférer des caractéristiques des vidéos appartenant au domaine de l'informatique, l'approche transdisciplinaire semble la plus adaptée dans ce cas.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Secondary goals}
%%% OTHER INTERSTING RESULTS
%will depend on our results
%\paragraph{Exploiting response time for memorability scores}
%D'autre part, nous voulons également estimer si le temps de réponse à un effet sur la mémorabilité, c'est-à-dire qu'il pourrait être pris en compte pour corriger les scores calculer. Note hypothèse est que, plus les participants sont sûrs d'eux, plus ils vont tendre à répondre rapidement. Att!! peut pas être tester directement : We hypothesized that the most memorable the sequence, the faster the participant will answer. Parce que les video les mieux représentés dans l'esprit sont les plus accessibles et que les participants ne veulent pas louper la préponse quand ils sont sûrs.
% Citer papiers - s'il en existe -- où les auteurs exploitent le temps de réponse pour corriger les scores de mémorabilité;

%\paragraph{Personalization of the data}
%Taking into account human factors
	%see results Logistic regression vs. SVM

il semble des dernière études qu'il serait intéressant de prendre en temps le dans le calcul de la mémorabilité [LaMem + papier sur la VM où ils prennent en compte le temps dans leurs scores]

%If results Repetition vs. memorability
The semantization that can affect long-term memorability of videos

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Construction of the dataset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Opening -- Ok for now
In this section, we describe the new protocol we propose to measure video memorability in order to collect the long-lasting memorability of the videos.
The protocol is also lighter than ordinarily since there is no need of learning step to encode the material before the recognition task, as we measure the participants' memory established prior the experiment.

The memory task involve three materials: videos sequences, questionnaire and participants.

%Insist on Contribution about the protocol (novel: how to measure long-term memory)

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memory task}
To measure the "real" long-term memory, we designed the recognition task presented in the figure ...%ref.
%figure Recognition task

% Targets and Fillers
The 10-sec video sequences are displayed one after another, separated by an inter-stimuli interval (ISI).

%define target and filler
Les vidéos cibles sont mêlées aux vidéos de remplissage, et le participant doit dire, pour chaque vidéo qui lui est présentée, s’il l’a déjà vue ou non.
To answer, he must press the spacebar during the displaying of the video sequences (than answer when they arrived during the 1-s inter-stimuli interval).
movies they have seen (the targets) and never seen (the fillers).
Even if the participant answers, the video continue to run until its end (to avoid the temptation of ansewering to faster finish the experiment).

"Respond ONLY if you remember the sequence AND NOT THE movie from which it comes".%reprendre phrase que j'ai écrite

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The video sequences}
%% movies
A list of 100 movies (available with the downloadable data) was first established, for which we mixed their popularity and their genres.
The movies quality was HD 720-1080, except for ??? dvdrip movies.
orginal-english version (without subtitles) (occidental movies, except for Slumdog millionaire) %to check

%%video sequences
To constitute our dataset, we manually selected 7 video sequences of 10 seconds from each movies (for a total of 700 videos).
The sequences had to meet several criteria to be selected.
First, not to be part of the official trailers, to be sure that, if a participant report not having seen the movie, he won't have seen the sequence presented for the recognition in a trailer.
Second, we wanted our sequences to constitute "logical" units, that is to say to "protect the sens" of the videos. Indeed, semantics is linked to memorability of images \cite{isola_2014_makes}, from which we can imagine that it is also the case for videos. But, contrary to images that are clearly cut units, sequences memorability depends on the cut. To maintain a relative semantic consistency intra-sequence, the cutters were provided with the instruction to avoid to cut in the middle of a sentence, of to agglomerate very different plans together. The idea was to increase the coherence of the sequences to diminish the difficulty of he computational models to learn without impairing ther capacity to generalize.

%%Generic vs. typical
During the manual selection of the sequence, we chose 127 sequences (among the 700) we called "generic" by contrast to "typical" ones. Generic sequences are parts of the movie which do not contain any elements that would enable someone to easily guess that the sequences belong to the particular movie. The list of undesirable includes but is not limited to: scenes with recognizable famous actors (e.g. in Kill Bill, avoid the scenes with Uma Thurman, Lucy Liu, David Carradine…); scenes famous for their music, gesture... (e.g. in Kill Bill, avoid the scenes with music that enable to easily guess that the sequence comes from Kill Bill); scenes with typical elements (e.g. a spaceship or an alien from which you can easily guess that you are watching a Star Wars movie). The cutters were provided with the instructions to ask themselves the following question when choosing a generic sequence in a movie: “Could I easily guess this sequence belongs to the movie I extracted it from?” -- if the answer was yes, so don’t select this sequence.
In some movies (e.g. Kill Bill 1, Star wars, 2001 A Space Odyssey, etc.), just a few 10-sec sequences of no sequence at all were found to meet these criteria (e.g. The lord of the ring); for these movies, we don't have generic sequences.
By contrast, non-generic sequences were namely two 10-sec parts of the movie that you think are very representative of the movie (you can choose any part of the movie without restriction).
These two types of sequences should be considered in link with the instruction of the task written above (even if you know that a sequence comes from a movie, just answer if you recognize THE sequence): it is just a supplementary control (qui jour un rôle symétrique au taux de FA pour la prise de risque).

%%Indoor/Outdoor
In Isola et al., 2014, the authors showed that memorability directly correlated with color. To explain thi result, they advanced the hypothesis that the difference in memorability between indoor and outdoor scenes was due to the fact that the later tend to be less memorable that the older, and that the color tend to linked with a differences in colors, could explain this result. To validate this hypothesis, we manually annotated our 700 sequences as consisting of an indoor or outdoor scene.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Questionnaire}
The targets and fillers that constitute the items of the recognition task were different for each participant.
Indeed, the video sequences viewed by a participant were selected based on response to a questionnaire presented in the figure ???
Questionnaire with the 100 movies.

%figure Questionnaire

Le nom du film, suivi de l’affiche, du nom du réalisateur et des acteurs principaux – qui permettent de s’assurer que le film que le participant a en tête est bien celui qui lui est présenté – est présenté au participant.

%Other questions associated with the questionnaire
For each movie, the participant had first to answer the following question: "Do you remember watching this movie?" -- with the fact that was madde clear that the film must been seen EN ENTIER.

If he answered "yes"
Please let us know how confident you are of having seen this movie (I am not/slightly/50\%/considerably/100\% confident)
Please let us know when did you last see this movie (<1 month; <1 year; <5 years; < 10 years; > 10 years)
Please let us know how many times did you watch this movie (once; 2-4 times; 5-9 times; 10-19 times; >20 times)

If he answered "no"
Please let us know how confident you are of NOT having seen this movie (I am not/slightly/50\%/considerably/100\% confident)

The goal of the confidence question was to be sure that we select video sequences from movie , to increase our certitude that target/filler were really targets/fillers.

The other two questions were used for the analysis of the results to assess factors a priori susceptible to influence video memorability. The question about the number of times the movie had been seen for us to evaluate the effect of repetition on memorization, and the question about the last time tha movie had been seen to evaluate the effect of time passage on long-term memorability, that continue to be affected in longterm memory %citer un papier de moi

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Participants, facilities and apparatus}
%participants
??? (105) participants ($22-58$ years of age; $mean = 37.1 $; $SD = 10.4 $; $26\%$ of them female) employees of [removed for double-blind review] participated in he experiment on a volunteer basis. % Technicolor, Rennes (France). 
All participants have either normal or corrected-to-normal visual acuity.

%facilities and apparatus
%tv
%lumière tamisée
The images were displayed on a 40 inch monitor (TV SONY Bravia ???) with a display resolution of 1,920 $\times$ 1,080. The participants were comfortably seated at a distance of 150 centimeters from the screen (three times the screen height). The 1,024 $\times$ 768 images were centered on a black background; at a viewing distance of 150 cm, the stimuli subtended 18.85 degrees of vertical visual angle.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Procedure}
%Questionnaire part
Participants first answered the questionnaire. During the whole time, the experimenter was near the participant in case he doubt about one movie.

%Algorithm selection
Based on the answers to the questionnaires, the algorithm selected the movies associated with the maximum certitude level (i.e. 5) to constitute  targets and fillers sets. // At the end, the algorithm selected 80 targets and 40 fillers for the participants, and place them in a random order for presentation
// The criteria of selection was the number after the response to the questionnaire the nb of annotations of the sequences, in order to harmonize the nb of annotations per video.
The sequences used as targets were selected among the sequences corresponded to the movies the participants were sure (i.e. rate of 9 in the scale) they have already seen. The algorithm selected the less annotated sequences to try maintaining a number of annotation equivalent between sequences.
This was to maximize the probability that targets really are targets, and fillers really are fillers. according to they number of annotations.
	%three sequences maximum from the same movie! %verif

%Instructions
Then the participants was provided with the instructions about how to complete the memory task. In particular, the experimenter insisted on the fact that we did not want the participant guess that the sequence was il the film they have already seen, but that we want they recognize.

// During about 24 minutes, you will see a series of 10-seconds video scenes. Press the SPACEBAR during the 10-seconds video scene if you RECOGNIZE/REMEMBER seeing it before. Be careful!! You only press the spacebar when you RECOGNIZE/REMEMBER the video scene, not the movie from which the video scene had been extracted. We do not want you guess that the video scene was in a movie you have seen: just press the spacebar if you remember that you saw a particular video scene.

%Recognition task
Then the experimental phase was then launched, corresponding to the memory task.

%Total time
The total time of the experiment was about 50 minutes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{STUDY OF THE MEMORABILITY SCORES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we describe in detail the ground truth data collected thanks to the protocol described in the previous section.
We conclude by calculating the memorability scores finally used to feed our prediction models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{}

%%%
\subsubsection{Sequences' memorabilities}
A sequence's memorability is defined as the percentage of correct detections by participants.
On average, each video was viewed as target by $9.3$ participants. Average sequence memorability was $48.25\%$ (SD of $28.21\%$).
On average, each video was viewed as filler by $9.2$ participants. Average false alarm rate was $5.95\%$ (SD of $14.47\%$). 
%corriger score par moyenne de mem du film -- est-ce que ça a un sens ?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Number of annotations per sequence}
The histogram presented in the figure \ref{fig:nb_annotations_per_sequence} corresponds to the number of sequences for each possible number of annotations.

%Histogram of number of nb of sequences per possibble nb of annotations
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=\columnwidth]{figures/histogram_nb_sequences_for_nb_of_annotations.png}
	\caption{\label{fig:nb_annotations_per_sequence}Nb of sequences for each possible nb of annotations per sequence (>=3).}
\end{figure}

%figure Rank corr for Isola et al. et han et al., pour montrer que le nombre d'annotation sva plus vite pour les videos (mais crowdsourcing vs. laboratoire -- et autres facteurs...)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Consistency analysis}
We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency. It answers the question: "Are the videos that are more memorable (or forgettable) for a group of observers also more likely to be remembered (or forgotten) by a different group of observers? The figure \ref{fig:human_consistency} presents the curve of the human consistency for each video sequence. It answers the question: "How many annotations for each do we need for the human consistency stop varying?'

%Curve of human consistency
\begin{figure}[!htbp]
	\centering
	\subfloat[(a) Human concistency for images in \cite{khosla_2015_understanding}]{\includegraphics[width=0.3\columnwidth]{figures/human_consistency_LaMem.png}}
	\quad
	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/human_consistency_nb_annotations.png}}
	\caption{\label{fig:human_consistency}Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes}.}
\end{figure}

This curve show us that from 14 annotations (using our method) video memorability data has enough consistency

Individual and contextual differences, besides mandatory random variability, explain the $1-.65$ part of the memorability that is not universally derivable from the intrinsic informations of the videos.

%Compare our and Khosla et al. curve dire que -- video + non-crowdsourcing + "real" long-term memory -- explique qu'on obtient avec beaucoup moins d'annotation la même human congruency.
If we compare the curve obtained by \cite{khosla_2015_understanding} (\ref{fig:human_consistency}(a))and the curve we obtained (b), we can note than we attain human consistency far more quickly than Khosla \textit{et al.}. At least three arguments go in the way of this results. First, we work with videos and they work with images ; maybe video memorability is more universal than image memorability because of the higher length of this kind of stimulus (but We can also note that the maximum human consistency is close for images and videos!!). Second, we obtain a measure of real long term memorability, and not a memorability measured some minutes after encoding step ; maybe this measure more representative of what is really memorable. Finally, we can advance the fact that in-lab experiment enable to obtain better memorability measure than crowdsourcing one, resulting to attain maximum human consistency earlier.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generic vs. Typical sequences}
% Tester s'il y a une différence entre les deux et en tirer des conséquence quant à leur validité --> How objective our measure of memorability is?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quality of the movies}
dvdrip vs. HD 720-1080 => Difference of memorability? => Interesting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Response time}
%How to exploit the response time in the prediction models? --> two potential models for memorabilit prediction

The question here is: "Could we exploit response time to correct memorability score?"

We hypothesized that the most memorable the sequences, the faster the participant will answer.

%Memorability of sequence vs. their 
We observed a Person's correlation of $-0.35 (p<.0001)$ between the response time on the target and their memorability scores. This means that the videos with the higher memorability score tended to be answered fastly when correct detection by the participants, suggesting that a sequence most memorable is also a sequence for which we rapidly detect that it is memorable.

The global mean response time was $4.87 sec$ on targets and $5.96 sec$ on fillers.
A Student t-test for different sample size show a significant difference ($t(2836)=-5.34, p<.0001$). This means that the participants globally answered more rapidly for targets (i.e. correct detections) than for fillers (i.e. false alarm), probably because of their hesitation for fillers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolution of the memorability along time}
%Use order positions for that (e.g. mean memorability for position 1, 2... 120)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%tester la mémorability de séquences très proches (prise à intervalles très proche + vérif = même scène)

%indoor/outdorr comme isola et al. mais en contrecarrant l'effet de la couleur -- car ils espliquent que c'est par la corrrélation entre couleur et mémorabilité est expliqué par le fait que les couleurs plus froides sont d'exté.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New manner to compute memorability scores: take into account time and FA}
%Une nouvelle manière pour prendre en compte les FA dans les scores de mémorabilité applicable à l'étude de la mémorabilité sur les images (si fonction : en parler dan sl'intro comme quoi les gens ne prennent jamais en compte les FA quoiqu'elles puissent être exploiter)
%Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.

%%%
\subsubsection{Participants' performance}
% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the film --> No' --> So the point here is just to provide an overview.
The average memory performance was the following: the average percentage of correct detection was $48.2\%$ (SD of $14.1\%$) and the average percentage of false alarms was $4.78\%$ (SD of $5.63\%$).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Logistic regression vs. SVM to personalize prediction model}
%Regression qui prend en compte tous les facteurs (voir ma thèse)
%See QoMEX paper to present the data

%We used the following factors:
	%Occidental/non-occidental, because the 100 movies are occidental (expept slumdog millionaires...)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Other results possible
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Film genre and IMDB ratings}
%test the momorability od the sequences compare to the mean one of the movie.
%Correct bu the mean participant's performance
%And maybe other IMDB annotations/labels

\subsection{Context}
%Analyse the context (order of video presentation) influent the memo: to be check if we have enough information?
%Une vidéo par rapport à toutes les autres différentes à un impact sur sa mem ?
%S'inspirer de la technique de Bylinskii

\subsection{Features linked to memorability}
%tester ces features comme dans 'What makes a photograph...'
% Indoor/Outdoor, low-level visal feautures (color... -- i.e. as in isola et al.), salinecy maps

\subsection{Indoor \textit{.vs} outdoor scenes}

%add other relevant points found in MIT papers e.g.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Memorability score calculation}
According to what was presented before in tis section, this is how we finally decided to compute our memorability scores...

%quality of the movies: dvdrip vs. HD 720-1080 => Difference of memorability? => Interesting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MEMORABILITY PREDICTION} %for Khartik
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%we used Soundnet...
% Nous construisons actuellement une base de données de grande ampleur avec une mesure purement objective de la mémoire.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{DISCUSSION}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cite QooMEX for personlization of memorability model

%D'abord, le test de mémoire interroge une mémoire de vidéos vues longtemps (en moyenne, probablement plusieurs mois ou années) auparavant.

%file to make the data + movies names available.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

%%%
 %Liste des features (comme/ceux de MediaEval 2015 à donner au téléchargement avec le début en sec de nos sequences)
 %Release text file for movie title, start/end time + features

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Appendixes should appear before the acknowledgment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{ACKNOWLEDGMENT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This section is optional; it is a location for you
%to acknowledge grants, funding, editing assistance and
%what have you.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BIBLIOGRAPHY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2018-01-09 ARTICLE.tex
Draft 2 of the paper

@@
 \begin{abstract}
-Following the interest of these last years for image memorability prediction, video memorability has recently attracted attention of researchers.
-The growing number of video shared everyday forces us to find new way to deal with them to make the most relevant their occurrence in our everyday lives.
+Following on from the study of image memorability prediction, which draw continuous attention over the past six years, the computational understanding of video memorability's search field has recently hatched.
+The growing number of shared videos foster us to find new ways to make their occurrence the most relevant in our everyday lives.
+There is no available dataset of videos annotated in terms of memorability; such dataset may probably result in a launching of the field, as it had been the case for images.
+The first goal of the pioneers of this search should be to succeed to propose a protocol to constitute such a dataset. 
+%our work:
+In this article, we propose a protocol of 700 videos annotated with memorability scores, constitute a dataset, study the quality of the collected data, and computationally modeling memorability to predict it.
+%long-term memory
 
-Recently, two studies have tried to predict video memorability ; however, the immaturity of this work is obvious.
-In particular, none of these studies on images nor videos tried to measure 'very' long-term memory: they just measured the memory of items some minutes after the memory encoding.
-In this paper, we propose a dataset of 700 videos annotated with memorability scores corresponding to real long-term memory.
-We discuss in details the method and the proponents.
-We then propose a model to predict memorability of videos and test it on the available dataset of images and videos.
-% Our approach improves over current memorability methods in this task.
-The material -- videos plus theirs memorability scores -- is made accessible for researchers.
+We finally propose a deep-learning model base to predict video memorability, comparing several methods and class of features.
+
 \end{abstract}
@@
 %citer Deep learning emotional images bias
-\paragraph{Opening} % oder Ok
-To deal with the constant increase in shared videos, we must imagine new ways to organize -- in particular, to retrieve -- videos with the aim of improving the relevance of their occurrences in our every day lives.
-Contrary to other metrics used to quantify image of video importance, such as aesthetics or interestingness, whose corresponding ground truths corresponds to subjective judgments, memorability can be objectively measured by a memory test.
-As such, memorability can be regarded as a particularly relevant element to help us picking a video among several ones.
 
-\paragraph{The field of video memorability is very immature}
+%%%%%%%%%%%%%%%%%%%%%%
+%Opening %Ok
+Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- contents.
+Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as a particularly relevant element to help us picking a video among several ones, with the advantage of a possibility to be measured less subjectively.
+
 %from images to video memorability
-The image memorability prediction challenge has attracted increasing attention from the seminal work of Isola \textit{et al.} released in 2001 \cite{isola_2011_makes}.
-Recently, models achieved very good results with the introduction of deep learning to address the challenge \cite{khosla_2015_understanding,baveye_2016_deep}% plus papier Hammad
-As a result of this success, the search field has been extended to videos.
-Video memorability prediction is however a very new field, and there are only two studies, to our knowledge, that address this issue \cite{han_2015_learning,shekhar_2017_show}.
+The computational understanding of video memorability's search field follows on from the study of image memorability prediction which has attracted increasing attention from the seminal work of Isola \textit{et al.} released in 2001 \cite{isola_2011_makes}.
+Recently, models of image memorability prediction achieved very good results with the introduction of deep learning to address the challenge of image memorability prediction \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
+This success resulted on the extension of the challenge to videos.
+Video memorability prediction is however a new field, and there are only two studies, to our knowledge, that addressed this issue \cite{han_2015_learning,shekhar_2017_show}.
 
+%No dataset available
 At least two important problems could explained this scarcity of studies.
-Firstly, protocols used in the existing studies are not smart enough, but are easily critisizeable.
+Firstly, protocols used in the existing studies are not smart enough, but are questionable.
 Secondly, authors didn't make their data accessible for community, so no database exist.
-
-\subsection{Problem 1: The problem of the weakness of the studies}
-%%% Critiques of previous studies %%%
-
-\paragraph{Critics of Han et al.}
-The previous attempt at creating memorability video dataset first approach to model video memorability had been propose by \cite{han_2015_learning}. 
-This first attempt is a "strange" adaptation of \cite{isola_2011_makes} protocols.
-
-% A memory game of 12 hours
-- Firstly, the authors call their task a "memory game" as Isola \textit{et al.}.
-If it were justified for Isola et al., 2011, la tâche dure pour 1 unique participant environ 12 heures.
-The authors do not provide this time, but it cans be deduced from the by the following calculation:
-
-Thenks to this maneer, participants obtained a dataset of videos with only 20 participants (we don\'t know if they are student or payed).
-
-répatit sur 10 jours diffférens (5 pour les phases dapprentissage + 5 pour les phase de mesure de la mémoire).
-
-uen autre aberration e
-
-%critique of Shekhar et al.
-In \cite{shekhar_2017_show}, 
-- des questions pas forcément aussi faciles/diffilec entre les vidéos, qui peuvent expliquer leur
-
-Enfin, dans les deux études précédentes, la congruence inter-humaine, minimum nécessaire pour s'assurer de la cohérence des données et avoir une base de comparaison pour les modèles, n'a pas été fournie.
-
-%Lasting long-term memory for IMAGES - critic of Isola et al.
-on comprend les auteur d'avoir agit ainsi pour consituter une base de grand ampleur (Khosla et al.)
-It is long-term memory a proprement parlé because.
-but memories evolove in l-g mem
-=> Ebbinghaus curve
-So memorabilities of souvenirs can change after a certain delays
-%as proposed in [ma thèse] 
-%Memorability continue to evolve in long-term memory --> Ebbinghaus -- or more recent similar -- curve
-%Memorability image/videos in computer vision never interested in REAL long-term memory
-%We want to correct this problem by proposing a new way to measure memorability.
-% saut de la tâche d'apprentissage (remplacée par le questionnaire)
-% To sum up, we want to obtain the "long-lasting memorability" of the videos
-
-\paragraph{Problem 2: No existing database}
-%there is no database
-=> Dataset non disponible pour les deux études.
-
 The fact that there exist no dataset to modeling video memorability is the most serious obstacle which prevent video memorability prediction to take off.
@@
 The second released database, far larger, launched new possibility for deep learning and enable to obtained the aforementioned really good results \cite{khosla_2015_understanding}.
-Nous pouvons parier qu'il en sera de même pour le cham de la mémorabilité des vidéos.
+Nous pouvons parier qu'il en sera de même pour le champ de la mémorabilité des vidéos.
+
+%A dataset as first objective to launch VM prediction --> Goal of this paper is to participant in th eestablishment of an adapted protocol
+We think that the existence of such a database could have the importance for video memorability that the ones of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} had on image memorability.%in particular for deep learning algortihm which perform the best for images
+This article wants to takes his part to develop such a protocol to collect high quality data, and such a dataset.
 
@@
 The establishment of a dataset, in particular of the first dataset of this kind, may not be self-evident.
-
 Elle nécessite de définir les points cardinaux qui vont guider – et contraindre – les chercheurs qui utiliseront ces premières données disponibles.
@@
 
-%Conclusion paragraph: we want to contribute to the protocol creation by proposing one method
-Aussi cette étude veut participer à réfléchir aux protocoles qui pourraient être employée pour contituer à l'avenir à un protocole de plus grande envergure.
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Previous work} %Critics of the 2 previous studies
+%Critics of studies on image memorability : the problem of sustainable long-term memorability
+The work on image memorability prediction, initiated with \cite{isola_2011_makes}, achieved great results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
 
-\paragraph{Goals of our study}
-%link the two reasons precebndtky evoque
-These two reasons are linked. They give us our priority: build a dataset and define a protocol to do that.
-This dataset -- and a fortiori the protocol -- should avoid the drawbacks of previous work in video and image memorability.
-%in image memorability, difficultys come from que'elle ont justement utilisés ces bases de données.
+Probably constraint by the difficulty to perform large long span studie in crowdsourcing, this lasts necessary to build a annotated database large enough to meet dep-learning, the authors of the major existing dataset \cite{isola_2011_makes,khosla_2015_understanding} measured memory some minutes after learning.
+Or, this could be a drawback, as propose and showed in \cite{cohendet_2016_prediction}.
+Indeed, long-term memorability continue to evolve gretly in the first day of long-term memory [Ebbinghaus's curves], which could change the order of memorability of images, alors que les goals proposés pour la mémorabilité des images et vidéos supposent d'être capables de prédire une mémorabilité à réel long terme.
 
-%2 resaons => our goal of our study => correct these drawbacks
-=> To propose a new protocol to collect data/for annotation
-=> To propose a dataset
-=> The dataset will have "lasting" long-term memorability annotations
-We want our work would benefit to video but also to image memorability.
+It would be cool to have a protocole that measures what we will later refer to as "long-term" memorability.
 
-%Ainsi, puisque les limitations d'un protocole destiné à collecter des données tendent à se répercuter dans les travaux qui utilisent celles-ci, un protocole destiné à constituer la base nécessaire pour répondre au défi de la prédiction de la mémorabilité des vidéos devra répondre à de strictes exigences de qualité.
-La qualité des données obtenues, primordiale, nécessite une connaissance profonde des phénomènes que ces données sont censées refléter.
-La mémoire étant un phénomène étudié (depuis plus d'un siècle) par la psychologie, et les modèles computationnels permettant de l'inférer des caractéristiques des vidéos appartenant au domaine de l'informatique, l'approche transdisciplinaire semble la plus adaptée dans ce cas.
+%Critics of Han et al.
+To our knowledge, the first attempt at predicting video memorability had been proposed by \cite{han_2015_learning}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\paragraph{Secondary goals}
-%%% OTHER INTERSTING RESULTS
-%will depend on our results
-%\paragraph{Exploiting response time for memorability scores}
-%D'autre part, nous voulons également estimer si le temps de réponse à un effet sur la mémorabilité, c'est-à-dire qu'il pourrait être pris en compte pour corriger les scores calculer.
-Note hypothèse est que, plus les participants sont sûrs d'eux, plus ils vont tendre à répondre rapidement.
-Att!!
-peut pas être tester directement : We hypothesized that the most memorable the sequence, the faster the participant will answer.
+The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
+In contrast to the "memory game" proposed by Isola \textit{et al.} to collect memorability data, their protocol is however much heavier. 
+They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
+
+We can infer from the sparse information authors gave that the task duration, for each of the 20 participants, was about 24 hours (by taking an average video duration half-way between 15 and 30 sec), spread over 10 sessions (5 free viewing and 5 recall tasks) of about 2 hours each.
+
+Authors used the same proportion of fillers in the free viewing and recall tasks (4/5 of fillers and 1/5 of targets), to "to guarantee that viewers are unaware of targets".
+If it was totally justified in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers (never reasked videos) in free viewing task would have very week impact on memorability scores (généralement on se passe même de fillers dans les tâches d'apprentissage du matériel).
+
+Still, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
+
+Regarding the prediction part, we also agree with these authors to say that "the method used fMRI measurements for predicting memorability, which would be difficult to generalize".
+
+%Critics of Shekhar et al.
+Another earlier approach was the one of \cite{shekhar_2017_show}.
+The participants passed in crowdsourcing a free viewing task followed by a recall task, for which they had to answer question (such as: "Did you see a man juggling?", "Did you see a car on road?").
+
+%the problem of the questions
+The major drawback of the study comes from this use of questions instead of a recognition task.
+The "memorability" measured may be due not to memorability of the videos but (also) to the differences between the questions rather than a difference of memorability of the videos.
+This bias could be of several kinds: difficulty, comprehension (non anglophone people in AMT), easility of imaging), indexage plus facile...
+C'est d'autant plu sproblématique que ces diffénrece could have drives to a high inter-human consistency not due to memorability attribué à tort à un phénomène de mémoire.
+C'est d'autant plus problématique que le temps de réponse des partcipants entre dans le calcul de la mémorabilité, ce qui ne peut q'uaccentuer ces biais, et est porbablement responsable de la high human response consistency they observe.
+
+Le fait que les questions doivent être créé à la main, et que les choix des questions sont assez limitatifs (par ex avec la question "Did you see a car on road?", il ne peut y avoir de toute la tâche qu'une seule fois une voiture) makes difficult the generalization of this protocol to build an extensive dataset.
+“It was manually ensured that no two textual questions nor any two videos in a sequence were similar in content.”
+
+%Conslusion critics
+Les critiques exposés montre à quel point il est difficile de construire la première base de données pour mesurer la mémorabilité des vidéos.
+Isola et al.
+S’en sont très bien sortis, en se cantonnant à une mesure purement objective de la mémoire – tâche de reconnaissance classique –, ce qui leur a permis de lancer le champ de recherche de la mémorabilité des images.
+Il faudrait en faire de même pour la mémorabilité des vidéos.
+Cette étude s’inscrit dans cette volonté – non pas de créer un nouveau datasets, mais d’essayer un protocole pour voir s’il est efficient.
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Goals of our study}
+%1/ we want a protocol
+To sum up, we want a protocol:
+- which could be extended to create a large dataset in case it shows its quality
+
+- that measure a long-term memorability: as a longitudinal study would be difficult to generalized to create a large dataset, we decide to measure the memoray of participants created previously to the experiment.
+
+%2/ we want an all-around paper
+D'autre part, all around paper
+
+%3/ we want to meet the challenge of video memorability prediction
+Finally, we want -- thanks to our protocole -- to participate in the VM prediction challenge.
+
+%Conclusion intro
+The rest of the article is as follows: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and part 5 the conlusion and intended future work.
+
+%%%%%%% POTENTIALLY %%%%%%%
+% Response time: must we exmploit it to calculate memorability scores?
+ %Because of \cite{shekhar_2017_show} and other papers in images which suppose that short time response equal to more memorable ALORS qu'ils n'ont pas pour tâche de répondre le plus rapidement possible [vérifier].
+ %La répartition du temps de réponse pourrait ne pas linéraire.
+	%We hypothesized that the most memorable the sequence, the faster the participant will answer.
 Parce que les video les mieux représentés dans l'esprit sont les plus accessibles et que les participants ne veulent pas louper la préponse quand ils sont sûrs.
@@
 
-%\paragraph{Personalization of the data}
-%Taking into account human factors
-	%see results Logistic regression vs. SVM
-
-il semble des dernière études qu'il serait intéressant de prendre en temps le dans le calcul de la mémorabilité [LaMem + papier sur la VM où ils prennent en compte le temps dans leurs scores]
-
-%If results Repetition vs. memorability
-The semantization that can affect long-term memorability of videos
+% Personalization of the data
+ %Taking into account human factors %see results Logistic regression vs. SVM
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Opening -- Ok for now
-In this section, we describe the new protocol we propose to measure video memorability in order to collect the long-lasting memorability of the videos.
-The protocol is also lighter than ordinarily since there is no need of learning step to encode the material before the recognition task, as we measure the participants' memory established prior the experiment.
-
-The memory task involve three materials: videos sequences, questionnaire and participants.
-
-%Insist on Contribution about the protocol (novel: how to measure long-term memory)
+Our first step was to implement a protocol to measure the long-term memorability of 700 video sequences, extracted from 100 mainstream films.
+The main characteristic of the protocol, in contrast with previously proposed methods to collect image and video memorability annotations, is the absence of learning/free viewing task, replaced by a questionnaire designed to collect information about the participants' prior memory.
+The way of proceeding aimed at measuring memory performance after a significant retention period.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Memory task}
-To measure the "real" long-term memory, we designed the recognition task presented in the figure ...%ref.
-%figure Recognition task
+\subsection{Dataset}
+%% films
+We mixed films' popularity and genres to establish our list. 
+%The movies quality was HD 720-1080, except for ??? dvdrip movies.
 
-% Targets and Fillers
-The 10-sec video sequences are displayed one after another, separated by an inter-stimuli interval (ISI).
-
-%define target and filler
-Les vidéos cibles sont mêlées aux vidéos de remplissage, et le participant doit dire, pour chaque vidéo qui lui est présentée, s’il l’a déjà vue ou non.
-To answer, he must press the spacebar during the displaying of the video sequences (than answer when they arrived during the 1-s inter-stimuli interval).
-movies they have seen (the targets) and never seen (the fillers).
-Even if the participant answers, the video continue to run until its end (to avoid the temptation of ansewering to faster finish the experiment).
-
-"Respond ONLY if you remember the sequence AND NOT THE movie from which it comes".%reprendre phrase que j'ai écrite
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The video sequences}
-%% movies
-A list of 100 movies (available with the downloadable data) was first established, for which we mixed their popularity and their genres.
-The movies quality was HD 720-1080, except for ??? dvdrip movies.
 orginal-english version (without subtitles) (occidental movies, except for Slumdog millionaire) %to check
@@
 %%video sequences
-To constitute our dataset, we manually selected 7 video sequences of 10 seconds from each movies (for a total of 700 videos).
+We then manually selected 7 video sequences of 10 seconds from each movies (for a total of 700 videos).
+
 The sequences had to meet several criteria to be selected.
@@
 To maintain a relative semantic consistency intra-sequence, the cutters were provided with the instruction to avoid to cut in the middle of a sentence, of to agglomerate very different plans together.
-The idea was to increase the coherence of the sequences to diminish the difficulty of he computational models to learn without impairing ther capacity to generalize.
+The idea was to increase the coherence of the sequences to diminish the difficulty of he computational models to learn without impairing their capacity to generalize.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Questionnaire}
-The targets and fillers that constitute the items of the recognition task were different for each participant.
-Indeed, the video sequences viewed by a participant were selected based on response to a questionnaire presented in the figure ???
-Questionnaire with the 100 movies.
+\subsection{Survey design}
+%questionnaire
+After they had provided basic demographics, participants filled out a questionnaire composed of 100 items, indicating whether they had seen the films of our list.
+Each item included a film poster with its name below, followed by the question: "Do you remember watching (fully) this movie?" The experimenter was present in case the participant had doubt about one movie.
+If the participant answered "yes", he had to answer three additional questions: "Please let us know how confident you are of having seen this movie." (I am not/slightly/50\%/considerably/100\% confident); "Please let us know when did you last see this movie." (<1 month; <1 year; <5 years; < 10 years; > 10 years); "Please let us know how many times did you watch this movie." (once; 2-4 times; 5-9 times; 10-19 times; >20 times).
+If he answered "no", he had to answer one more question: "Please let us know how confident you are of NOT having seen this movie". (I am not/slightly/50\%/considerably/100\% confident).
+The question relative to confidence aimed to ensure the targets and fillers corresponded to seen and non-seen films respectively.
+The other questions (i.e.
+when and how many times the films had been seen) were used during result analysis to evaluate the likely effects of visualization repetition and passage on long-term memorability.
+The questionnaire required about 20 minutes to complete.
 
-%figure Questionnaire
+%Algorithm selection
+Based on the answers to the questionnaire, an algorithm selected 80 targets and 40 fillers among the films associated with the highest degree of certitude, with a maximum of two sequences from the same film.
+Another criterion of selection for the target was the number of annotations of the sequences, in order to harmonize the nb of annotations per video.
+The sequences were presented in a random order for presentation
 
-Le nom du film, suivi de l’affiche, du nom du réalisateur et des acteurs principaux – qui permettent de s’assurer que le film que le participant a en tête est bien celui qui lui est présenté – est présenté au participant.
-
-%Other questions associated with the questionnaire
-For each movie, the participant had first to answer the following question: "Do you remember watching this movie?" -- with the fact that was madde clear that the film must been seen EN ENTIER.
-
-If he answered "yes"
-Please let us know how confident you are of having seen this movie (I am not/slightly/50\%/considerably/100\% confident)
-Please let us know when did you last see this movie (<1 month; <1 year; <5 years; < 10 years; > 10 years)
-Please let us know how many times did you watch this movie (once; 2-4 times; 5-9 times; 10-19 times; >20 times)
-
-If he answered "no"
-Please let us know how confident you are of NOT having seen this movie (I am not/slightly/50\%/considerably/100\% confident)
-
-The goal of the confidence question was to be sure that we select video sequences from movie , to increase our certitude that target/filler were really targets/fillers.
-
-The other two questions were used for the analysis of the results to assess factors a priori susceptible to influence video memorability.
-The question about the number of times the movie had been seen for us to evaluate the effect of repetition on memorization, and the question about the last time tha movie had been seen to evaluate the effect of time passage on long-term memorability, that continue to be affected in longterm memory %citer un papier de moi
+%Recognition task
+The recognition task lasted 24 minutes. Participants saw 10-seconds sequences separated by an inter-stimuli interval of 2 second. Their task was to press the spacebar when they recognized a sequence.
+During the instruction, the experimenter insisted on the fact that the only task of the participant was to answer only when he recognized the particular sequence, but not answer if he recognized not even when he could guess that the sequence came from a film he had seen.
+This, because for number of sequences (i.e.
+the typical ones), it was possible to guess that the sequence came from a particular film.
+Even if the participant answers, the video continue to run until its end.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Participants, facilities and apparatus}
-%participants
-???
-(105) participants ($22-58$ years of age; $mean = 37.1 $; $SD = 10.4 $; $26\%$ of them female) employees of [removed for double-blind review] participated in he experiment on a volunteer basis.
+\subsubsection{Participants, facilities and apparatus}
+104 participants ($22-58$ years of age; $mean = 37.1 $; $SD = 10.4 $; $26\%$ of them female), employees of [removed for double-blind review] participated in he experiment on a volunteer basis.
 % Technicolor, Rennes (France).
 All participants have either normal or corrected-to-normal visual acuity.
-
-%facilities and apparatus
-%tv
-%lumière tamisée
-The images were displayed on a 40 inch monitor (TV SONY Bravia ???) with a display resolution of 1,920 $\times$ 1,080.
-The participants were comfortably seated at a distance of 150 centimeters from the screen (three times the screen height).
-The 1,024 $\times$ 768 images were centered on a black background; at a viewing distance of 150 cm, the stimuli subtended 18.85 degrees of vertical visual angle.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Procedure}
-%Questionnaire part
-Participants first answered the questionnaire. During the whole time, the experimenter was near the participant in case he doubt about one movie.
-
-%Algorithm selection
-Based on the answers to the questionnaires, the algorithm selected the movies associated with the maximum certitude level (i.e.
-5) to constitute  targets and fillers sets.
-// At the end, the algorithm selected 80 targets and 40 fillers for the participants, and place them in a random order for presentation
-// The criteria of selection was the number after the response to the questionnaire the nb of annotations of the sequences, in order to harmonize the nb of annotations per video.
-The sequences used as targets were selected among the sequences corresponded to the movies the participants were sure (i.e.
-rate of 9 in the scale) they have already seen.
-The algorithm selected the less annotated sequences to try maintaining a number of annotation equivalent between sequences.
-This was to maximize the probability that targets really are targets, and fillers really are fillers. according to they number of annotations.
-	%three sequences maximum from the same movie! %verif
-
-%Instructions
-Then the participants was provided with the instructions about how to complete the memory task.
-In particular, the experimenter insisted on the fact that we did not want the participant guess that the sequence was il the film they have already seen, but that we want they recognize.
-
-// During about 24 minutes, you will see a series of 10-seconds video scenes.
-Press the SPACEBAR during the 10-seconds video scene if you RECOGNIZE/REMEMBER seeing it before.
-Be careful!!
-You only press the spacebar when you RECOGNIZE/REMEMBER the video scene, not the movie from which the video scene had been extracted.
-We do not want you guess that the video scene was in a movie you have seen: just press the spacebar if you remember that you saw a particular video scene.
-
-%Recognition task
-Then the experimental phase was then launched, corresponding to the memory task.
-
-%Total time
-The total time of the experiment was about 50 minutes.
+The videos were displayed on a ?????
+inch monitor (TV SONY Bravia ?????) with a display resolution of 1,920 $\times$ 1,080.
+The participants were seated at a distance of 150 centimeters from the screen (three times the screen height) in a room equipped with subdued lights.
+%The 1,024 $\times$ 768 videos were centered on a black background; at a viewing distance of 150 cm, the stimuli subtended 18.85 degrees of vertical visual angle.
 
@@
 \subsection{}
+
+%We finally work on 660 movies (sequences annotated at least 4 times).
 
@@
 	\quad
-	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/human_consistency_nb_annotations.png}}
-	\caption{\label{fig:human_consistency}Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes}.}
+	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/Measure_of_human_consistency.png}}
+	\caption{\label{fig:human_consistency}Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error).}
 \end{figure}
 
-This curve show us that from 14 annotations (using our method) video memorability data has enough consistency
+%compounding factor: most annotated sequences are also the most memorable group, which could artificially increase the Spearman's correlation).
 
@@
 \subsection{Response time}
-%How to exploit the response time in the prediction models? --> two potential models for memorabilit prediction
+Previous authors have integrated response time in their score of memorability.
+We pose here the following question: "Is the degree of memorability related to response time?" to answer the question: "Should we exploit response time to correct memorability score?" We hypothesized that the most memorable the sequences, the faster the participant will answer (no instructions to answer faster: just to answer durgin the 10-sec display).
 
-The question here is: "Could we exploit response time to correct memorability score?"
+%figure Mean response time vs. degrees of memorability
+	%with SEM
+\begin{figure}[!htbp]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/Response_time_vs_Memorability.png}
+	\caption{\label{fig:response_time_vs_memorability}Mean response time for each memorability degree (error bars correspond to SEM).}
+\end{figure}
 
-We hypothesized that the most memorable the sequences, the faster the participant will answer.
+We observed a Person's correlation of $-0.36 (p<.0001)$ between the response time on the target and their memorability scores.
+This is consistent with the figure \ref{fig:response_time_vs_memorability}.
+In other words, the higher the memorability, the lower the response time tends to be, suggesting that a sequence most memorable is also a sequence for which we rapidly detect that it is memorable.
+This is consistent with what we expected: empirically, we observe during passation that participant tend to answer as soon as they recognized the sequence.
 
-%Memorability of sequence vs. their 
-We observed a Person's correlation of $-0.35 (p<.0001)$ between the response time on the target and their memorability scores.
-This means that the videos with the higher memorability score tended to be answered fastly when correct detection by the participants, suggesting that a sequence most memorable is also a sequence for which we rapidly detect that it is memorable.
-
+%target vs. fillers
+It is interesting to link it to the response time for answer on targets (i.e.
+correct detections) vs.
+on fillers (i.e.
+false alarms).
 The global mean response time was $4.87 sec$ on targets and $5.96 sec$ on fillers.
@@
 correct detections) than for fillers (i.e.
-false alarm), probably because of their hesitation for fillers.
+false alarm).
+One explanation is that participants hesitated more for fillers they answered on, increasing their response time.
+
+It tells us something on memory, that the memories are not evident but more blurred.
+This is an important point, taht explained why, in our discussion, we propose to turn to a "totally objective" measure of memory to constitute a dataset for video memorability prediction: if the participants hesitated, so they choice to answer or not was probably impacted by the way they answered the instructions and their own sensibility to the risk.%=> corrigé mémorabilité par individu ?
+
+%How to exploit the response time in the prediction models? --> two potential models for memorabilit prediction
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%we used Soundnet...
-% Nous construisons actuellement une base de données de grande ampleur avec une mesure purement objective de la mémoire.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{DISCUSSION}
+\section{CONCLUSIONS AND FUTURE WORKS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Nous construisons actuellement une base de données de grande ampleur avec une mesure purement objective de la mémoire.
+%Nulle part les auteurs ne parlent de mémoire à long terme, à court terme… => Ils appréhendent un problème par nature interdisciplinaire d’un point de vue purement computer science => Or la qualité des données est ce qui détermine ce que les modèles vont finalement prédire.
+// %D'abord, le test de mémoire interroge une mémoire de vidéos vues longtemps (en moyenne, probablement plusieurs mois ou années) auparavant.
 %cite QooMEX for personlization of memorability model
 
-%D'abord, le test de mémoire interroge une mémoire de vidéos vues longtemps (en moyenne, probablement plusieurs mois ou années) auparavant.
-
-%file to make the data + movies names available.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{CONCLUSIONS}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % A conclusion section is not required.
@@
  %Release text file for movie title, start/end time + features
+
+%file to make the data + movies names available.
+
+%In particular, our protocol try to correct to integrate the concept of "very long-term memory", that escaped to previous studie in image and videos, which measured the memory of items some minutes after the memory encoding.
+
+%anntation des séquences pas à même vitesse, en cas de généralisation : changer les séquences annoté plus de 15 fois par exemple par de nouvelles.
+
+% Conclusion => Nos travaux futurs
 

### 2018-01-11 ARTICLE.tex
Draft -- Add som text for intro + for dataset construction + some results for Analysis of the collected data part

@@
 %%%%%%%%%%%%%%%%%%%%%%
-%Opening %Ok
+%Opening
 Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- contents.
@@
 %Critics of studies on image memorability : the problem of sustainable long-term memorability
-The work on image memorability prediction, initiated with \cite{isola_2011_makes}, achieved great results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
+The work on image memorability prediction, initiated by \cite{isola_2011_makes}, produced great results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
+The availability of two large datasets \cite{isola_2011_makes,khosla_2015_understanding} has been critical in this achievement. 
+However, possibly constrained by the difficulty in conducting long crowdsourcing studies, the authors measured memory performance a few minutes after memorization to obtain their memorability annotations.
+As shown in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
+Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation \cite{mcgaugh_2000_memory}.
 
-Probably constraint by the difficulty to perform large long span studie in crowdsourcing, this lasts necessary to build a annotated database large enough to meet dep-learning, the authors of the major existing dataset \cite{isola_2011_makes,khosla_2015_understanding} measured memory some minutes after learning.
-Or, this could be a drawback, as propose and showed in \cite{cohendet_2016_prediction}.
-Indeed, long-term memorability continue to evolve gretly in the first day of long-term memory [Ebbinghaus's curves], which could change the order of memorability of images, alors que les goals proposés pour la mémorabilité des images et vidéos supposent d'être capables de prédire une mémorabilité à réel long terme.
+In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance is particularly strong in the few days immediately following the memorization.
 
-It would be cool to have a protocole that measures what we will later refer to as "long-term" memorability.
+, with a memory performance in recall that asymptotically tends its long-term performance \cite{ebbinghaus_1913_memory}.
+
+Because several factors are susceptible to influence the consolidation process (such as the emotional content, the semantics, the attention...), which is not a linear decrease for all the memories, it follows that the order of memorability of images is susceptible to change from some minutes to a day after memorization, as it has been found in \cite{cohendet_2016_prediction} with emotional image from of the International Affective Picture System \cite{lang_2005_international}.
+
+A protocol to collect memorability annotations for videos would benefit from the capacity to measure what we will later refer to as "long-term memorability".
 
@@
 %3/ we want to meet the challenge of video memorability prediction
-Finally, we want -- thanks to our protocole -- to participate in the VM prediction challenge.
+Finally, we want -- thanks to our protocol -- to participate in the VM prediction challenge.
 
 %Conclusion intro
-The rest of the article is as follows: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and part 5 the conlusion and intended future work.
+The rest of the article is as follows: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and part 5 the conclusion and intended future work.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Our first step was to implement a protocol to measure the long-term memorability of 700 video sequences, extracted from 100 mainstream films.
+Our first step was to implement a protocol to measure the long-term memorability of 700 video sequences, extracted from 100 mainstream movies.
 The main characteristic of the protocol, in contrast with previously proposed methods to collect image and video memorability annotations, is the absence of learning/free viewing task, replaced by a questionnaire designed to collect information about the participants' prior memory.
@@
 \subsection{Dataset}
-%% films
-We mixed films' popularity and genres to establish our list. 
-%The movies quality was HD 720-1080, except for ??? dvdrip movies.
+%% films and sequences
+We mixed popularity and genres to establish our list of movies.% occidental movies, except for Slumdog millionaire)%to check%(list added to the appendix [ref_appendix])
+Then, we manually selected seven sequences of 10 seconds from each movie (for a total of 700 sequences).
+%with HD or dvdrip quality, in orginal (i.e.english) version, without without subtitles [add if results HD vs.
+DVDrip]
 
-orginal-english version (without subtitles) (occidental movies, except for Slumdog millionaire) %to check
+%cuts
+We made efforts to maintain a high intra-sequence semantic cohesion; in particular, we did not cut through meaningful vectors nor we agglomerated shots that belonged to different scenes.
+Indeed, we think the images' characteristic by which they constitute obvious semantic units is an important element in the capacity of the prediction models to generalize.
+By contrast, the meaning a sequence conveyed was susceptible to change depending on where we decided to cut the movie.
 
-%%video sequences
-We then manually selected 7 video sequences of 10 seconds from each movies (for a total of 700 videos).
+Because the semantics is linked to the memorability of images \cite{isola_2014_makes}, so probably to videos, therefore their memorabilities, we chose this definition susceptible to be extensive to predict other dataset.
 
-The sequences had to meet several criteria to be selected.
-First, not to be part of the official trailers, to be sure that, if a participant report not having seen the movie, he won't have seen the sequence presented for the recognition in a trailer.
-Second, we wanted our sequences to constitute "logical" units, that is to say to "protect the sens" of the videos.
-Indeed, semantics is linked to memorability of images \cite{isola_2014_makes}, from which we can imagine that it is also the case for videos.
-But, contrary to images that are clearly cut units, sequences memorability depends on the cut.
-To maintain a relative semantic consistency intra-sequence, the cutters were provided with the instruction to avoid to cut in the middle of a sentence, of to agglomerate very different plans together.
-The idea was to increase the coherence of the sequences to diminish the difficulty of he computational models to learn without impairing their capacity to generalize.
+%hence, in our idea, fostering semantics coherence of the sequences equal to diminish the difficulty of the computational models to learn without impairing their capacity to generalize.
 
-%%Generic vs. typical
-During the manual selection of the sequence, we chose 127 sequences (among the 700) we called "generic" by contrast to "typical" ones.
-Generic sequences are parts of the movie which do not contain any elements that would enable someone to easily guess that the sequences belong to the particular movie.
-The list of undesirable includes but is not limited to: scenes with recognizable famous actors (e.g.
-in Kill Bill, avoid the scenes with Uma Thurman, Lucy Liu, David Carradine…); scenes famous for their music, gesture...
-(e.g.
-in Kill Bill, avoid the scenes with music that enable to easily guess that the sequence comes from Kill Bill); scenes with typical elements (e.g.
-a spaceship or an alien from which you can easily guess that you are watching a Star Wars movie).
-The cutters were provided with the instructions to ask themselves the following question when choosing a generic sequence in a movie: “Could I easily guess this sequence belongs to the movie I extracted it from?” -- if the answer was yes, so don’t select this sequence.
-In some movies (e.g.
-Kill Bill 1, Star wars, 2001 A Space Odyssey, etc.), just a few 10-sec sequences of no sequence at all were found to meet these criteria (e.g.
-The lord of the ring); for these movies, we don't have generic sequences.
-By contrast, non-generic sequences were namely two 10-sec parts of the movie that you think are very representative of the movie (you can choose any part of the movie without restriction).
-These two types of sequences should be considered in link with the instruction of the task written above (even if you know that a sequence comes from a movie, just answer if you recognize THE sequence): it is just a supplementary control (qui jour un rôle symétrique au taux de FA pour la prise de risque).
+%neutral vs. typical
+We also gave preference to the sequences we called "neutral", by contrast to the "typical" ones.
+According to our definition, a neutral sequence is a part of a movie which contains no element that would enable someone to easily guess it belongs to a particular movie.
+
+The list of undesirable includes but is not limited to: recognizable famous actors, typical music, style, or any element...
+
+The question we asked ourselves to select the neutral sequence was the following: "Could I guess this sequence belongs to the movie I extracted it from if I didn't know it?"
+
+In most movies, just a few or no 10-sec sequences of this sort exist. For example, in Kill Bill 1, we can
+
+That explains why, at the end, we obtained only 127 neutral sequences for 573 typical ones, that are defined as being non neutral sequences that meet the first criterion.
+
+These two types of sequences -- neutral \textit{vs.} typical -- should be considered in conjunction with the instruction of the task:
+
+neutral sequences are control.
+
+ of the task written above (even if you know that a sequence comes from a movie, just answer if you recognize THE sequence):
+
+They only are used as controls is just a supplementary control (qui joue un rôle symétrique au taux de FA pour la prise de risque).
 
 %%Indoor/Outdoor
-In Isola et al., 2014, the authors showed that memorability directly correlated with color.
-To explain thi result, they advanced the hypothesis that the difference in memorability between indoor and outdoor scenes was due to the fact that the later tend to be less memorable that the older, and that the color tend to linked with a differences in colors, could explain this result.
+ %In Isola et al., 2014, the authors showed that memorability directly correlated with color.
+To explain this result, they advanced the hypothesis that the difference in memorability between indoor and outdoor scenes was due to the fact that the later tend to be less memorable that the older, and that the color tend to linked with a differences in colors, could explain this result.
 To validate this hypothesis, we manually annotated our 700 sequences as consisting of an indoor or outdoor scene.
+
+%fig with video frame memorability scores ranked from most t oless memorable.
+
+% potentially
+ %First, not to be part of the official trailers, to be sure that, if a participant report not having seen the movie, he won't have seen the sequence presented for the recognition in a trailer.
+=> Bof: this doesn't matter if False Alarms becauses of seen trailer.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{STUDY OF THE MEMORABILITY SCORES}
+\section{STUDY OF THE MEMORABILITY ANNOTATIONS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{}
-
-%We finally work on 660 movies (sequences annotated at least 4 times).
-
-%%%
-\subsubsection{Sequences' memorabilities}
-A sequence's memorability is defined as the percentage of correct detections by participants.
-On average, each video was viewed as target by $9.3$ participants. Average sequence memorability was $48.25\%$ (SD of $28.21\%$).
-On average, each video was viewed as filler by $9.2$ participants. Average false alarm rate was $5.95\%$ (SD of $14.47\%$). 
-%corriger score par moyenne de mem du film -- est-ce que ça a un sens ?
+For the results we retain only the 660 sequences that had been seen at least 4 times as targets, and remove sequence selected from films too few participants had seen.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Number of annotations per sequence}
-The histogram presented in the figure \ref{fig:nb_annotations_per_sequence} corresponds to the number of sequences for each possible number of annotations.
+\subsection{Overview}
+The average memorability of the test was $46.71\%$ (SD of $14.65\%$). The average false alarm rate was $4.16\%$ (SD of $5.27\%$).
 
+On average, each sequence was viewed as target by $10.7$ participants. 
+On average, each sequence was viewed as filler by $10.5$ participants.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sequences' memorabilities}
+%A sequence's memorability is defined as the percentage of correct detections by participants.
+
+%Corriger score par moyenne de mem du film -- est-ce que ça a un sens ? Par le nombre de vue du film ? par le moment où il a été vu?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Consistency analysis}
+The histogram presented in the figure \ref{fig:human_consistency}(a) corresponds to the number of sequences for each possible number of annotations.
+
+%Fig
 %Histogram of number of nb of sequences per possibble nb of annotations
+%Curve of human consistency
 \begin{figure}[!htbp]
 	\centering
-	\includegraphics[width=\columnwidth]{figures/histogram_nb_sequences_for_nb_of_annotations.png}
-	\caption{\label{fig:nb_annotations_per_sequence}Nb of sequences for each possible nb of annotations per sequence (>=3).}
+	\subfloat[Number of sequences associated with number of annotations.]{\includegraphics[width=0.3\columnwidth]{figures/histogram_nb_sequences_for_nb_of_annotations.png}}
+	\quad
+	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/Measure_of_human_consistency.png}}
+	\caption{\label{fig:human_consistency}(a) Number of sequences for each possible nb of annotations per sequence.
+(b) Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error).}
 \end{figure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Consistency analysis}
 We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency.
 It answers the question: "Are the videos that are more memorable (or forgettable) for a group of observers also more likely to be remembered (or forgotten) by a different group of observers?
-The figure \ref{fig:human_consistency} presents the curve of the human consistency for each video sequence.
+The figure \ref{fig:human_consistency}(b) presents the curve of the human consistency for each video sequence.
 It answers the question: "How many annotations for each do we need for the human consistency stop varying?'
-
-%Curve of human consistency
-\begin{figure}[!htbp]
-	\centering
-	\subfloat[(a) Human concistency for images in \cite{khosla_2015_understanding}]{\includegraphics[width=0.3\columnwidth]{figures/human_consistency_LaMem.png}}
-	\quad
-	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/Measure_of_human_consistency.png}}
-	\caption{\label{fig:human_consistency}Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error).}
-\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Generic vs. Typical sequences}
+\subsection{Neutral and typical sequences}
 % Tester s'il y a une différence entre les deux et en tirer des conséquence quant à leur validité --> How objective our measure of memorability is?

### 2018-01-15 ARTICLE.tex
Draft 1 -- ideas + figures put in orderly manner

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%citer Qomex
-%citer Deep learning emotional images bias
-
-%%%%%%%%%%%%%%%%%%%%%%
 %Opening
@@
 At least two important problems could explained this scarcity of studies.
-Firstly, protocols used in the existing studies are not smart enough, but are questionable.
-Secondly, authors didn't make their data accessible for community, so no database exist.
+
+%1st point: no available dataset
+Firstly, there is no available dataset to train and test models.
 The fact that there exist no dataset to modeling video memorability is the most serious obstacle which prevent video memorability prediction to take off.
@@
 The second released database, far larger, launched new possibility for deep learning and enable to obtained the aforementioned really good results \cite{khosla_2015_understanding}.
-Nous pouvons parier qu'il en sera de même pour le champ de la mémorabilité des vidéos.
+We could expect a similar success for video memorability prediction research.
+%A dataset as first objective to launch VM prediction --> Goal of this paper is to participant in the establishment of an adapted protocol
+We think that the existence of such a database with high quality data could have the importance for video memorability that the ones of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} had on image memorability.%in particular for deep learning algortihm which perform the best for images
 
-%A dataset as first objective to launch VM prediction --> Goal of this paper is to participant in th eestablishment of an adapted protocol
-We think that the existence of such a database could have the importance for video memorability that the ones of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} had on image memorability.%in particular for deep learning algortihm which perform the best for images
-This article wants to takes his part to develop such a protocol to collect high quality data, and such a dataset.
+%second point: difficulty to define video unity and VM
+This drives us to the second point: the difficulty to build such a dataset.
+The principle difficulty to define what video memorability and unity should be, and a protocol to measure items according to these definition.
 
-%it's not easy to construct the first dataset
-The establishment of a dataset, in particular of the first dataset of this kind, may not be self-evident.
-Elle nécessite de définir les points cardinaux qui vont guider – et contraindre – les chercheurs qui utiliseront ces premières données disponibles.
+If we consider the studies on image memorability prediction, we could see that, with their database, the cardinal points defined in the very first study \cite{isola_2011_makes} were widely accepted by the community without re-interrogation of what memorability -- of what the models -- really predicted.
+First, it is because Isola \textit{et al.} find a proper and objective way to measure memory performance at a large scale; but it remains that how the authors proceeded is just one way to measure memorability.
 
-Pour preuve, l'influence considérable qu'Isola et al.
-(puis Khosla et al., de la même équipe, avec leur base de données beaucoup plus conséquente rendue disponible en 2015) ont eu sur les études réalisées dans leur sillage : jusqu'à aujourd'hui, la quasi-totalité des études ayant porté sur la mémorabilité des images ont adopté – avec les bases de données de ces auteurs – leurs définitions de la mémorabilité.
+From what we can imagine that such a dataset for VM will probably have an important influence on future research, one should define very carefully what memorability of video is and what the video as unity means.
+It requires to define the cardinal points that will guide -- and constraint -- researchers that will used the dataset.
 
-Or, la manière dont les auteurs ont mesuré la mémorabilité des images, qui n'est pratiquement jamais rediscutée, n'est qu'une des manières de procéder.
-
-Aussi, si les modèles de prédiction de la mémorabilité des images ont obtenu d'excellents résultats (très proches de ceux qu'on obtiendrait en inférant des résultats d'un groupe d'observateur les résultats d'un autre groupe d'observateurs), qu'est-ce que ces modèles prédisent si bien, en fait ?
-
-%Difficulty: define video unity; complexity of the videos compare to images
+First, if image in image memorability was obvious, video in video memorability is less obvious.
 Videos are more complex than images.
@@
 %from ahn et al. Videos contain a synthesis of visual, aural, textual, and 	motion information and generally convey relatively richer 	media content than images.
+By an extreme example, is it possible for a model to transfer his capacity to predict memorability from movies-like videos to Youtube-like videos?
 
@@
 Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation \cite{mcgaugh_2000_memory}.
-
-In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance is particularly strong in the few days immediately following the memorization.
-
-, with a memory performance in recall that asymptotically tends its long-term performance \cite{ebbinghaus_1913_memory}.
-
+In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall is particularly strong in the few days immediately following the memorization, after which it asymptotically tends its the almost permanent long-term performance \cite{ebbinghaus_1913_memory}.
 Because several factors are susceptible to influence the consolidation process (such as the emotional content, the semantics, the attention...), which is not a linear decrease for all the memories, it follows that the order of memorability of images is susceptible to change from some minutes to a day after memorization, as it has been found in \cite{cohendet_2016_prediction} with emotional image from of the International Affective Picture System \cite{lang_2005_international}.
-
 A protocol to collect memorability annotations for videos would benefit from the capacity to measure what we will later refer to as "long-term memorability".
@@
 To our knowledge, the first attempt at predicting video memorability had been proposed by \cite{han_2015_learning}. 
-
 The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
@@
 They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
-
 We can infer from the sparse information authors gave that the task duration, for each of the 20 participants, was about 24 hours (by taking an average video duration half-way between 15 and 30 sec), spread over 10 sessions (5 free viewing and 5 recall tasks) of about 2 hours each.
-
-Authors used the same proportion of fillers in the free viewing and recall tasks (4/5 of fillers and 1/5 of targets), to "to guarantee that viewers are unaware of targets".
-If it was totally justified in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers (never reasked videos) in free viewing task would have very week impact on memorability scores (généralement on se passe même de fillers dans les tâches d'apprentissage du matériel).
-
-Still, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
-
-Regarding the prediction part, we also agree with these authors to say that "the method used fMRI measurements for predicting memorability, which would be difficult to generalize".
+Authors used the same proportion of fillers (i.e.
+non repeated videos) in the free viewing and recall tasks (i.e.
+4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
+If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in free viewing task would have very little impact on memorability scores (often authors even use only material interrogated later in learning/free viewing task).
+Furthermore, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
+Regarding the prediction part, we also agree with these authors to say that the method would be difficult to generalize as it used fMRI measurements for predicting memorability.
 
@@
 Another earlier approach was the one of \cite{shekhar_2017_show}.
-The participants passed in crowdsourcing a free viewing task followed by a recall task, for which they had to answer question (such as: "Did you see a man juggling?", "Did you see a car on road?").
-
-%the problem of the questions
-The major drawback of the study comes from this use of questions instead of a recognition task.
-The "memorability" measured may be due not to memorability of the videos but (also) to the differences between the questions rather than a difference of memorability of the videos.
-This bias could be of several kinds: difficulty, comprehension (non anglophone people in AMT), easility of imaging), indexage plus facile...
-C'est d'autant plu sproblématique que ces diffénrece could have drives to a high inter-human consistency not due to memorability attribué à tort à un phénomène de mémoire.
-C'est d'autant plus problématique que le temps de réponse des partcipants entre dans le calcul de la mémorabilité, ce qui ne peut q'uaccentuer ces biais, et est porbablement responsable de la high human response consistency they observe.
-
-Le fait que les questions doivent être créé à la main, et que les choix des questions sont assez limitatifs (par ex avec la question "Did you see a car on road?", il ne peut y avoir de toute la tâche qu'une seule fois une voiture) makes difficult the generalization of this protocol to build an extensive dataset.
-“It was manually ensured that no two textual questions nor any two videos in a sequence were similar in content.”
-
-%Conslusion critics
-Les critiques exposés montre à quel point il est difficile de construire la première base de données pour mesurer la mémorabilité des vidéos.
-Isola et al.
-S’en sont très bien sortis, en se cantonnant à une mesure purement objective de la mémoire – tâche de reconnaissance classique –, ce qui leur a permis de lancer le champ de recherche de la mémorabilité des images.
-Il faudrait en faire de même pour la mémorabilité des vidéos.
-Cette étude s’inscrit dans cette volonté – non pas de créer un nouveau datasets, mais d’essayer un protocole pour voir s’il est efficient.
+The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequences a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on road?").
+The major drawback of the study comes from the use of questions instead of a recognition task.
+Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty, ease of understanding (in particular for the non-anglophone people that works with Amazon Mechanical Turk), of imaging, ease to retrieve a scene by some words more the by others...
+Especially since the authors took into account the response time of the participants to calculate the memorability scores of the videos.
+One will note that this potential bias could have also affect the measure of inter-human consistency.
+Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limiting (e.g.
+the question "Did you see a car on road?" implies that in a whole experimental session you can have just one car on a road), and the authors manually ensured that no textual questions nor videos in a sequence were similar in content.
+This makes difficult the generalization of this protocol to build an extensive dataset.
+%%% ATT!! The have tested complexity of textual questions, but there is the other problems! mesure insufisante pour rendre compte de la difficulté réelle des questions.
 
@@
 \subsection{Goals of our study}
+%Conclusion critics
+It appears from the review of previous work that video memorability prediction research is in its early stage.
+We still need to find a proper way to collect quality data in order to build a large dataset to bring video memorability prediction research to a new level.
 %1/ we want a protocol
-To sum up, we want a protocol:
-- which could be extended to create a large dataset in case it shows its quality
-
-- that measure a long-term memorability: as a longitudinal study would be difficult to generalized to create a large dataset, we decide to measure the memoray of participants created previously to the experiment.
-
+The present study aims to be part of these efforts to reach such a way.
+As aforementioned, we also want this protocol to measure a "long-term" memorability.
+This can be achieved with a study that have a long time span, where learning and recall tasks are seperated by a certain amount of time, or by measuring a memory created prior to the experiment.
+Because crowdsourcing is a priori the only practicable way to experimentally collect great amount of annotations, we focused on the last choice more suitable for crowdsourcing as it requires only one connexion to the measurement application.
 %2/ we want an all-around paper
-D'autre part, all around paper
-
+To assess the quality of the data produced by the protocol, we analyzed the data obtained from different perspectives, in particular in order to shed light on some potential human or task-related factors susceptible to influence memorability as it is measured by the protocol, that are numerous \ref{cohendet_2016_using}.
 %3/ we want to meet the challenge of video memorability prediction
-Finally, we want -- thanks to our protocol -- to participate in the VM prediction challenge.
-
+Thanks to the data collected, we participate in the VM prediction challenge.
 %Conclusion intro
-The rest of the article is as follows: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and part 5 the conclusion and intended future work.
+The structure of the rest of the article follows these points: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and a conclusive part 5 where we talked about intended future work related to video memorability prediction.
 
@@
 Because the semantics is linked to the memorability of images \cite{isola_2014_makes}, so probably to videos, therefore their memorabilities, we chose this definition susceptible to be extensive to predict other dataset.
+
+%PARCE QUE CEST UNE PETITE BASE DE DONNEES => POUR UNE PLUS GROSSE, CE PROBLEME N4AURA PAS ROP D'EFFET.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we describe in detail the ground truth data collected thanks to the protocol described in the previous section.
-We conclude by calculating the memorability scores finally used to feed our prediction models.
+In this section, we analyze the ground truth data collected thanks to the protocol described in the previous section.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Memorability scores calculation}
+After collecting the data, we measured the average percentage of correct detections for all participants, of $46.71\%$ (SD of $14.65\%$), and the average false alarm rate, of $4.16\%$ (SD of $5.27\%$).
+Then we assigned a memorability score to each sequence, defined as the percentage of correct detections by participants.
+We retained only the 660 sequences that had been seen at least 4 times as targets, and remove the sequences chosen from films seen by too few participants.
+On average, each sequence was viewed as target by $10.7$ participants, and as filler by $10.5$ participants.
 
-For the results we retain only the 660 sequences that had been seen at least 4 times as targets, and remove sequence selected from films too few participants had seen.
+The figure \ref{fig:overview} provides a sample of sequences sorted according to their memorability (a) and a distribution of the sequences according to their degree of memorability (b).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Overview}
-The average memorability of the test was $46.71\%$ (SD of $14.65\%$). The average false alarm rate was $4.16\%$ (SD of $5.27\%$).
-
-On average, each sequence was viewed as target by $10.7$ participants. 
-On average, each sequence was viewed as filler by $10.5$ participants.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sequences' memorabilities}
-%A sequence's memorability is defined as the percentage of correct detections by participants.
-
-%Corriger score par moyenne de mem du film -- est-ce que ça a un sens ? Par le nombre de vue du film ? par le moment où il a été vu?
+%Examples of ++/-- memorable sequences and Distribution of memorability scores 
+\begin{figure}[!htbp]
+	\centering
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/key_frame_mem_examples.png}}
+	\quad
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/memorability_scores_distribution.png}}
+	\quad
+	\caption{\label{fig:overview}(a) Sample of the dataset with key frames of sequences sorted from most memorable (left) to less memorable (right). (b) Distribution of the memorability scores.}
+\end{figure}
 
@@
 \subsection{Consistency analysis}
-The histogram presented in the figure \ref{fig:human_consistency}(a) corresponds to the number of sequences for each possible number of annotations.
+%global consistency calculation
+We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency.
+We randomly split our 104 participants into two independent halves, and calculate how well sequence memorability scores from the first half of the participants match with sequence memorability scores from the second half of the participants.
+Averaging over 25 random split half trials, we calculated a Spearman's rank correlation $\rho$ of $0.57$ between these two sets of scores.
+%Individual and contextual differences, besides mandatory random variability, explain the $1-.57$ part of the memorability that is not universally derivable from the intrinsic informations of the videos.
 
-%Fig
+%Fig Consistency
 %Histogram of number of nb of sequences per possibble nb of annotations
@@
 	\centering
-	\subfloat[Number of sequences associated with number of annotations.]{\includegraphics[width=0.3\columnwidth]{figures/histogram_nb_sequences_for_nb_of_annotations.png}}
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/histogram_nb_sequences_for_nb_of_annotations.png}}
 	\quad
-	\subfloat[(b) Human consistency for videos for us]{\includegraphics[width=0.3\columnwidth]{figures/Measure_of_human_consistency.png}}
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Measure_of_human_consistency.png}}
+	\quad
 	\caption{\label{fig:human_consistency}(a) Number of sequences for each possible nb of annotations per sequence.
-(b) Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error).}
+(b) Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error and linear trendline).}
 \end{figure}
 
-%figure Rank corr for Isola et al. et han et al., pour montrer que le nombre d'annotation sva plus vite pour les videos (mais crowdsourcing vs. laboratoire -- et autres facteurs...)
-
-We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency.
-It answers the question: "Are the videos that are more memorable (or forgettable) for a group of observers also more likely to be remembered (or forgotten) by a different group of observers?
-The figure \ref{fig:human_consistency}(b) presents the curve of the human consistency for each video sequence.
-It answers the question: "How many annotations for each do we need for the human consistency stop varying?'
-
-%compounding factor: most annotated sequences are also the most memorable group, which could artificially increase the Spearman's correlation).
-
-Individual and contextual differences, besides mandatory random variability, explain the $1-.65$ part of the memorability that is not universally derivable from the intrinsic informations of the videos.
-
-%Compare our and Khosla et al. curve dire que -- video + non-crowdsourcing + "real" long-term memory -- explique qu'on obtient avec beaucoup moins d'annotation la même human congruency.
-If we compare the curve obtained by \cite{khosla_2015_understanding} (\ref{fig:human_consistency}(a))and the curve we obtained (b), we can note than we attain human consistency far more quickly than Khosla \textit{et al.}.
-At least three arguments go in the way of this results.
-First, we work with videos and they work with images ; maybe video memorability is more universal than image memorability because of the higher length of this kind of stimulus (but We can also note that the maximum human consistency is close for images and videos!!).
-Second, we obtain a measure of real long term memorability, and not a memorability measured some minutes after encoding step ; maybe this measure more representative of what is really memorable.
-Finally, we can advance the fact that in-lab experiment enable to obtain better memorability measure than crowdsourcing one, resulting to attain maximum human consistency earlier.
+%per nb of annotations consistency %comparison with image memorability
+We reproduced this calculation to obtain 25 Spearman's correlation coefficients as a function of the mean number of scores per video, presented in figure \ref{fig:human_consistency}(b).
+This curve is to be compared with the histogram presented in figure \ref{fig:human_consistency}(a), which shows that the number of sequences for each number of annotations was unequal.
+According to the curve, we achieved a mean consistency of $.70$ around 18 annotations, which is consistent with the previous attempt of \cite{han_2015_learning}, were it is the maximum consistency they achieve.
+It is interesting to compare these results to the one obtained by authors that have collected image memorability annotations \cite{isola_2011_makes,khosla_2015_understanding}.
+In these studies, the authors reached such a correlation of $.70$ after 80 annotations per image, which is much later.
+The experimental protocols and conditions are different between all these studies; in particular, emphasis must be placed on two points.
+First, in the present studies and \cite{han_2015_learning}, we measured a long-term memory performance later than two days after the memorization, that is to say after the biggest losses in memory performance, as was previously stated.
+By contrast, \cite{isola_2011_makes,khosla_2015_understanding} measured memory performance from dozens of seconds to a few minutes after memorization.
+Second, videos memorability annotations were collected through in-lab experiments, and images annotations through crowdsourcing experiments.
+These two confound factors may have contributed to the shortest number of annotations necessary to reach a high human consistency.
+However, it would be interesting in future work to confirm if an important difference exists between images and videos regarding to the number of annotations necessary to achieve a high human response consistency.
+Apart from the conclusions we could draw about the universality of the intrinsic memorability of videos compare to images, this would mean that the magnitude of the work to carry out to build an extensive database for video memorability prediction is substantially smaller than one could expect from work on image memorability prediction.
 
@@
 % Tester s'il y a une différence entre les deux et en tirer des conséquence quant à leur validité --> How objective our measure of memorability is?
+We select two sorts of sequences from the films: neutral and typical ones.
+For neutral sequences, participant had no element to guess that the sequence came from a particular film; for typical sequence, participants could guess that the sequence came from a particular film.
+The instructions was made clear, that we wanted that participants answered when they recognized a sequence, not when they guess that they came from a particular film they answered during the previous questionnaire film they had seen.
+Thus, the response for the neutral sequences is objective, but there is subjectivity in the typical sequence.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Quality of the movies}
-dvdrip vs. HD 720-1080 => Difference of memorability? => Interesting
+A non-parametric Wilcoxon rank sum test showed a significant difference in memorability between the neutral ($\mu=.24$) and typical ($\mu=.53$) sequences: $Z=10.22, p<.00001$.
+We expected a result in this sens because neutral sequence contained less contextual elements useful to recognize that a segment comes from a particular film.
+Because of this compounding factor this results does not mean that participants tended to guess -- rather to purely recognize -- sequences drawn from films they have seen.
+
+We also observed a difference in human consistency for memorability responses between the neutral ($\mu=.45$) and typical sequences ($\mu=.41$): $Z=2.75, p<.01$.
+The human congruency was slightly higher for neutral than for typical sequences.
+Along with the comments collected from the participants, who have as a majority reported difficulty to know if they guess or just recognized, this suggests that human congruence is higher for the objective measure than for a one with a part of subjectivity.
+In another context than with colleagues of our company, who highly probably respected the instructions -- in particular in crowdsourcing, which as we aforementioned stated that it must be used to built an extensive database -- it is probable that this difference would increase.
+Certainly, this result has to be confirmed by further studies, but it could be one of the weakness of our protocol to collect extensive data.
+
+We can also note that the false alarm rate was low for neutral sequences ($\mu=.05$) as well as for typical sequences ($\mu=.03$).
+Specifically, we expect lucky confusions to account for little of correct detections on average for the two sorts of sequences.
 
@@
 \subsection{Response time}
-Previous authors have integrated response time in their score of memorability.
-We pose here the following question: "Is the degree of memorability related to response time?" to answer the question: "Should we exploit response time to correct memorability score?" We hypothesized that the most memorable the sequences, the faster the participant will answer (no instructions to answer faster: just to answer durgin the 10-sec display).
-
 %figure Mean response time vs. degrees of memorability
@@
 	\includegraphics[width=\columnwidth]{figures/Response_time_vs_Memorability.png}
-	\caption{\label{fig:response_time_vs_memorability}Mean response time for each memorability degree (error bars correspond to SEM).}
+	\caption{\label{fig:response_time_vs_memorability}Mean response time for each memorability degree for targets for which participants answered(error bars correspond to SEM).}
 \end{figure}
 
-We observed a Person's correlation of $-0.36 (p<.0001)$ between the response time on the target and their memorability scores.
-This is consistent with the figure \ref{fig:response_time_vs_memorability}.
-In other words, the higher the memorability, the lower the response time tends to be, suggesting that a sequence most memorable is also a sequence for which we rapidly detect that it is memorable.
-This is consistent with what we expected: empirically, we observe during passation that participant tend to answer as soon as they recognized the sequence.
+The figure \ref{fig:response_time_vs_memorability} shows that the response time to do a correct detection decreases when the memorability of the sequence increase.
+We also observed a  Person's correlation of $-0.36 (p<.0001)$ between the response time on the target and their memorability scores.
+These result show that the participants tended to answer quicker when the sequences were memorable than non memorable, even though the participants did not receive any instruction to answer quickly.
+This suggest two things.
+First, that participants tends to naturally answer rapidly after having recognized the sequence.
+Second, either that the most memorable sequences are also the sequences the most accessible in memory, or that the most memorable sequences have more early recognizable elements than less memorable sequences.
+However, the way we select sequences and the linear decrease in response time with memorability increasing showed in figure \ref{fig:response_time_vs_memorability} go in the direction of the first hypothesis.
 
 %target vs. fillers
-It is interesting to link it to the response time for answer on targets (i.e.
-correct detections) vs.
-on fillers (i.e.
-false alarms).
-The global mean response time was $4.87 sec$ on targets and $5.96 sec$ on fillers.
-A Student t-test for different sample size show a significant difference ($t(2836)=-5.34, p<.0001$).
-This means that the participants globally answered more rapidly for targets (i.e.
+It is interesting to link it to the response time for answer on targets (i.e. correct detections) vs. on fillers (i.e. false alarms).
+The global mean response time was $4.88 sec$ on targets and $5.90 sec$ on fillers.
+A Wilcoxon rank sum test showed a significant difference ($Z=-5.10, p<.00001$), in the way participants globally answered more rapidly for targets (i.e.
 correct detections) than for fillers (i.e.
 false alarm).
-One explanation is that participants hesitated more for fillers they answered on, increasing their response time.
+One explanation is that participants hesitated more for fillers they answered on than to make a correct detection, increasing their response time.
+This hesitation would reinforce the hypothesis that the most memorable sequences are most easily accessible in memory: if it has just been a question of a difference in the early recognizable elements that had explained that participants answered faster for the most memorable sequences, and not a question of hesitation -- i.e.
+of accessibility in memory --, the response time would have been the same for false alarm than for correct detection.%consistent with memory accessibility in psychology
+%L’étude de la morphologie des distributions nous enseigne que le temps de réponse tend à être uniforme pour les fillers, alors que les participants répondent plus vite pour les fillers.
 
-It tells us something on memory, that the memories are not evident but more blurred.
-This is an important point, taht explained why, in our discussion, we propose to turn to a "totally objective" measure of memory to constitute a dataset for video memorability prediction: if the participants hesitated, so they choice to answer or not was probably impacted by the way they answered the instructions and their own sensibility to the risk.%=> corrigé mémorabilité par individu ?
+It also point out the fact that memories of videos are generally not obvious but more blurred.
+This is an interesting point to connect with the fact that human consistency is lower for typical than for neutral sequences: because of the non obviousness of if they had memories or not, the part of subjectivity for typical sequences could have fostered participants to answer when they hesitated if they were guessing that a sequence came from a film they had seen or remembering it.
+The higher degree of memorability of typical sequences compare to neutral ones could be partly due to this factor too.
+Among other, this is also an argument in favour of the most "objective" possible measure to choose to constitute an extensive dataset for video memorability prediction.
 
-%How to exploit the response time in the prediction models? --> two potential models for memorabilit prediction
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evolution of the memorability along time}
-%Use order positions for that (e.g. mean memorability for position 1, 2... 120)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%tester la mémorability de séquences très proches (prise à intervalles très proche + vérif = même scène)
-
-%indoor/outdorr comme isola et al.
-mais en contrecarrant l'effet de la couleur -- car ils espliquent que c'est par la corrrélation entre couleur et mémorabilité est expliqué par le fait que les couleurs plus froides sont d'exté.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New manner to compute memorability scores: take into account time and FA}
-%Une nouvelle manière pour prendre en compte les FA dans les scores de mémorabilité applicable à l'étude de la mémorabilité sur les images (si fonction : en parler dan sl'intro comme quoi les gens ne prennent jamais en compte les FA quoiqu'elles puissent être exploiter)
-%Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
-
-%%%
-\subsubsection{Participants' performance}
-% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the film --> No' --> So the point here is just to provide an overview.
-The average memory performance was the following: the average percentage of correct detection was $48.2\%$ (SD of $14.1\%$) and the average percentage of false alarms was $4.78\%$ (SD of $5.63\%$).
+To conclude this section, the degree of memorability sounds to be naturally related to response time.
+Therefore, we could imagine to exploit in the calculation of memorability score.
+That is what \cite{shekhar_2017_show} did in their previous attempt to construct their dataset to predict video memorability: they integrated response time in the calculation of the memorability score of their videos as the principal element (even if, in this case, the aforementioned problem of the use of questions could have influenced their results).
 
@@
 \subsection{Logistic regression vs. SVM to personalize prediction model}
-%Regression qui prend en compte tous les facteurs (voir ma thèse)
 %See QoMEX paper to present the data
+
+%figure Logistic regression
+\begin{figure}[!htbp]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/Logistic_regression.png}
+	\caption{\label{fig:logistic_regression}.}
+\end{figure}
 
@@
 
+%Corr Evolution of the memorability along time
+ %Use order positions for that (e.g. mean memorability for position 1, 2... 120)
+
+%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Film genre and IMDB ratings}
+%\subsection{Mean memorability of the sequences compare to the film they come from}
+	%Corriger score par moyenne de mem du film -- est-ce que ça a un sens ? Par le nombre de vue du film ? par le moment où il a été vu?
+
+%\subsection{Quality of the movies}
+	%dvdrip vs. HD 720-1080 => Difference of memorability? => Interesting
+
+%\subsection{Film genre and IMDB ratings}
 	%test the momorability od the sequences compare to the mean one of the movie.
@@
 
-\subsection{Context}
+%\subsection{Context}
 	%Analyse the context (order of video presentation) influent the memo: to be check if we have enough information?
@@
 	%S'inspirer de la technique de Bylinskii
+	%tester la mémorability de séquences très proches (prise à intervalles très proche + vérif = même scène)
 
-\subsection{Features linked to memorability}
+%\subsection{Features linked to memorability}
 	%tester ces features comme dans 'What makes a photograph...'
@@
 
-\subsection{Indoor \textit{.vs} outdoor scenes}
+%\subsection{Indoor \textit{.vs} outdoor scenes}
+	%indoor/outdoor comme isola et al.
+mais en contrecarrant l'effet de la couleur -- car ils expliquent que c'est par la corrrélation entre couleur et mémorabilité est expliqué par le fait que les couleurs plus froides sont d'exté.
 
+%\subsection{Other points}
 	%add other relevant points found in MIT papers e.g.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Memorability score calculation}
-According to what was presented before in tis section, this is how we finally decided to compute our memorability scores...
-
-%quality of the movies: dvdrip vs. HD 720-1080 => Difference of memorability? => Interesting
+%\subsection{New manner to compute memorability scores: take into account time and FA}
+%Une nouvelle manière pour prendre en compte les FA dans les scores de mémorabilité applicable à l'étude de la mémorabilité sur les images
+%Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
+% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the film --> No' --> So the point here is just to provide an overview.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%avantages de notre dataset: long-term memorabity
+%Inconvénient: mesure avec une part de subjectivité
+
+%we propose to turn to a "totally objective" measure of memory to constitute a dataset for video memorability prediction
+
 % Nous construisons actuellement une base de données de grande ampleur avec une mesure purement objective de la mémoire.

### 2018-01-15 ARTICLE.tex
Added Modelling part of the paper

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Until now, we have seen how we collected data and memorability annotations from users.
+We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
+In this section, we move towards building a machine learning model that can learn and then predict the memorability scores of a video from its audio-visual features. 
+We pose the problem of predicting memorability scores as a standard regression problem and Figure \ref{prop-apprch} illustrates the different steps in our method.
+From the figure, we can see that the video sequences are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new video sequences.
+In this following sub-sections, we explain our choice of features, models to address the problem in hand.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.9\columnwidth]{figures/approach.pdf}
+		\caption{Proposed approach for automatically predicting memorability scores.}
+    \label{prop-apprch}
+\end{figure}
+
+Before we go into the technical details about the different audio-visual features and regression models we explored, we will talk about how we split our dataset.
+We split our dataset, at the level of movies, into training (70\%), validation (15\%) and test (15\%) data, which translates into 70 movies in the training set and 15 movies each in the validation and test sets.
+We chose to split our dataset at the level of movies, instead of the video sequences, in order to avoid sequences from the same movie being present in the train as well as the evaluation set (validation+test).
+
+\subsection{Feature extraction}
+\label{feat-extr}
+We explore a variety of audio-visual features in building a model that can predict memorability scores for new video sequences.
+We mostly stick to state-of-the-art deep learning features that have been proposed in vision and audio research communities for tasks like video classification \cite{c3d-feat}, event detection \cite{audioset-feat}, image captioning \cite{caption-feat}, and concept detection \cite{sb-feat}.
+
+We use the features extracted from the C3D model, a 3-dimensional convolutional network proposed for generic video analysis \cite{c3d-feat}. 
+The main motivation to use these features is that these features try to encode the motion information in the video.
+The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
+We use the model trained on the Sports-1M dataset \cite{c3d-feat} that is provided by the authors of the paper.
+We explore the output of the 3 fully connected layers of the network: fc6, fc7, and fc8 with a dimensionality of 4096.
+Since the dimensionality of the features is very high, we also explore a dimension reduction method (Principal Component Analysis) for modelling.
+
+Google Inc. recently released a large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat}. 
+The dataset consists of 10-second audio clips taken from YouTube videos along with code to extract 128-dimensional embeddings for your own audio clips.
+We use these features as they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.
+We extract the 128-dimensional embeddings for our dataset and use these embeddings in training the regression models.
+
+Now, we move onto features related to emotion.
+Emotion and memory are correlated \cite{}, which is why we look at emotion related features in building a model for memorability prediction.
+For emotion from visual content, we resort to a visual sentiment concept detector: Sentibank \cite{sb-feat}.
+SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
+We use the binary code for concept detection, from images, provided by the authors.
+We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
+For each of these 10 frames, we run the sentibank concept detector and obtain two pieces of information: concepts with probabilities and features.
+Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
+We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
+For each of the 50 concepts, we extract 300-dimensional word2vec \cite{word2vec} embeddings by training on english wikipedia corpus and taking an average of all the vectors to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
+Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
+In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors as features to build a model for memorability prediction.
+Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video sequence to obtain arousal and valence scores \cite{affect}.
+Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video happy or sad, following a circumplex model of affect \cite{affect-model}.
+For each frame in the video sequence, we compute the arousal and valence scores using the methods proposed in \cite{affect}.
+We concatenate the arousal and valence scores for the first 200 frames in each video sequence resulting in a 400-dimensional feature vector (200 for arousal and 200 for valence) for a video.
+
+In addition to emotion, semantics also play a role in memory.
+We utilize the ongoing research in image captioning to capture the semantics \cite{caption-feat}.
+We use the code provided by the authors to extract captions for videos in our dataset.
+We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
+For each of these 10 frames, we run the caption detector and obtain a caption for the frame.
+For each word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings by training on english wikipedia corpus and taking an average of all the words to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
+We use these 300-dimensional vectors for building our model.
+
+\subsection{Modelling and evaluation}
+\label{model-eval}
+We discussed the variety of features that we extract from the audio-visual content.
+Now, we turn to picking a regression model that can effectively predict memorability scores for videos in our dataset.
+We choose to go with the Support vector regression model (SVR) \cite{svr}, with a linear and an RBF kernel, for its popularity and wide applicability.
+Since we are posing the memorability prediction problem as a regression, we use the standard regression metric: Mean squared error ($mse$) for evaluation.
+Additionally, we also use the spearmann correlation ($spcorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
+This metric gives us an indication of the correlation between the predicted memorability scores and the human annotated memorability scores.
+
+\section{Memorability prediction results}
+\label{pred-res}
+In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new video sequences.
+Each of the video sequences in our dataset has multiple annotations and we pick sequences with atleast 4 annotations for training the models.
+In our dataset, 660 sequences (need to confirm) out of the 700 sequences have atleast 4 annotations.
+One set of results correspond to using these 660 sequences. 
+Another set of results correspond to using 330 sequences (need to confirm) which have atleast 10 annotations.
+
+Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies).
+We pick the model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $mse$ and $spcorr$.
+We retrain the model with these parameters and evaluate on the final test set.
+In the experiments where we use a dimension reduction method (PCA): C3D and SentiBank features, we try to retain 95\% variance in the data while reducing the feature dimensions.
+Tables \ref{res-4-ann} and \ref{res-10-ann} report the memorability prediction results for different features using models trained on sequences with atleast 4 and 10 annotations.
+Observing the tables, you can clearly see that the image captions perform the best among all the features.
+These features capture the semantics in the video and are helping in a better prediction of the memorability scores.
+The next best performance (apart from the combination of all the features) is observed for the features from the AudioSet embeddings.
+Since these embeddings encode the event related information, we can say that the events happening in a video play a significant role in how the video is remembered.
+A combination of all the features is the second best performing feature combination in terms of the metrics.
+We concatenate all the features into a single feature vector and train a model which is then used to predict the memorability scores on the test data. 
+We observe that there is a decline in performance when compared to the best performing features: image captions and this is because the other features are hindering the performance of the regression model.
+
+One of our initial hypotheses was that emotion would play a big role in memorability of a video sequence, supported by literature from psychology.
+We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
+Observing the tables, you can say that the SentiBank features perform reasonably well, but it is not close to the performance of features encoding the semantics.
+This could be because of the choice of features that encode emotion or the performance of the emotion models not being good enough for such an application.
+From the analysis on this dataset and the features we used for encoding emotion, we cannot conclude that emotion plays an important role in memorability.
+
+Comparing the results across the two tables, we observe that the model trained on sequences with atleast 4 annotations performs better than the model trained on sequences with atleast 10 annotations.
+The only exception being the C3D feature, where the model trained on sequences with 10 annotations performs marginally better than the model trained on 4 annotations. 
+Now, we will investigate how the performance changes with increasing number of annotations: from 4 to 15.
+For this, we take the best performing feature: image captions and analyze its performance with models trained on sequences with different number of annotations.
+We take sequences with 4 annotations and keep increasing the number of annotations per sequence to 15.
+In each step, we repeat the training-validation-testing procedure of the previous experiment and compute the evaluation metrics.
+We observed that the $mse$ does not change significantly in each of the steps and hence we provide an demonstration of how $spcorr$ varies with increasing number of annotations.
+Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $spcorr$ keeps going down.
+
+\begin{table*}[!ht]
+\centering
+    \begin{tabular}{| c | c | c | c |}
+    \hline
+				Feature& Model& $mse$& $spcorr$\\ \hline
+				C3D (4096)& Linear, $C$=1.0& 0.08& 0.26\\ \hline
+				C3D (PCA) (225)& Linear, $C$=1.0& 0.08& 0.21\\ \hline
+				AudioSet (128)& RBF, $C$=1.0, $\gamma$=0.01& 0.06& 0.22\\ \hline
+				Affect (400)& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.17\\ \hline
+				SentiBanl Conc. (300)& RBF, $C$=1.0, $\gamma$=10& 0.08& 0.13\\ \hline
+				SentiBank Feat. (4096)& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.21\\ \hline
+				SentiBank Feat. (PCA) (225)& RBF, $C$=1.0, $\gamma$=0.01& 0.07& 0.21\\ \hline
+				\textbf{Captions (300)}& \textbf{RBF, $C$=1.0, $\gamma$=0.01}& \textbf{0.06}& \textbf{0.31}\\ \hline
+				All (1578)& RBF, $C$=1.0, $\gamma$=0.1& 0.06& 0.23\\ \hline
+    \end{tabular}		
+		\caption{Regression results for different features with models trained on sequences that have atleast 4 annotations.}
+		\label{res-4-ann}	
+\end{table*}
+
+\begin{table*}[!ht]
+\centering
+    \begin{tabular}{| c | c | c | c |}
+    \hline
+				Feature& Model& MSE& SpearCorr\\ \hline
+				C3D (4096)& Linear, $C$=1.0& 0.08& 0.24\\ \hline
+				C3D (PCA) (225)& Linear, $C$=1.0& 0.09& 0.23\\ \hline
+				AudioSet (128)& RBF, $C$=1.0, $\gamma$=0.1& 0.1& 0.15\\ \hline
+				Affect (400)& RBF, $C$=1.0, $\gamma$=0.01& 0.09& 0.17\\ \hline
+				SentiBanl Conc. (300)& Linear, $C$=1.0& 0.08& 0.14\\ \hline
+				SentiBank Feat. (4096)& RBF, $C$=1.0, $\gamma$=10& 0.09& 0.14\\ \hline
+				SentiBank Feat. (PCA) (225)& Linear, $C$=1.0& 0.08& 0.03\\ \hline
+				\textbf{Captions (300)}& \textbf{RBF, $C$=1.0, $\gamma$=0.01}& \textbf{0.06}& \textbf{0.22}\\ \hline
+				All (1578)& RBF, $C$=, $\gamma$=& 0.07& 0.21\\ \hline
+    \end{tabular}		
+		\caption{Regression results for different features with models trained on sequences that have atleast 10 annotations.}
+		\label{res-10-ann}	
+\end{table*}
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
+		\caption{Spearmann correlation on test data with models trained using features from image captions for varying number of annotations.}
+    \label{num-ann}
+\end{figure}
+
+\subsection{Cross-validation}
+\label{cross-val}
+Since we have a relatively small dataset, we perform a cross-validation experiment in order to ensure that the model is not overfitting.
+We perform a 10-fold cross-validation on the cross-validation set: combining the training and validation sets.
+We report the results (both $mse$ and $spcorr$) of the cross-validation for the following parameters: features from image captions, model trained on sequences with atleast 4 annotations, RBF kernel.
+Figure \ref{cross-val} reports the results of the cross-validation experiment and we observe that the evaluation metrics do not change much across the different folds.
+We can conclude that the trained model is generalizable and there is not much overfitting.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.8\columnwidth]{figures/cross-val.pdf}
+		\caption{Cross-validation performance on a model trained on sequences with atleast 4 annotations with features from image captions and an RBF kernel.}
+    \label{cross-val}
+\end{figure}
 

### 2018-01-18 ARTICLE.tex
Changes to the modelling part

--- 2018-01-15 ARTICLE.tex
@@
 \usepackage{graphicx}
+\usepackage{hhline}
+\usepackage{multirow}
 
@@
 	\centering
-	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/key_frame_mem_examples.png}}
+%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/key_frame_mem_examples.png}}
 	\quad
-	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/memorability_scores_distribution.png}}
+%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/memorability_scores_distribution.png}}
 	\quad
@@
 	\quad
-	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Measure_of_human_consistency.png}}
+%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Measure_of_human_consistency.png}}
 	\quad
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/Response_time_vs_Memorability.png}
+%	\includegraphics[width=\columnwidth]{figures/Response_time_vs_Memorability.png}
 	\caption{\label{fig:response_time_vs_memorability}Mean response time for each memorability degree for targets for which participants answered(error bars correspond to SEM).}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/Logistic_regression.png}
+%	\includegraphics[width=\columnwidth]{figures/Logistic_regression.png}
 	\caption{\label{fig:logistic_regression}.}
@@
 
+Until now, we have presented the video collection and the annotation protocol. 
+We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
+In this section, we move towards building a machine learning model that can learn and then predict the memorability score of a video from its audio-visual features. 
+We pose the problem as a standard regression problem and Figure \ref{prop-apprch} illustrates different steps in our method.
+First, we can see that the video sequences are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new video sequences.
+In this following sub-sections, we explain our choice of features, models to address the problem in hand.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.9\columnwidth]{figures/approach.pdf}
+		\caption{Proposed approach for memorability score prediction.}
+    \label{prop-apprch}
+\end{figure}
+
+Before we go into the technical details about the different audio-visual features and regression models we explored, we will talk about how we split our dataset.
+We split our dataset, at the level of movies, into training (70\%), validation (15\%) and test (15\%) data, which translates into 70 movies in the training set and 15 movies each in the validation and test sets.
+We chose to split our dataset at the level of movies, instead of the video sequences, in order to avoid sequences from the same movie being present in the training as well as the evaluation (validation+test) set.
+
+\subsection{Feature extraction}
+\label{feat-extr}
+We explore a variety of audio-visual features in building a model that can predict memorability scores for new video sequences.
+We mostly stick to state-of-the-art deep learning features that have been proposed in vision and audio research communities for tasks like video classification \cite{c3d-feat}, event detection \cite{audioset-feat}, image captioning \cite{caption-feat}, and concept detection \cite{sb-feat}.
+
+\subsubsection{Spatio-temporal visual features (C3D)}
+\label{c3d-feat}
+These features are extracted from the C3D model, a 3-dimensional convolutional network proposed for generic video analysis \cite{c3d-feat}. 
+The main motivation to use C3D is that it encodes both the spatial and temporal information in the video.
+The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
+We use the model trained on the Sports-1M dataset \cite{c3d-feat} that is provided by the authors of the paper.
+We use the output of the fully connected layer -- fc6 of the network with a dimensionality of 4096.
+Since such dimensionality is high compared to other features, we explore the use of principal component analysis (PCA) for the dimension reduction.
+
+\subsubsection{Audio features (AudioSet)}
+\label{as-feat}
+Using a recently released large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat} and the corresponding event detection models, we extract 128-dimensional embeddings for each audio track associated with a video in our dataset.
+We use these embeddings for training the regression models.
+The motivation to use these features is that they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.
+
+\subsubsection{Emotion related features (SentiBank and Affect)}
+\label{emo-feat}
+As research in psychology showed that emotion and memory are highly correlated \cite{emo-mem}, we investigate the use of emotion-related feature in our prediction system.
+For emotion from visual content, we resort to a visual sentiment concept detector: SentiBank \cite{sb-feat}.
+SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
+We use the binary code for concept detection, from images, provided by the authors.
+We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
+For each of these 10 frames, we run the SentiBank concept detector and obtain two pieces of information: concepts with probabilities and features.
+Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
+Examples of some concepts in the SentiBank ontology are: young driver, scary face, terrible pain etc.
+We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
+For each of the 50 concepts, we extract 300-dimensional word2vec \cite{word2vec} embeddings, using a model trained on english wikipedia corpus.
+We then take an average of all the vectors to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
+Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
+In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors as features to build a model for memorability prediction.
+
+Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video sequence to obtain arousal and valence scores \cite{affect}.
+Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video happy or sad, following a circumplex model of affect \cite{affect-model}.
+For each frame in the video sequence, we compute the arousal and valence scores using the method proposed in \cite{affect}.
+We concatenate the arousal and valence scores for the first 200 frames in each video sequence resulting in a 400-dimensional feature vector (200 for arousal and 200 for valence) for a video.
+
+\subsubsection{Visual semantic features (Image captions)}
+\label{sem-feat}
+Visual semantics are known to play an important role in memorability \cite{sem-mem}.
+We utilize the state-of-the-art research in image captioning to capture the semantics \cite{caption-feat}.
+We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
+For each of these 10 frames, we run the caption detector (code provided by the authors) and obtain a caption for the frame.
+For each word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings and taking an average of all the words to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
+We use these 300-dimensional vectors for building our model.
+
+\subsection{Modelling and evaluation}
+\label{model-eval}
+We used the features discussed in Section \ref{feat-extr} to train an SVR model for the memorability score prediction. 
+We studied the use of two kernels: linear and RBF for their popularity and wide applicability.
+We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ and $\gamma$.
+
+Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
+Additionally, we also use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
+This metric gives us an indication of the correlation between the predicted memorability scores and the ground-truth memorability scores.
+
+\subsection{Memorability prediction results}
+\label{pred-res}
+In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new video sequences.
+In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating all the feature vectors into a single feature vector.
+While concatenating the feature vectors, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set of features.
+Each of the video sequences in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
+In this way, we eliminate movies that were not seen by a majority of the users who participated in our experiment.
+Additionally, we investigate the effect of number of annotations, for video sequences in the training set, on the performance of the model.
+
+Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
+We pick the model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $MSE$.
+We retrain the model with these parameters and evaluate on the final test set.
+In the experiments where we use a dimension reduction method (PCA): C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
+
+Based on our experiments on the validation set, we observe that the image caption feature performs the best while using training video sequences that have atleast 4 annotations.
+Based on this result, we train different SVR models with varying number of annotations on the training set using image caption features.
+For example, we train a model, using image caption features, on training sequences with at least 4 annotations and use this model to predict the memorability score for video sequences in the validation set.
+We repeat this process for different number of annotations: from 4 to 15 annotations increasing the number of annotations by 1 in each iteration.
+Please note that the validation set in each of the repetitions is fixed and only the training set changes.
+We observed that the $MSE$ does not change significantly in each of the steps and hence we provide a demonstration of how $SpCorr$ varies with increasing number of annotations in the training set.
+Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $SpCorr$ keeps going down.
+In the wake of this observation, we investigate the performance of all the features when we train the regression model with video sequences that have at least 8 annotations.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
+		\caption{Spearmann correlation on validation set with models trained using image caption features for varying number of annotations.}
+    \label{num-ann}
+\end{figure}
+
+\begin{table*}
+  \centering
+  \renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|c|c|c|c|c|c|c|c|}
+    \hline
+    \textbf{Feature} & \textbf{Dimension}& \multicolumn{3}{c|}{\textbf{SVR Model (4 annotations)}} & \multicolumn{3}{c|}{\textbf{SVR Model (8 annotations)}}\\
+    % \hline
+    % \textbf{Inactive Modes} & \textbf{Description}\\
+    \cline{3-8}
+    & & Parameters& $MSE$ & $SpCorr$ & Parameters & $MSE$ & $SpCorr$\\ \hline
+    C3D& 4096& Linear, $C$=1.0& 0.08& 0.26& Linear, $C$=1.0& 0.07& 0.34\\ \hline
+		C3D (PCA)& 225& Linear, $C$=1.0& 0.08& 0.21& RBF, $C$=1.0, $\gamma$=0.01& 0.08& 0.17\\ \hline
+		AudioSet& 128& RBF, $C$=1.0, $\gamma$=0.01& 0.06& 0.22& RBF, $C$=1.0, $\gamma$=0.01& 0.05& 0.24\\ \hline
+		Affect& 400& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.17& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.23\\ \hline
+		SentiBank concepts& 300& RBF, $C$=1.0, $\gamma$=10& 0.08& 0.13& RBF, $C$=1.0, $\gamma$=0.1& 0.08& 0.17\\ \hline
+		SentiBank features& 4096& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.21& RBF, $C$=1.0, $\gamma$=0.01& 0.07& 0.26\\ \hline
+		SentiBank features (PCA)& 225& RBF, $C$=1.0, $\gamma$=0.01& 0.07& 0.21& RBF, $C$=1.0, $\gamma$=0.01& 0.07& 0.23\\ \hline
+		Captions& 300& RBF, $C$=1.0, $\gamma$=0.01& 0.06& 0.31& RBF, $C$=1.0, $\gamma$=0.01& 0.05& 0.38 \\ \hline
+		Combination& 1578& RBF, $C$=1.0, $\gamma$=0.1& 0.06& 0.23& RBF, $C$=1.0, $\gamma$=0.1& 0.05& 0.27\\ \hline		
+    \hline
+  \end{tabular}
+  \caption{Prediction results on the test data for different features with models trained on sequences that have at least 4 (columns 3-5) or 10 (columns 6-8) annotations.}
+	\label{res-4-10-ann}	
+\end{table*}
+
+Now, reverting to the main experiment to measure how each feature performs in predicting the memorability scores for video sequences in the test set. 
+We compute the average number of annotations per sequence in training and test set for both the cases in order to check if there is a mismatch.
+On an average, each sequence in the test set has around 10 annotations and the test set remains the same across different experiments.
+Similarly, each sequence in the training set has 9 annotations on an average in the case where we train the models with sequences having at least 4 annotations and 11 annotations in the other case.
+Looking at these numbers, we can see that the average number of annotations per sequence is quite balanced across the training and test sets.
+Table \ref{res-4-10-ann} reports the memorability prediction results for different features using models trained on sequences with at least 4 (columns 3-5) and 10 (columns 6-8) annotations.
+Observing the table, we can clearly see that the image captions perform the best, among all the features, in terms of $MSE$ and $SpCorr$.
+These features capture the semantics in the video and are helping in better predicting the memorability scores on new data.
+Though similar $MSE$ scores are obtained for other features like AudioSet, the memorability scores, predicted by the model trained on image caption features, are better correlated with the ground-truth ($SpCorr$ = 0.31 and 0.39).
+A combination of all the features is the second best performing feature combination in terms of $MSE$, but falls behind other individual features (AudioSet and C3D) in terms of $SpCorr$.
+
+One of our initial hypotheses was that emotion would play a important role in memorability of a video sequence, supported by literature from psychology.
+We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
+Observing the scores in Table \ref{res-4-10-ann}, we can say that the SentiBank features (fourth row from bottom) and affect features (sixth row from bottom) perform reasonably well, when the video sequences in the training set have at least 8 annotations.
+But the models trained on image caption features out-perform those trained on emotion features.
+This could be because of the choice of features that encode emotion or the performance of the emotion models not being good enough for such an application.
+From the analysis on this dataset and the features we used for encoding emotion, we \emph{cannot} conclude that emotion plays an important role in memorability.
+
+Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the model trained on sequences with at least 8 annotations performs better than the model trained on sequences with at least 4 annotations.
+The only exception being the C3D (PCA) feature, where the model trained on sequences with at least 4 annotations performs marginally better than the model trained on sequences with at least 8 annotations.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %Inconvénient: mesure avec une part de subjectivité
-
+We presented a machine learning model that can predict memorability score for a given video sequence.
+After exploring a variety of state-of-the-art deep learning audio-visual features, we can say that a model trained with features from image captions (semantic features) can provide predictions that are most correlated with ground-truth memorability scores.
+We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
+An immediate direction for future research would be to explore other approaches to encode emotion in videos and investigate whether we can improve the prediction performance using these approaches.
 %we propose to turn to a "totally objective" measure of memory to constitute a dataset for video memorability prediction

### 2018-01-26 ARTICLE.tex
Changes in modelling part according to suggestions from Ngoc and Claire-Helene

--- 2018-01-15 ARTICLE.tex
@@
 \section{MEMORABILITY PREDICTION} %for Khartik
+\label{mem-pred}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Until now, we have presented the video collection and the annotation protocol. 
+%We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
+In this section, we move towards building a machine learning model that can learn and then predict the memorability score of a video from its audio-visual features.
+The main goal of modelling is to identify which features: generic, perceptual, or semantic, are important for predicting the memorability scores of a video and \emph{not} the best learner/classifier.
+We pose the problem as a standard regression problem and Figure \ref{prop-apprch} illustrates different steps in our method.
+First, we see that the videos are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new videos.
+In the following sub-sections, we explain our choice of features and models to address the problem in hand.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.9\columnwidth]{figures/approach.pdf}
+		\caption{Proposed approach for memorability score prediction.}
+    \label{prop-apprch}
+\end{figure}
+
+%Before we go into the technical details about the different audio-visual features and regression models we explored, we will talk about how we split our dataset.
+To build our predictive model, we split the dataset, at the level of movies, into training (70\%), validation (15\%) and test (15\%) data, which translates into 70 movies in the training set and 15 movies each in the validation and test sets.
+We chose to split our dataset at the level of movies, instead of the videos, in order to avoid videos from the same movie being present in the training as well as the evaluation (validation+test) set.
+
+\subsection{Feature extraction}
+\label{feat-extr}
+%We explore a variety of audio-visual features in building a model that can predict memorability scores for new videos.
+The task of remembering a specific sequence from a movie has a high cognitive complexity in general, and is difficult for humans.
+Many users who participated in our experiments indicated that it is a difficult task.
+While trying to build a machine learning model for such a task, we wanted to explore different kinds of features that could be extracted from the audio-visual signal.
+We investigate a variety of generic state-of-the-art features (\cite{c3d-feat}, \cite{audioset-feat}) and compare them with other semantic (\cite{caption-feat}) and perceptual (emotion) features (\cite{sb-feat}).
+
+%We mostly stick to state-of-the-art deep learning features that have been proposed in vision and audio research communities for tasks like video classification \cite{c3d-feat}, event detection \cite{audioset-feat}, image captioning \cite{caption-feat}, and concept detection \cite{sb-feat}.
+
+\subsubsection{Spatio-temporal visual features (C3D)}
+\label{c3d-feat}
+These features are extracted from the C3D model, a 3-dimensional convolutional network proposed for generic video analysis \cite{c3d-feat}. 
+The main motivation to use C3D is that it encodes both the spatial and temporal information in the video.
+The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
+We use the publicly available model trained on the Sports-1M dataset \cite{c3d-feat}.
+We use the output of the fully connected layer -- fc6 of the network with a dimensionality of 4096.
+Since such dimensionality is high compared to other features, we explore the use of principal component analysis (PCA) for the dimension reduction.
+
+\subsubsection{Audio features (AudioSet)}
+\label{as-feat}
+Using a recently released large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat} and the corresponding event detection models, we extract 128-dimensional embeddings for each audio track associated with a video in our dataset.
+We use these embeddings for training the regression models.
+The motivation to use these features is that they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.
+
+\subsubsection{Emotion related features (SentiBank and Affect)}
+\label{emo-feat}
+As research in psychology showed that emotion and memory are highly correlated \cite{emo-mem}, we investigate the use of emotion-related feature in our prediction system.
+For emotion from visual content, we resort to a visual sentiment concept detector: SentiBank \cite{sb-feat}.
+SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
+We use the binary code for concept detection, from images, provided by the authors.
+We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
+For each of these 10 frames, we run the SentiBank concept detector and obtain two pieces of information: concepts with probabilities and features.
+Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
+Examples of some concepts in the SentiBank ontology are: young driver, scary face, terrible pain, etc.
+We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
+We extract 300-dimensional word2vec \cite{word2vec} embeddings, for each of the 50 concepts and take an average to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
+Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
+In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors.
+
+Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video to obtain arousal and valence scores \cite{affect}.
+Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video invokes positive or negative emotion, following a circumplex model of affect \cite{affect-model}.
+For each frame in the video, we compute the arousal and valence scores using the method proposed in \cite{affect}.
+In order to keep a fixed dimensionality of the feature vector, we take the first 200 frames in the video because of the varying frame rates across the videos.
+We concatenate the arousal and valence scores for the first 200 frames in each video resulting in a 400-dimensional feature vector (200 for arousal and 200 for valence) for a video.
+
+\subsubsection{Visual semantic features (Image captions)}
+\label{sem-feat}
+Visual semantics are known to play an important role in memorability \cite{sem-mem}.
+We utilize the state-of-the-art research in image captioning to capture the semantics \cite{caption-feat}.
+We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
+For each of these 10 frames, we run the caption detector (code provided by the authors) and obtain a caption for the frame.
+For each non-functional word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings and taking an average of all the words to obtain a single vector per frame.
+We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single 300-dimensional feature vector for each video.
+
+\subsection{Modelling and evaluation}
+\label{model-eval}
+We used the features discussed in Section \ref{feat-extr} to train an SVR model with two different kernels (linear and RBF) for the memorability score prediction. 
+We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ = ${0.1, 1, 10, 100, 1000}$ and $\gamma$ = ${0.01, 0.1, 1, 10, 100}$.
+The choice of SVR is guided by the small size of the dataset and we stick to the two standard kernels because our focus was mainly on identifying which features are more important for memorability prediction.
+We have chosen to go with the same classifier for all the features so as to ensure that the differences in performances is because of the feature themselves.
+
+Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
+Additionally, we also use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
+This metric gives us an indication of the correlation between the predicted memorability scores and the ground-truth memorability scores.
+
+\subsection{Memorability prediction results}
+\label{pred-res}
+%In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new videos.
+In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating all the feature vectors into a single feature vector.
+While concatenating the feature vectors, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set of features.
+%Each of the videos in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
+%In this way, we eliminate movies that were not seen by a majority of the users who participated in our experiment.
+%Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
+We pick the SVR model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $MSE$.
+We retrain the model with these parameters on the training set and evaluate on the final test set.
+In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
+
+We report the results of the prediction on both the validation and test sets, so as to investigate whether there is overfitting of the model.
+We also report only correlation scores: $SpCorr$ as there is not much change in the $MSE$ scores across different features.
+On an average, the $MSE$ score varies between 0.05 and 0.08 across different features.
+Table \ref{res-4-10-ann} reports the the results of prediction capability of the model trained on different features.
+There are two sets of results reported in the table: $SpCorr$ (4 annotations) and $SpCorr$ (8 annotations).
+The first set of results correspond to the prediction capability of the model when trained on videos with at least 4 annotations and the second set corresponds to the results when trained on videos with at least 8 annotations.
+We will further discuss why we report the results with the model trained on videos with at least 8 annotations.
+Observing the first set of results in Table \ref{res-4-10-ann}, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
+Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
+These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
+Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
+Other features like AudioSet come close to the performance of the above features (third best feature), but only audio signal does not seem to be enough for predicting video memorability.
+Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features.
+One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector.
+In future, we could look at selectively combining the features to investigate if that improves the performance.
+
+One of our initial hypotheses was that emotion would play a important role in memorability of a video, supported by literature from psychology.
+We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
+Observing the scores in Table \ref{res-4-10-ann}, we can say that the SentiBank features (fourth row from bottom) perform slightly better than the affect features (sixth row from bottom).
+But the models trained on image caption features clearly out-perform those trained on emotion features.
+This could be because of the following two reasons: our choice of features to capture emotion related information is not suitable for our task, or the performance of the emotion prediction models is not good enough for memorability prediction.
+From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
+In the future, we could investigate more robust emotion/affect prediction models to improve the performance of memorability prediction.
+
+We further investigate the effect of number of annotations, for videos in the training set, on the performance of the model.
+Based on our experimental results, we observe that the image caption feature performs the best while using training videos that have at least 4 annotations.
+Based on this result, we train different SVR models with varying number of annotations on the training set using image caption features.
+For example, we train a model, using image caption features, on training sequences with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
+We repeat this process for different number of annotations: from 4 to 15 annotations increasing the number of annotations by 1 every time.
+Please note that the validation set in each of the repetitions is fixed and only the training set changes.
+We observed that the $MSE$ does not change significantly in each of the steps and hence we provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set.
+Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $SpCorr$ keeps going down.
+In the wake of this observation, we investigate the performance of all the features when we train the regression model with videos that have at least 8 annotations.
+
+\begin{figure}[h]	  
+  \centering
+    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
+		\caption{Spearmann correlation on the validation set with models trained using image caption features for varying number of annotations.}
+    \label{num-ann}
+\end{figure}
+
+\begin{table*}
+  \centering
+  \renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|c|c|c|c|c|c|c|}
+    \hline
+    \textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$(4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ (8 annotations)}}\\
+    % \hline
+    % \textbf{Inactive Modes} & \textbf{Description}\\
+    \cline{4-7}
+    & & & \emph{validation set}& \emph{test set} & \emph{validation set}& \emph{test set}\\ \hline
+    C3D& visual-spatio-temporal& 4096& 0.20& 0.26& 0.31& 0.34\\ \hline
+		C3D (PCA)& visual-spatio-temporal (PCA)& 225& 0.24& 0.21& 0.18& 0.17\\ \hline
+		AudioSet& audio-related& 128& 0.23& 0.22& 0.21& 0.24\\ \hline
+		Affect& affect-related& 400& 0.19& 0.17& 0.26& 0.23\\ \hline
+		SentiBank concepts& emotion-related& 300& 0.16& 0.13& 0.15& 0.17\\ \hline
+		SentiBank features& emotion-related& 4096& 0.25& 0.21& 0.27& 0.26\\ \hline
+		SentiBank features (PCA)& emotion-related (PCA)& 225& 0.22& 0.21& 0.22& 0.23\\ \hline
+		Captions& visual-semantics & 300& 0.29& 0.31& 0.39& 0.38 \\ \hline
+		Combination& combine-all-features (PCA)& 1578& 0.24& 0.23& 0.29& 0.27\\ \hline		
+    \hline
+  \end{tabular}
+  \caption{Prediction results (Spearmann correlation) on validation and test data for different features with models trained on sequences that have at least 4 (columns 4-5) or 10 (columns 6-7) annotations.}
+	\label{res-4-10-ann}	
+\end{table*}
+
+These results are reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ (8 annotations).
+Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the model trained on sequences with at least 8 annotations performs better than the model trained on sequences with at least 4 annotations.
+The only exception being the C3D (PCA) feature, where the model trained on sequences with at least 4 annotations performs marginally better than the model trained on sequences with at least 8 annotations.
+
+Additionally, we also observe that the scores for validation set and test set are very similar across the two results and hence we can say that there is no overfitting.
+We also compute the average number of annotations per sequence in the training and test set for both the cases (4 and 8 annotations) in order to check if there is a mismatch.
+On an average, each sequence in the test set has around 10 annotations and the test set remains the same across different experiments.
+Similarly, each sequence in the training set has 9 annotations on an average in the case where we train the models with sequences having at least 4 annotations and 11 annotations in the other case.
+Looking at these numbers, we can see that the average number of annotations per sequence is quite balanced across the training and test sets and there is indeed no mismatch.
 
@@
 %Inconvénient: mesure avec une part de subjectivité
+After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground-truth memorability scores.
+At this moment, we cannot claim that this would hold true for any other type of videos because we have only investigated one specific type of videos i.e., movies.
+Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
+We also found that emotion features are not that well correlated with ground-truth memorability scores, as indicated in psychology literature.
+An immediate direction for future research would be to explore other approaches to encode emotion in videos and investigate whether we can improve the prediction performance using these approaches.
+We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
 

### 2018-01-29 ARTICLE.tex
Karthik changes 29 january

@@
 \usepackage{graphicx}
-\usepackage{hhline}
-\usepackage{multirow}
 
@@
 \title{Annotating, understanding, and predicting long-term video memorability}
-\titlenote{Produces the permission block, and copyright information}
-\subtitle{Extended Abstract}
-\subtitlenote{The full version of the author's guide is available as
-\texttt{acmart.pdf} document}
+%\titlenote{}
+%\subtitle{}
+%\subtitlenote{}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Romain Cohendet}
+%author 1
+%\author{Romain Cohendet}
 %\authornote{}
 %\orcid{1234-5678-9012}
-\affiliation{%
-	\institution{Technicolor}
-	\streetaddress{}
-	\city{Rennes} 
-	\state{France} 
-	\postcode{}
-}
-\email{romain.cohendet@technicolor.com}
+%\affiliation{%
+%	\institution{Technicolor}
+%	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
+%	\city{Rennes} 
+%	\state{France} 
+%	\postcode{}
+%}
+%\email{romain.cohendet@technicolor.com}
 
-\author{Karthik Yadati}
-\affiliation{%
-	\institution{University of Delft}
-	\streetaddress{}
-	\city{Delft} 
-	\state{} 
-	\postcode{}
-}
-\email{N.K.Yadati@tudelft.nl}
+%author 2
+%\author{Karthik Yadati}
+%\affiliation{%
+%	\institution{University of Delft}
+%	\streetaddress{}
+%	\city{Delft} 
+%	\state{} 
+%	\postcode{}
+%}
+%\email{N.K.Yadati@tudelft.nl}
 
-\author{Ngoc Khanh Duong}
-\affiliation{%
-	\institution{Technicolor}
-	\city{Rennes} 
-	\country{France}}
-\email{Quang-Khanh-Ngoc.Duong@technicolor.com}
+%author 3
+%\author{Ngoc Khanh Duong}
+%\affiliation{%
+%	\institution{Technicolor}
+%	\city{Rennes} 
+%	\country{France}}
+%\email{Quang-Khanh-Ngoc.Duong@technicolor.com}
 
-\author{Claire-Helene Demarty}
-\affiliation{%
-	\institution{Technicolor}
-	\city{Rennes}
-	\country{France}
-}
-\email{Claire-Helene.Demarty@technicolor.com}
+%author 4
+%\author{Claire-Helene Demarty}
+%\affiliation{%
+%	\institution{Technicolor}
+%	\city{Rennes}
+%	\country{France}
+%}
+%\email{Claire-Helene.Demarty@technicolor.com}
+
+%%%%%%%%%%%%
+%%% for blind review %%%
+%author 1
+\author{Removed for review}
+%author 2
+\author{Removed for review}
+%author 3
+\author{Removed for review}
+%author 4
+\author{Removed for review}
 
@@
 \begin{abstract}
-Following on from the study of image memorability prediction, which draw continuous attention over the past six years, the computational understanding of video memorability's search field has recently hatched.
-The growing number of shared videos foster us to find new ways to make their occurrence the most relevant in our everyday lives.
-There is no available dataset of videos annotated in terms of memorability; such dataset may probably result in a launching of the field, as it had been the case for images.
-The first goal of the pioneers of this search should be to succeed to propose a protocol to constitute such a dataset. 
-%our work:
-In this article, we propose a protocol of 700 videos annotated with memorability scores, constitute a dataset, study the quality of the collected data, and computationally modeling memorability to predict it.
-%long-term memory
-
-We finally propose a deep-learning model base to predict video memorability, comparing several methods and class of features.
-
+The growing number of video contents on the Internet encourages us to find new ways to make more relevant their occurrences in our everyday life.
+In this context, memorability can be regarded as a useful metric of video importance to help us make a choice between competing videos.
+However, the research on computational understanding of video memorability is in its early stages.
+There is no available dataset for modelling purposes, and the few previous attempts provided protocols not generalizable to collect data at a large scale.
+Furthermore, the video computational features valuable to build a robust memorability predictor remain largely undiscovered.
+In this article, we propose a new protocol to collect long-term memorability annotations, that we use to measure memory performances of 104 participants from weeks to years after memorization.
+We then analyze the data collected for 660 videos by focusing on its quality, and test it against different characteristics such as response time, duration of memory retention and repetition of visualization.
+We finally conduct an extensive feature analysis, comparing several methods and classes of features, to propose a computational model for the prediction of video memorability.
 \end{abstract}
@@
 
-%% About the ICMR article on Media retrieval with video memorability
+%% About the ICMR article on Media retrieval with VM
 	% focus on "Multimedia content-based retrieval"
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{INTRODUCTION}%nb of sections %2 PAGES
+\section{Introduction}%nb of sections %2 PAGES
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Opening
-Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- contents.
-Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as a particularly relevant element to help us picking a video among several ones, with the advantage of a possibility to be measured less subjectively.
+%opening
+Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- digital contents.
+Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as useful to help us make a choice between competing videos.
+In addition, memorability has the advantage of being clearly definable and objectively measurable (i.e., using a measure that is not influenced by the observer's personal judgment).
+Thus, memorability has initially be defined as the probability for an image to be recognized \textbf{a few minutes} after a single view, when presented amidst a stream
+of images \cite{isola_2011_makes}.
+This definition has been widely accepted within subsequent work (e.g., \cite{mancas_2013_memorability,kim_2013_relative,celikkale_2013_visual,khosla_2015_understanding,lahrache_2016_bag}).
 
-%from images to video memorability
-The computational understanding of video memorability's search field follows on from the study of image memorability prediction which has attracted increasing attention from the seminal work of Isola \textit{et al.} released in 2001 \cite{isola_2011_makes}.
-Recently, models of image memorability prediction achieved very good results with the introduction of deep learning to address the challenge of image memorability prediction \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
-This success resulted on the extension of the challenge to videos.
-Video memorability prediction is however a new field, and there are only two studies, to our knowledge, that addressed this issue \cite{han_2015_learning,shekhar_2017_show}.
+%from images to VM
+The computational understanding of video memorability (VM) follows on from the study of image memorability prediction which has attracted increasing attention from the 2001 seminal work of Isola \textit{et al.} \cite{isola_2011_makes}.
+With the recent introduction of deep learning to address the challenge of image memorability prediction, models also achieved very good results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
+As a result of this success, researchers have recently extended this challenge to the videos.
+However, to the best of our knowledge, only two available studies focused on VM prediction \cite{han_2015_learning,shekhar_2017_show}.
 
-%No dataset available
-At least two important problems could explained this scarcity of studies.
+%explain the scarcity of studies on VM
+Several problems could explained this scarcity of studies.
+Firstly, there is no publicly available dataset to train and test models.
+This is probably the most serious obstacle to the rapid expansion of the VM prediction's search field.
+Accordingly, to provide researchers with ground truth data should be our very first objective, similarly with the work of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} which have enabled research on image memorability to flourish.
+The second point is closely related to the first one: there is no widely accepted definition of VM.
+The previous attempts to predict VM \cite{han_2015_learning,shekhar_2017_show} were based on different measures of memorability.
+Furthermore, contrary to images, videos do not constitute clearly defined units.
+They have supplementary dimensions -- sound and movement -- that critically contribute to the semantic and emotional information conveyed.
+If harmonized, the videos used and the way memorability is measured will have a critical impact on what we will understand by VM.
+The definition of image memorability by \cite{isola_2011_makes} had a great impact on subsequent work as aforementioned.
+But it also inevitably limited researchers.
+In particular, in previous research image memorability corresponds to memory performances measured only a few minutes after memorization.
+But these might be poor predictors of longer term memory performances (at least in some instances; e.g., for emotional images \cite{cohendet_2016_prediction}).
+Thus, VM data would benefit from a protocol that would measure lasting long-term memory performance.
 
-%1st point: no available dataset
-Firstly, there is no available dataset to train and test models.
-The fact that there exist no dataset to modeling video memorability is the most serious obstacle which prevent video memorability prediction to take off.
-This should be the very first goal of pioneers of this new research field: to provide data to community.
-For images, the pioneers created and made downloadable such a database \cite{isola_2011_makes}, which enable to made image memorability research flourishing.
-The second released database, far larger, launched new possibility for deep learning and enable to obtained the aforementioned really good results \cite{khosla_2015_understanding}.
-We could expect a similar success for video memorability prediction research.
-%A dataset as first objective to launch VM prediction --> Goal of this paper is to participant in the establishment of an adapted protocol
-We think that the existence of such a database with high quality data could have the importance for video memorability that the ones of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} had on image memorability.%in particular for deep learning algortihm which perform the best for images
+%modelling
+Regarding modelling, our capacity to propose efficient computational models is also important to meet the challenge of VM prediction. 
+The previous attempts at predicting VM \cite{han_2015_learning,shekhar_2017_show} shed light on several features which have a predictive power of VM.
+However, videos contain a great volume of data which can be used for feature extraction purposes, and the work is far for complete.
 
-%second point: difficulty to define video unity and VM
-This drives us to the second point: the difficulty to build such a dataset.
-The principle difficulty to define what video memorability and unity should be, and a protocol to measure items according to these definition.
+%objectives of the present study
+The aim of this article is to participate in the expansion of the VM prediction emerging search field.
+In particular, we:
+\begin{itemize}
+	\item propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect quality ground truth data (section 2);
+	\item assess and analyze the collected data to better understand VM (section 3);
+	\item model VM in an attempt to understand which computation features are better for this task (section 4).
+\end{itemize}
+In the next subsection, we review previous work on image and video memorability prediction, focusing on protocol and modelling questions.
 
-If we consider the studies on image memorability prediction, we could see that, with their database, the cardinal points defined in the very first study \cite{isola_2011_makes} were widely accepted by the community without re-interrogation of what memorability -- of what the models -- really predicted.
-First, it is because Isola \textit{et al.} find a proper and objective way to measure memory performance at a large scale; but it remains that how the authors proceeded is just one way to measure memorability.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Previous work}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Measurement of memorability}
+%Images work -- towards a long-term memorability
+Almost all studies on image memorability prediction made use of one or the other of the two available large datasets designed specifically to meet this challenge \cite{isola_2011_makes,khosla_2015_understanding}.
+To build these datasets, the authors, possibly constrained by the difficulty in conducting long crowdsourcing studies, measured memory performance a few minutes after memorization to obtain their memorability annotations.
+As proposed in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
+Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation, which lasts weeks to years \cite{mcgaugh_2000_memory}.
+%In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall follow an exponential decay, and is particularly strong in the few days immediately following the memorization \cite{ebbinghaus_1913_memory}.
+Because several factors influence the consolidation process (e.g., emotions, sleep, re-evocation), which does not equally affect all memories, the order of memorability ratings measured for videos is susceptible to change over time \cite{cohendet_2016_prediction}.
+A protocol to collect memorability annotations would benefit from the capacity to capture long-lasting memory performances, averaged to obtain what we will later refer to as "long-term memorability".
 
-From what we can imagine that such a dataset for VM will probably have an important influence on future research, one should define very carefully what memorability of video is and what the video as unity means.
-It requires to define the cardinal points that will guide -- and constraint -- researchers that will used the dataset.
-
-First, if image in image memorability was obvious, video in video memorability is less obvious.
-Videos are more complex than images.
-Contrary to images, videos do not constitute clearly defined units, but have supplementary dimensions -- sound and movement -- that makes difficult the definition of what a video is.
-%from ahn et al. Videos contain a synthesis of visual, aural, textual, and 	motion information and generally convey relatively richer 	media content than images.
-By an extreme example, is it possible for a model to transfer his capacity to predict memorability from movies-like videos to Youtube-like videos?
-
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Previous work} %Critics of the 2 previous studies
-%Critics of studies on image memorability : the problem of sustainable long-term memorability
-The work on image memorability prediction, initiated by \cite{isola_2011_makes}, produced great results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
-The availability of two large datasets \cite{isola_2011_makes,khosla_2015_understanding} has been critical in this achievement. 
-However, possibly constrained by the difficulty in conducting long crowdsourcing studies, the authors measured memory performance a few minutes after memorization to obtain their memorability annotations.
-As shown in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
-Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation \cite{mcgaugh_2000_memory}.
-In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall is particularly strong in the few days immediately following the memorization, after which it asymptotically tends its the almost permanent long-term performance \cite{ebbinghaus_1913_memory}.
-Because several factors are susceptible to influence the consolidation process (such as the emotional content, the semantics, the attention...), which is not a linear decrease for all the memories, it follows that the order of memorability of images is susceptible to change from some minutes to a day after memorization, as it has been found in \cite{cohendet_2016_prediction} with emotional image from of the International Affective Picture System \cite{lang_2005_international}.
-A protocol to collect memorability annotations for videos would benefit from the capacity to measure what we will later refer to as "long-term memorability".
-
-%Critics of Han et al.
-To our knowledge, the first attempt at predicting video memorability had been proposed by \cite{han_2015_learning}. 
+%Han et al.
+To our knowledge, the first attempt at predicting VM had been proposed by \cite{han_2015_learning}. 
 The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
@@
 They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
-We can infer from the sparse information authors gave that the task duration, for each of the 20 participants, was about 24 hours (by taking an average video duration half-way between 15 and 30 sec), spread over 10 sessions (5 free viewing and 5 recall tasks) of about 2 hours each.
-Authors used the same proportion of fillers (i.e.
-non repeated videos) in the free viewing and recall tasks (i.e.
-4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
-If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in free viewing task would have very little impact on memorability scores (often authors even use only material interrogated later in learning/free viewing task).
+The task duration, for each of the 20 participants, was about 24 hours, spread over 10 sessions (five free viewing tasks and five recall tasks) of about two hours each.
+The authors used the same proportion of fillers (i.e., non repeated videos) in the free viewing and recall tasks (i.e., 4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
+If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in the free viewing task would have very little impact on the memorability scores (often authors even use only material interrogated later in the learning/free viewing task).
 Furthermore, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
-Regarding the prediction part, we also agree with these authors to say that the method would be difficult to generalize as it used fMRI measurements for predicting memorability.
+Furthermore, authors measure memory after two days, but the consolidation process lasts weeks to years: it would be interesting to collect even longer term memory performance measures.
 
-%Critics of Shekhar et al.
+%Shekhar et al.
 Another earlier approach was the one of \cite{shekhar_2017_show}.
-The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequences a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on road?").
-The major drawback of the study comes from the use of questions instead of a recognition task.
-Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty, ease of understanding (in particular for the non-anglophone people that works with Amazon Mechanical Turk), of imaging, ease to retrieve a scene by some words more the by others...
+The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequence a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on a road?").
+The major drawback of the study comes from the use of questions instead of a classic visual recognition task.
+Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty
+The authors have tested complexity of textual questions, but their measure is not sufficient to represent the real difficulty of the question in its complexity (ease of understanding, in particular for the non-anglophone people that works with Amazon Mechanical Turk, ease of imaging, ease to retrieve a scene by some words more the by others...).
 Especially since the authors took into account the response time of the participants to calculate the memorability scores of the videos.
-One will note that this potential bias could have also affect the measure of inter-human consistency.
-Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limiting (e.g.
-the question "Did you see a car on road?" implies that in a whole experimental session you can have just one car on a road), and the authors manually ensured that no textual questions nor videos in a sequence were similar in content.
-This makes difficult the generalization of this protocol to build an extensive dataset.
-%%% ATT!! The have tested complexity of textual questions, but there is the other problems! mesure insufisante pour rendre compte de la difficulté réelle des questions.
+One will note that this potential bias could have also affected the measure of inter-human consistency.
+Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limited (e.g., the question "Did you see a car on a road?" implies that in a whole experimental session one can have just one car on a road).
+The authors also manually ensured that no textual questions nor videos in a session were similar in content.
+This makes it difficult to generalize this protocol to the construction of an extensive dataset.
 
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Goals of our study}
-%Conclusion critics
-It appears from the review of previous work that video memorability prediction research is in its early stage.
-We still need to find a proper way to collect quality data in order to build a large dataset to bring video memorability prediction research to a new level.
-%1/ we want a protocol
-The present study aims to be part of these efforts to reach such a way.
-As aforementioned, we also want this protocol to measure a "long-term" memorability.
-This can be achieved with a study that have a long time span, where learning and recall tasks are seperated by a certain amount of time, or by measuring a memory created prior to the experiment.
-Because crowdsourcing is a priori the only practicable way to experimentally collect great amount of annotations, we focused on the last choice more suitable for crowdsourcing as it requires only one connexion to the measurement application.
-%2/ we want an all-around paper
-To assess the quality of the data produced by the protocol, we analyzed the data obtained from different perspectives, in particular in order to shed light on some potential human or task-related factors susceptible to influence memorability as it is measured by the protocol, that are numerous \ref{cohendet_2016_using}.
-%3/ we want to meet the challenge of video memorability prediction
-Thanks to the data collected, we participate in the VM prediction challenge.
-%Conclusion intro
-The structure of the rest of the article follows these points: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and a conclusive part 5 where we talked about intended future work related to video memorability prediction.
+\subsection{Memorability modelling}
+%intro on focus on features to discover
+Previous attempts at predicting image and video memorability highlighted quite a few features that correlates with memorability.
 
-%%%%%%% POTENTIALLY %%%%%%%
-% Response time: must we exmploit it to calculate memorability scores?
- %Because of \cite{shekhar_2017_show} and other papers in images which suppose that short time response equal to more memorable ALORS qu'ils n'ont pas pour tâche de répondre le plus rapidement possible [vérifier].
- %La répartition du temps de réponse pourrait ne pas linéraire.
-	%We hypothesized that the most memorable the sequence, the faster the participant will answer.
-Parce que les video les mieux représentés dans l'esprit sont les plus accessibles et que les participants ne veulent pas louper la préponse quand ils sont sûrs.
-% Citer papiers - s'il en existe -- où les auteurs exploitent le temps de réponse pour corriger les scores de mémorabilité;
+%image memorability prediction
+The pioneering work of Isola \textit{et al.} focused primarily on building computational models to predict image memorability from low-level visual features \cite{isola_2011_makes}.
+It appeared from this first attempt that it is possible to predict to some extent the degree of picture's memorability.
+Several characteristics have been found to be relevant for predicting memorability in subsequent work, for example saliency features \cite{mancas_2013_memorability}, interestingness and aesthetic \cite{isola_2014_makes}, or emotions \cite{khosla_2015_understanding}.
+The best results was finally obtained by using fine-tuned deep features, which outperformed all other features in \cite{khosla_2015_understanding}, reaching a rank correlation of $.64$ which is near human consistency ($.68$) measured for ground truth collected in the study.
+This result was confirmed in \cite{baveye_2016_deep,squalli_2017_deep}.
 
-% Personalization of the data
- %Taking into account human factors %see results Logistic regression vs. SVM
+%VM prediction
+Regarding the VM prediction, Han \textit{et al.} propose a method which combines the power of audiovisual and fMRI-derived features \cite{shekhar_2017_show}.
+They preliminary built a computational model learned with fMRI features, which supposedly convey the brain activity of memorizing videos, which enable them to finally predict VM without fMRI scans.
+However, the method would be difficult to generalize
+
+Shekhar \textit{et al.} conducted a performance analysis of several computationally extracted features before building their memorability predictor \cite{shekhar_2017_show}.
+The analysis encompassed C3D deep learning feature for video classification, video semantics obtained thanks to video captioning method, saliency features, dense trajectories, and color features.
+They found that the most predictive feature combination used video semantics, spatio-temporal, saliency and color features.
+The feature that performed the best when tested alone was video semantics.
+The authors used a captioning method to generate a semantic description of the videos.
+The text generated by this method was then processed by a recursive auto-encoder network that outputted a 100-dimensional representation of the videos.
+Due to the particularity of the dataset of Shekhar \textit{et al.} -- that is, their aforementioned use of questions to measure memorability --, it would be interesting to confirm if this combination works equally well on another dataset, and in particular if images captioning features also perform the best.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Construction of the dataset}
+\section{Memorability dataset construction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Our first step was to implement a protocol to measure the long-term memorability of 700 video sequences, extracted from 100 mainstream movies.
-The main characteristic of the protocol, in contrast with previously proposed methods to collect image and video memorability annotations, is the absence of learning/free viewing task, replaced by a questionnaire designed to collect information about the participants' prior memory.
-The way of proceeding aimed at measuring memory performance after a significant retention period.
+\subsection{Video collection}
+%why did we choose movie
+We wanted our protocol to measure memory performance after a significant retention period.
+This can be achieved either with a longitudinal study, or by measuring a memory created prior to the experiment.
+We chose the latter because it enabled us to immediately measure very long-term memory.
+Thus, the main characteristic of the proposed protocol, in contrast with previous work, is the absence of learning (often, free viewing) task, replaced by a questionnaire designed to collect information about the participants' prior memory.
+
+%movies and videos selections
+ %with HD or dvdrip quality, in orginal (i.e.,english) version, without without subtitles [add if results HD vs. DVDrip]
+We first established a list of 100 movies, taking care to mix popularity and genres.
+We then manually selected seven videos of 10 seconds from each movie.
+To maintain a high intra-video semantic cohesion, we did not make cuts that would impair the understanding of the scene, nor did we aggregate shots that belong to different scenes.
+Indeed, since the semantics is linked to the memorability of images \cite{isola_2014_makes}, we can expect it is linked to the memorability of videos too.
+
+%neutral vs. typical
+We also gave preference to the videos we called "neutral", by contrast to the "typical" ones.
+According to our definition, a neutral video is a part of a movie which contains no element that would enable someone to easily guess it belongs to a particular movie.
+The list of undesirable elements includes but is not limited to: recognizable famous actors, typical music, style, etc.
+Typical videos are simply defined as non-neutral videos.
+In most movies, just a few or no 10-sec neutral videos exist.
+That explains why we obtained only 127 neutral videos for 573 typical ones (while we expected two neutral and five typical videos per movie).
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dataset}
-%% films and sequences
-We mixed popularity and genres to establish our list of movies.% occidental movies, except for Slumdog millionaire)%to check%(list added to the appendix [ref_appendix])
-Then, we manually selected seven sequences of 10 seconds from each movie (for a total of 700 sequences).
-%with HD or dvdrip quality, in orginal (i.e.english) version, without without subtitles [add if results HD vs.
-DVDrip]
+\subsection{Annotation protocol}
+%presentation protocol
+The protocol is composed of two tasks.
+Firstly, participants had to fill a questionnaire intended to collect data about whether they know the 100 selected movies.
+Secondly, participants performed a recognition task on videos selected from their responses to the questionnaire.
 
-%cuts
-We made efforts to maintain a high intra-sequence semantic cohesion; in particular, we did not cut through meaningful vectors nor we agglomerated shots that belonged to different scenes.
-Indeed, we think the images' characteristic by which they constitute obvious semantic units is an important element in the capacity of the prediction models to generalize.
-By contrast, the meaning a sequence conveyed was susceptible to change depending on where we decided to cut the movie.
+%participants, facilities and apparatus
+104 participants ($22-58$ years of age; $\mu_{age}=37.1 $; $\sigma_{age}=10.4$; $26\%$ of them female), mostly educated persons (engineers or researchers mainly), participated in the experiment on a volunteer basis.
+The experiment was taking place in a room insulated from noise and equipped with subdued lights.
+The videos, of HD or DVD quality, were displayed on a 60 inch monitor (Sony Bravia).
+The participants were seated at a distance of about 220 centimeters from the screen (that is three times the screen height).
 
-Because the semantics is linked to the memorability of images \cite{isola_2014_makes}, so probably to videos, therefore their memorabilities, we chose this definition susceptible to be extensive to predict other dataset.
-
-%PARCE QUE CEST UNE PETITE BASE DE DONNEES => POUR UNE PLUS GROSSE, CE PROBLEME N4AURA PAS ROP D'EFFET.
-
-%hence, in our idea, fostering semantics coherence of the sequences equal to diminish the difficulty of the computational models to learn without impairing their capacity to generalize.
-
-%neutral vs. typical
-We also gave preference to the sequences we called "neutral", by contrast to the "typical" ones.
-According to our definition, a neutral sequence is a part of a movie which contains no element that would enable someone to easily guess it belongs to a particular movie.
-
-The list of undesirable includes but is not limited to: recognizable famous actors, typical music, style, or any element...
-
-The question we asked ourselves to select the neutral sequence was the following: "Could I guess this sequence belongs to the movie I extracted it from if I didn't know it?"
-
-In most movies, just a few or no 10-sec sequences of this sort exist. For example, in Kill Bill 1, we can
-
-That explains why, at the end, we obtained only 127 neutral sequences for 573 typical ones, that are defined as being non neutral sequences that meet the first criterion.
-
-These two types of sequences -- neutral \textit{vs.} typical -- should be considered in conjunction with the instruction of the task:
-
-neutral sequences are control.
-
- of the task written above (even if you know that a sequence comes from a movie, just answer if you recognize THE sequence):
-
-They only are used as controls is just a supplementary control (qui joue un rôle symétrique au taux de FA pour la prise de risque).
-
-%%Indoor/Outdoor
- %In Isola et al., 2014, the authors showed that memorability directly correlated with color.
-To explain this result, they advanced the hypothesis that the difference in memorability between indoor and outdoor scenes was due to the fact that the later tend to be less memorable that the older, and that the color tend to linked with a differences in colors, could explain this result.
-To validate this hypothesis, we manually annotated our 700 sequences as consisting of an indoor or outdoor scene.
-
-%fig with video frame memorability scores ranked from most t oless memorable.
-
-% potentially
- %First, not to be part of the official trailers, to be sure that, if a participant report not having seen the movie, he won't have seen the sequence presented for the recognition in a trailer.
-=> Bof: this doesn't matter if False Alarms becauses of seen trailer.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Survey design}
 %questionnaire
-After they had provided basic demographics, participants filled out a questionnaire composed of 100 items, indicating whether they had seen the films of our list.
-Each item included a film poster with its name below, followed by the question: "Do you remember watching (fully) this movie?" The experimenter was present in case the participant had doubt about one movie.
-If the participant answered "yes", he had to answer three additional questions: "Please let us know how confident you are of having seen this movie." (I am not/slightly/50\%/considerably/100\% confident); "Please let us know when did you last see this movie." (<1 month; <1 year; <5 years; < 10 years; > 10 years); "Please let us know how many times did you watch this movie." (once; 2-4 times; 5-9 times; 10-19 times; >20 times).
-If he answered "no", he had to answer one more question: "Please let us know how confident you are of NOT having seen this movie". (I am not/slightly/50\%/considerably/100\% confident).
-The question relative to confidence aimed to ensure the targets and fillers corresponded to seen and non-seen films respectively.
-The other questions (i.e.
-when and how many times the films had been seen) were used during result analysis to evaluate the likely effects of visualization repetition and passage on long-term memorability.
+Having provided basic demographics, participants filled out a questionnaire on the selected movies.
+For each of them, they were asked whether they remembered watching fully the movie.%Each item included a movie poster with its name below, followed by the question.
+In case of a positive answer, three additional questions followed on: 1/ their confidence of watching the movie (Not confident / slightly confident / 50\% confident / considerably confident / 100\% confident), 2/ the duration since they saw the movie (less than month / 1 year / 5 years / 10 years / more than 10 years), and 3/ the number of times they saw the movie (once / 2-4 times / 5-9 times / 10-19 times / more than 20 times).
+In case of a negative answer, only one question followed on, relating to their confidence of not having seen the movie.
 The questionnaire required about 20 minutes to complete.
@@
 %Algorithm selection
-Based on the answers to the questionnaire, an algorithm selected 80 targets and 40 fillers among the films associated with the highest degree of certitude, with a maximum of two sequences from the same film.
-Another criterion of selection for the target was the number of annotations of the sequences, in order to harmonize the nb of annotations per video.
-The sequences were presented in a random order for presentation
+Based on the answers to the questionnaire, an algorithm selected 80 targets and 40 fillers (i.e.
+videos from never seen movies) among the movies associated with the highest degree of certitude, with a maximum of two videos from the same movie.
+Another selection criterion was the videos' number of annotations, so that it was balanced.
 
 %Recognition task
-The recognition task lasted 24 minutes. Participants saw 10-seconds sequences separated by an inter-stimuli interval of 2 second. Their task was to press the spacebar when they recognized a sequence.
-During the instruction, the experimenter insisted on the fact that the only task of the participant was to answer only when he recognized the particular sequence, but not answer if he recognized not even when he could guess that the sequence came from a film he had seen.
-This, because for number of sequences (i.e.
-the typical ones), it was possible to guess that the sequence came from a particular film.
-Even if the participant answers, the video continue to run until its end.
+The questionnaire was followed by the recognition task itself, in which participants saw the 120 randomly chosen 10-seconds videos, separated by an inter-stimuli interval of 2 seconds.
+They had to press the space bar when they recognized a video in particular, and not when they guessed that the particular video came from a movie they had seen (which was possible only for the typical videos).
+In case a participant pressed the space bar, the video continued until its end.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Participants, facilities and apparatus}
-104 participants ($22-58$ years of age; $mean = 37.1 $; $SD = 10.4 $; $26\%$ of them female), employees of [removed for double-blind review] participated in he experiment on a volunteer basis.
-% Technicolor, Rennes (France).
-All participants have either normal or corrected-to-normal visual acuity.
-The videos were displayed on a ?????
-inch monitor (TV SONY Bravia ?????) with a display resolution of 1,920 $\times$ 1,080.
-The participants were seated at a distance of 150 centimeters from the screen (three times the screen height) in a room equipped with subdued lights.
-%The 1,024 $\times$ 768 videos were centered on a black background; at a viewing distance of 150 cm, the stimuli subtended 18.85 degrees of vertical visual angle.
+\subsection{Memorability scores calculation}
+%first discuss why we filtered the data from 700 to 660 and give other statistics.
+After collecting the data, we kept only the 660 videos that had been seen at least 4 times as targets (from the initial set of 700 videos).
+On average, each video of our dataset has been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enter in the calculation of a memorability score.
+Each video has also been viewed as a filler by $10.5$ participants.
+
+%Then explain how the scores are computed
+We then assigned a memorability score to each video, defined as the correct recognition rate of the video.
+The average percentage of correct detections for all participants was $46.71\%$ ($\sigma=14.65\%$), and the average false alarm (i.e., answer on a filler) rate was $4.16\%$ ($\sigma=5.27\%$).
+Figure \ref{fig:response_time_vs_memorability}(a) provides a distribution of the videos according to their degree of memorability.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{STUDY OF THE MEMORABILITY ANNOTATIONS}
+\section{Study of the memorability annotations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we analyze the ground truth data collected thanks to the protocol described in the previous section.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Memorability scores calculation}
-After collecting the data, we measured the average percentage of correct detections for all participants, of $46.71\%$ (SD of $14.65\%$), and the average false alarm rate, of $4.16\%$ (SD of $5.27\%$).
-Then we assigned a memorability score to each sequence, defined as the percentage of correct detections by participants.
-We retained only the 660 sequences that had been seen at least 4 times as targets, and remove the sequences chosen from films seen by too few participants.
-On average, each sequence was viewed as target by $10.7$ participants, and as filler by $10.5$ participants.
+In this section, we conduct an analysis of the ground truth data collected thanks to the protocol described in the previous section.
+We firstly focus on quality of data through a human consistency on memorability analysis, and by comparing neutral videos with typical videos.
+We then test the data against different characteristics such as response time, duration of memory retention and repetition of visualization.
+In what follows, error bars in the graphs correspond to standard error of the mean, $\mu$ to the mean, $\sigma$ to the standard deviation ans $N$ to the number of observations in the statistics.
 
-The figure \ref{fig:overview} provides a sample of sequences sorted according to their memorability (a) and a distribution of the sequences according to their degree of memorability (b).
-
-%Examples of ++/-- memorable sequences and Distribution of memorability scores 
-\begin{figure}[!htbp]
-	\centering
-%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/key_frame_mem_examples.png}}
-	\quad
-%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/memorability_scores_distribution.png}}
-	\quad
-	\caption{\label{fig:overview}(a) Sample of the dataset with key frames of sequences sorted from most memorable (left) to less memorable (right). (b) Distribution of the memorability scores.}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Consistency analysis}
@@
 We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency.
-We randomly split our 104 participants into two independent halves, and calculate how well sequence memorability scores from the first half of the participants match with sequence memorability scores from the second half of the participants.
-Averaging over 25 random split half trials, we calculated a Spearman's rank correlation $\rho$ of $0.57$ between these two sets of scores.
-%Individual and contextual differences, besides mandatory random variability, explain the $1-.57$ part of the memorability that is not universally derivable from the intrinsic informations of the videos.
+We randomly split our 104 participants into two independent halves, and calculate how well video memorability scores from the first half of the participants match with video memorability scores from the second half of the participants.
+Averaging over 25 random half-split trials, we calculated a Spearman's rank correlation $\rho$ of $0.57$ between these two sets of scores.
 
 %Fig Consistency
-%Histogram of number of nb of sequences per possibble nb of annotations
+ %Histogram of number of nb of videos per possibble nb of annotations
  %Curve of human consistency
@@
 	\quad
-%	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Measure_of_human_consistency.png}}
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Measure_of_human_consistency.png}}
 	\quad
-	\caption{\label{fig:human_consistency}(a) Number of sequences for each possible nb of annotations per sequence.
-(b) Human consistency averaged over 25 random splits (right) obtained using the method proposed by \cite{isola_2014_makes} for sequences with at least 5 annotations (with standard error and linear trendline).}
+	\caption{\label{fig:human_consistency}(a) Number of videos for each possible number of annotations per video.
+(b) Human consistency averaged over 25 random splits obtained using the method proposed by \cite{isola_2014_makes} (with linear trendline).}
 \end{figure}
@@
 %per nb of annotations consistency %comparison with image memorability
-We reproduced this calculation to obtain 25 Spearman's correlation coefficients as a function of the mean number of scores per video, presented in figure \ref{fig:human_consistency}(b).
-This curve is to be compared with the histogram presented in figure \ref{fig:human_consistency}(a), which shows that the number of sequences for each number of annotations was unequal.
-According to the curve, we achieved a mean consistency of $.70$ around 18 annotations, which is consistent with the previous attempt of \cite{han_2015_learning}, were it is the maximum consistency they achieve.
-It is interesting to compare these results to the one obtained by authors that have collected image memorability annotations \cite{isola_2011_makes,khosla_2015_understanding}.
-In these studies, the authors reached such a correlation of $.70$ after 80 annotations per image, which is much later.
-The experimental protocols and conditions are different between all these studies; in particular, emphasis must be placed on two points.
-First, in the present studies and \cite{han_2015_learning}, we measured a long-term memory performance later than two days after the memorization, that is to say after the biggest losses in memory performance, as was previously stated.
-By contrast, \cite{isola_2011_makes,khosla_2015_understanding} measured memory performance from dozens of seconds to a few minutes after memorization.
-Second, videos memorability annotations were collected through in-lab experiments, and images annotations through crowdsourcing experiments.
-These two confound factors may have contributed to the shortest number of annotations necessary to reach a high human consistency.
-However, it would be interesting in future work to confirm if an important difference exists between images and videos regarding to the number of annotations necessary to achieve a high human response consistency.
-Apart from the conclusions we could draw about the universality of the intrinsic memorability of videos compare to images, this would mean that the magnitude of the work to carry out to build an extensive database for video memorability prediction is substantially smaller than one could expect from work on image memorability prediction.
+We reproduced this calculation to obtain 25 Spearman's correlation coefficients as a function of the mean number of annotations per video, presented in figure \ref{fig:human_consistency}(b).
+This curve is to be compared with the histogram presented in figure \ref{fig:human_consistency}(a), which shows that the number of videos for each number of annotations was unequal.
+According to the curve, we achieved a mean consistency of $.70$ from about 18 annotations, which is consistent with the previous attempt of \cite{han_2015_learning}.
+$.70$ corresponds also to the maximum consistency obtained when collecting image memorability scores \cite{isola_2011_makes,khosla_2015_understanding}, but for a much bigger number of annotations (80) per image.
+It must be noted that the protocols are different between the image memorability experiments conducted in \cite{isola_2011_makes,khosla_2015_understanding} and ours or the work in \cite{han_2015_learning}.
+We conducted a measure of long-term memory performance after at least two days of memorization, whereas in \cite{isola_2011_makes,khosla_2015_understanding} it is measured after a dozen of seconds to a few minutes.
+In addition, VM annotations were collected through in-lab experiments, and images annotations through crowdsourcing experiments.
+These factors might have contributed to the shortest number of annotations necessary to reach a high human consistency for videos.
+However, it would be interesting in the future to confirm if an important difference exists between images and videos regarding the number of annotations necessary to achieve a high human response consistency.
+Apart from the conclusions we could draw about the universality of the intrinsic memorability of videos compared to images, this would mean that the magnitude of the work to carry out to build an extensive database for VM prediction is substantially smaller than one could expect from work on image memorability prediction.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Neutral and typical sequences}
-% Tester s'il y a une différence entre les deux et en tirer des conséquence quant à leur validité --> How objective our measure of memorability is?
-We select two sorts of sequences from the films: neutral and typical ones.
-For neutral sequences, participant had no element to guess that the sequence came from a particular film; for typical sequence, participants could guess that the sequence came from a particular film.
-The instructions was made clear, that we wanted that participants answered when they recognized a sequence, not when they guess that they came from a particular film they answered during the previous questionnaire film they had seen.
-Thus, the response for the neutral sequences is objective, but there is subjectivity in the typical sequence.
+\subsection{Neutral and typical videos}
+In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie which title was proposed in the questionnaire.
+We perform an analysis to compare neutral videos, which contain no element that would enable participants to guess that a video belongs to a particular movie, and typical videos.
+Indeed, if neutral videos received objective answers from participants, it might be more subjective for typical videos, that could be more or less easily related to the movie they belong to.
 
-A non-parametric Wilcoxon rank sum test showed a significant difference in memorability between the neutral ($\mu=.24$) and typical ($\mu=.53$) sequences: $Z=10.22, p<.00001$.
-We expected a result in this sens because neutral sequence contained less contextual elements useful to recognize that a segment comes from a particular film.
-Because of this compounding factor this results does not mean that participants tended to guess -- rather to purely recognize -- sequences drawn from films they have seen.
+A Wilcoxon rank-sum test indicated that the memorability was greater for neutral ($Mdn=.2, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, $Z=10.22, p<.00001$.
+Apart from the subjectivity aspect, we expected such a result because neutral videos contain less contextual elements, useful for recognition.
+Thus, this result does not necessarily mean that participants tended to guess -- rather to purely recognize -- videos drawn from movies they have seen.
 
-We also observed a difference in human consistency for memorability responses between the neutral ($\mu=.45$) and typical sequences ($\mu=.41$): $Z=2.75, p<.01$.
-The human congruency was slightly higher for neutral than for typical sequences.
-Along with the comments collected from the participants, who have as a majority reported difficulty to know if they guess or just recognized, this suggests that human congruence is higher for the objective measure than for a one with a part of subjectivity.
-In another context than with colleagues of our company, who highly probably respected the instructions -- in particular in crowdsourcing, which as we aforementioned stated that it must be used to built an extensive database -- it is probable that this difference would increase.
-Certainly, this result has to be confirmed by further studies, but it could be one of the weakness of our protocol to collect extensive data.
+A Wilcoxon rank-sum test indicated that the human consistency on memorability was slightly greater for neutral ($Mdn=.44, \mu=.45$) than for typical ($Mdn=.40, \mu=.41$) videos, $Z=2.75, p<.01$.
+Along with the comments collected from the participants, who have as a majority reported difficulty to know if they were guessing or really recognizing the videos, this result suggests that human congruency is higher for more 'objectively' recognized segments than for ones with subjectivity as part of the equation.
+One could note there is probably not just pure subjectivity here, but also a bias: some participants could have helped themselves with the context.
+If confirmed, this would constitute a weakness of our protocol to collect extensive data, that one should in that case counteract by adapted measures of control.
 
-We can also note that the false alarm rate was low for neutral sequences ($\mu=.05$) as well as for typical sequences ($\mu=.03$).
-Specifically, we expect lucky confusions to account for little of correct detections on average for the two sorts of sequences.
+We can also note that the false alarm rate was low for neutral videos ($\mu=.05$) as well as for typical videos ($\mu=.03$).
+Specifically, we expect lucky confusions to account for little of correct detections on average for the two sorts of videos.
 
@@
 \subsection{Response time}
-%figure Mean response time vs. degrees of memorability
-	%with SEM
+%figure Distribution of memorability scores + Mean response time vs. degrees of memorability
 \begin{figure}[!htbp]
 	\centering
-%	\includegraphics[width=\columnwidth]{figures/Response_time_vs_Memorability.png}
-	\caption{\label{fig:response_time_vs_memorability}Mean response time for each memorability degree for targets for which participants answered(error bars correspond to SEM).}
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/memorability_scores_distribution.png}}
+	\quad
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/Response_time_vs_Memorability.png}}
+	\quad
+	\caption{\label{fig:response_time_vs_memorability}(a) Distribution of the memorability scores. (b) Mean response time for correct recognitions against videos' memorability scores.}
+\end{figure}
+%(a) Sample of the dataset with key frames of videos sorted from most memorable (left) to less memorable (right).
+
+Figure \ref{fig:response_time_vs_memorability} (b) shows that the response time to do a correct detection decreases when the memorability of the video increases.
+We also observed a Pearson's correlation of $-0.36$ ($p<.0001$) between the response time on the targets and their memorability scores.
+These two results indicate that participants tended to answer quicker when the videos were more memorable, even though the participants did not receive any instruction to answer quickly.
+This suggests that people tend to naturally answer rapidly after having recognized the video.
+This also suggests either that the most memorable videos are also the most accessible in memory, and/or that the most memorable videos contain more early recognizable elements than the less memorable ones.
+In \cite{shekhar_2017_show}, the response time of the participants was taken to be the measure of video memorability.
+The authors chose this measure to avoid a long gap between viewing and recall stage.
+Our results validate -- to some extent -- their modus operandi: the fact that response time decreases linearly when the memorability increases suggest that the response time is a good indicator of the memorability of the videos (at least, in a recognition task).
+ %However, the linear continuous decrease in response time when memorability score values increase supports the first hypothesis: one can hardly imagine that the early recognizable elements (e.g., the face of a person, a particular place...) decreases linearly with the memorability of our 10-sec semantically consistent videos.
+ %The hypothesis of a linear relation between the accessibility of the videos in memory and their memorability is consistent with the associative functioning of long-term memory \cite{tulving_1974_cue}.
+ %Information stored in long-term memory is retrieved by way of association with other memories.
+ %One can hypothesize that the memorability of a video depends on how interconnected with peoples' prior memory the memories of this video is: in our experiment, the most memorable videos would have been the most easily accessible in memory, thus the most rapidly retrieved.
+%target vs. fillers
+ %We observed no significant correlation ($r=-0.09, p=.276$) between the mean response time to do a false alarm and the memorability score of the sequence.
+ %In conjunction with our previous hypothesis, only early recognizable elements could have played a role if we had observed such a correlation, not the accessibility in memory (because fillers are, by definition, videos that are not in memory).
+ %Let us imagine that the videos' early recognizable elements explain alone that the most memorable videos are recognized quickly than the less memorable ones.
+ %If we assume that recognizable elements are responsible of the false alarms by misleading the participants, we could have expected that early recognizable elements would have encouraged the participants to wrongly recognized fillers quickly when they were most memorable, since, as was suggested, people tend to naturally answer rapidly after having recognized the video.
+ %But we observed no effect of early recognizable elements on response time for fillers, while we observed an effect of either this factor of accessibility in memory for targets, in the same conditions (outside of this factor).
+ %This suggests that the presence of early recognizable elements does not explain the relation we observed between the response time and the memorability of the target videos, and thus that the hypothesis of the accessibility in memory is better.
+%response time for correct recognitions vs. false alarms 
+ %A Wilcoxon rank-sum test indicated that the response time was greater to do a correct recognition ($Mdn=4.5, \mu=4.88 sec$) than to do a false alarm ($Mdn=5.8, \mu=5.90 sec$), $Z=-5.10, p<.00001$.
+ %One explanation is that participants tended to hesitate more before generating a false alarm than when correctly recognizing a target.
+ %It points out the fact that memories of videos are generally not obvious but more blurred.
+ %This is an interesting point to connect with the fact that human consistency is lower for typical than for neutral videos: because of the non obviousness of if they had memories or not, the part of subjectivity for typical videos could have fostered participants to answer when they hesitated if they were guessing that a video came from a movie they had seen or remembering it.
+ %The higher degree of memorability of typical videos compare to the one for neutral videos could be partly due to this factor too.
+ %Among other, this is also an argument in favour of the most "objective" possible measure to choose to constitute an extensive dataset for VM prediction.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{User context and memorability}
+To provide us with an estimation context-related factors collected through our questionnaire, we processed to a logistic regression, using demographics and answers to the questionnaire as regressors, and the detection of a target video (with two possible discrete outcomes, detected or not) as observations to fit.
+Regarding the participants' nationality, we grouped them into occidental (69 pers) and non-occidental (35 pers) categories, motivated by our use of occidental movies, which could have more meaning for occidental than for non-occidental people.
+We also tested age and gender to reveal a potential bias in our movies' choice, maybe more interesting for people of a certain age and gender.
+The results of the logistic regression are shown in Figure \ref{fig:logistic_regression} (a).
+The method also provides a measure of the statistical significance of each feature in the model through their \textit{p}-value.
+
+The model, in case of a single observation, can be written as:
+\begin{equation}
+	y_n=\beta_0 + \sum\limits_{k=1}^K x_{nk} \beta_k+\epsilon_n
+\end{equation}
+where $y$ is the dependent variable -- the probability to correctly recognize a video --, $x$ are our predictor values, $\beta$ are the coefficients to be estimated, and $\epsilon$ indicates the error term.
+
+%figure logistic regression (a), and movie last viewing and nb of viewing (b)
+\begin{figure}[!htbp]
+	\centering
+	\subfloat[]{\includegraphics[width=0.75\columnwidth]{figures/logistic_regression.png}}
+	\\
+	%\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/film_last_viewing.png}}
+	\quad
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/nb_of_viewing.png}}
+	\quad
+	\caption{\label{fig:logistic_regression}(a) Features regression coefficients for the probability of detecting the repetition of a video (Features significance: *$p<.0001$).
+(b) Mean memorability depending on when occurred the last viewing, and (c) how many times the movie had been seen.}
 \end{figure}
 
-The figure \ref{fig:response_time_vs_memorability} shows that the response time to do a correct detection decreases when the memorability of the sequence increase.
-We also observed a  Person's correlation of $-0.36 (p<.0001)$ between the response time on the target and their memorability scores.
-These result show that the participants tended to answer quicker when the sequences were memorable than non memorable, even though the participants did not receive any instruction to answer quickly.
-This suggest two things.
-First, that participants tends to naturally answer rapidly after having recognized the sequence.
-Second, either that the most memorable sequences are also the sequences the most accessible in memory, or that the most memorable sequences have more early recognizable elements than less memorable sequences.
-However, the way we select sequences and the linear decrease in response time with memorability increasing showed in figure \ref{fig:response_time_vs_memorability} go in the direction of the first hypothesis.
+%memorability change over time
+Firstly, the retention duration is highly negatively correlated with the probability to recognize a video.
+Figure \ref{fig:logistic_regression} (b) shows that this decrease in memory for videos over time is continuous.
+This result indicates that long-term memory of videos continue to be altered over times for years.
+It implies that a memorability score, to provide an accurate representation of an average long-term memory performance, should correspond to a memory measure carried out as late as possible after the memorization.
 
-%target vs. fillers
-It is interesting to link it to the response time for answer on targets (i.e. correct detections) vs. on fillers (i.e. false alarms).
-The global mean response time was $4.88 sec$ on targets and $5.90 sec$ on fillers.
-A Wilcoxon rank sum test showed a significant difference ($Z=-5.10, p<.00001$), in the way participants globally answered more rapidly for targets (i.e.
-correct detections) than for fillers (i.e.
-false alarm).
-One explanation is that participants hesitated more for fillers they answered on than to make a correct detection, increasing their response time.
-This hesitation would reinforce the hypothesis that the most memorable sequences are most easily accessible in memory: if it has just been a question of a difference in the early recognizable elements that had explained that participants answered faster for the most memorable sequences, and not a question of hesitation -- i.e.
-of accessibility in memory --, the response time would have been the same for false alarm than for correct detection.%consistent with memory accessibility in psychology
-%L’étude de la morphologie des distributions nous enseigne que le temps de réponse tend à être uniforme pour les fillers, alors que les participants répondent plus vite pour les fillers.
+%memorability and viewving repetition
+Secondly, the number of views is highly correlated with the probability to recognize a video.
+As expected, the more a movie was seen, the better the videos were memorized.
+Figure \ref{fig:logistic_regression} (c) shows that this continues to be true even with more than 9 viewings.
+(However, the number of observations was very low (12) for videos which belongs to movies with 10 or more views.)
+One should note that the repetition of a viewing could not be the (only) factor involved in the above phenomenon; in particular, viewing again a movie may be the sign of a special interest which would explained a better memorization (e.g., via a greater attentional and emotional investment).
+The fact remains that repetition is an important factor to ask people when measuring their prior memory.
+Furthermore, a protocol used to build an extensive database for VM prediction should, in case of multiple measures of memory (e.g., after the memorization and then after a longer delay), avoid measuring twice the same items, because this repetition could artificially increase the performance measured for the last items.
 
-It also point out the fact that memories of videos are generally not obvious but more blurred.
-This is an interesting point to connect with the fact that human consistency is lower for typical than for neutral sequences: because of the non obviousness of if they had memories or not, the part of subjectivity for typical sequences could have fostered participants to answer when they hesitated if they were guessing that a sequence came from a film they had seen or remembering it.
-The higher degree of memorability of typical sequences compare to neutral ones could be partly due to this factor too.
-Among other, this is also an argument in favour of the most "objective" possible measure to choose to constitute an extensive dataset for video memorability prediction.
-
-To conclude this section, the degree of memorability sounds to be naturally related to response time.
-Therefore, we could imagine to exploit in the calculation of memorability score.
-That is what \cite{shekhar_2017_show} did in their previous attempt to construct their dataset to predict video memorability: they integrated response time in the calculation of the memorability score of their videos as the principal element (even if, in this case, the aforementioned problem of the use of questions could have influenced their results).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Logistic regression vs. SVM to personalize prediction model}
-%See QoMEX paper to present the data
-
-%figure Logistic regression
-\begin{figure}[!htbp]
-	\centering
-%	\includegraphics[width=\columnwidth]{figures/Logistic_regression.png}
-	\caption{\label{fig:logistic_regression}.}
-\end{figure}
-
-%We used the following factors:
-	%Occidental/non-occidental, because the 100 movies are occidental (expept slumdog millionaires...)
-
-%Corr Evolution of the memorability along time
- %Use order positions for that (e.g. mean memorability for position 1, 2... 120)
-
-%
+%Ajouter Age et expliquer que les movies sont plutôt jeune génération -- donner corrélation Age*Memorability => pas possible
+According to Figure \ref{fig:logistic_regression} (a), we observed no significant effects of the demographic factors (nationality, age and gender).
+This suggests that the videos were about equally susceptible to be recognized by the different participants (or, that the relations between these factors and the observations are too complex to have been captured by the model).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Mean memorability of the sequences compare to the film they come from}
-	%Corriger score par moyenne de mem du film -- est-ce que ça a un sens ? Par le nombre de vue du film ? par le moment où il a été vu?
+%\subsection{Mean memorability of the videos compare to the movie they come from}
+	%Corriger score par moyenne de mem du movie -- est-ce que ça a un sens ? Par le nombre de vue du movie ? par le moment où il a été vu?
 
@@
 
-%\subsection{Film genre and IMDB ratings}
-	%test the momorability od the sequences compare to the mean one of the movie.
+%\subsection{movie genre and IMDB ratings}
+	%test the momorability od the videos compare to the mean one of the movie.
 	%Correct bu the mean participant's performance
@@
 	%tester ces features comme dans 'What makes a photograph...'
-	% Indoor/Outdoor, low-level visal feautures (color... -- i.e. as in isola et al.), salinecy maps
+	% Indoor/Outdoor, low-level visal feautures (color... -- i.e., as in isola et al.), salinecy maps
 
@@
 %\subsection{Other points}
-	%add other relevant points found in MIT papers e.g.
+	%add other relevant points found in MIT papers e.g.,
+
+%Corr Evolution of the memorability along time for videos scores
+%Use order positions for that (e.g., mean memorability for position 1, 2... 120)
 
@@
 %Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
-% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the film --> No' --> So the point here is just to provide an overview.
+% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the movie --> No' --> So the point here is just to provide an overview.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{MEMORABILITY PREDICTION} %for Khartik
+\section{Memorability prediction} %for Khartik
 \label{mem-pred}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Until now, we have presented the video collection and the annotation protocol. 
 %We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
-In this section, we move towards building a machine learning model that can learn and then predict the memorability score of a video from its audio-visual features.
-The main goal of modelling is to identify which features: generic, perceptual, or semantic, are important for predicting the memorability scores of a video and \emph{not} the best learner/classifier.
+In this section, we move towards building a machine learning model that can learn and then predict the VM score of a video from its audio-visual features.
+The main goal of modelling is to understand if VM is predictable, and if yes identify which features: generic, perceptual, or semantic, are suitable for such prediction.
 We pose the problem as a standard regression problem and Figure \ref{prop-apprch} illustrates different steps in our method.
-First, we see that the videos are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new videos.
+%First, we see that the videos are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new videos.
 In the following sub-sections, we explain our choice of features and models to address the problem in hand.
@@
 %We explore a variety of audio-visual features in building a model that can predict memorability scores for new videos.
-The task of remembering a specific sequence from a movie has a high cognitive complexity in general, and is difficult for humans.
+The task of remembering a specific video has a high cognitive complexity in general, suggesting that it requires a semantic understanding of the content and/or some other perceptual factors such as the emotion conveyed by the video.
 Many users who participated in our experiments indicated that it is a difficult task.
-While trying to build a machine learning model for such a task, we wanted to explore different kinds of features that could be extracted from the audio-visual signal.
+While trying to build a machine learning model for such a task, we explore different kinds of features that can be extracted from the audio-visual signal.
 We investigate a variety of generic state-of-the-art features (\cite{c3d-feat}, \cite{audioset-feat}) and compare them with other semantic (\cite{caption-feat}) and perceptual (emotion) features (\cite{sb-feat}).
@@
 The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
-We use the publicly available model trained on the Sports-1M dataset \cite{c3d-feat}.
-We use the output of the fully connected layer -- fc6 of the network with a dimensionality of 4096.
-Since such dimensionality is high compared to other features, we explore the use of principal component analysis (PCA) for the dimension reduction.
+We use the publicly available model trained on the Sports-1M dataset \cite{c3d-feat} and extract the output of the fully connected layer -- fc6 of the network with a dimensionality of 4096.
+We additionally explore the use of principal component analysis (PCA) (named C3D (PCA) in Table \ref{res-4-10-ann}) for the dimension reduction, as the original dimensionality is very high when compared to other features.
 
@@
 \label{as-feat}
-Using a recently released large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat} and the corresponding event detection models, we extract 128-dimensional embeddings for each audio track associated with a video in our dataset.
+Using a recently released AudioSet \cite{audioset-feat} model, which was trained on a large dataset for event detection, we extract 128-dimensional embeddings for each audio track associated with a video in our dataset.
 We use these embeddings for training the regression models.
 The motivation to use these features is that they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.
+Additionally, we wanted to investigate how the audio channel contributes to building a model for VM prediction.
 
@@
 \label{emo-feat}
-As research in psychology showed that emotion and memory are highly correlated \cite{emo-mem}, we investigate the use of emotion-related feature in our prediction system.
-For emotion from visual content, we resort to a visual sentiment concept detector: SentiBank \cite{sb-feat}.
+As research in psychology showed that emotion and memory are correlated \cite{emo-mem}, we investigate the use of emotion-related feature in our prediction system.
+For modelling emotion from the visual content, we resort to a visual sentiment concept detector: SentiBank \cite{sb-feat}.
 SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
 We use the binary code for concept detection, from images, provided by the authors.
+The SentiBank concept detector provides two pieces of information: concepts with probabilities and features.
+Concepts are adjective--noun pairs and the probability represents how likely each concept is depicted visually in an image.
+Examples of some concepts in the SentiBank ontology are: \emph{young driver}, \emph{scary face}, \emph{terrible pain}, etc.
+
 We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
-For each of these 10 frames, we run the SentiBank concept detector and obtain two pieces of information: concepts with probabilities and features.
-Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
-Examples of some concepts in the SentiBank ontology are: young driver, scary face, terrible pain, etc.
-We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
-We extract 300-dimensional word2vec \cite{word2vec} embeddings, for each of the 50 concepts and take an average to obtain a single vector per frame.
+We run the SentiBank concept detector on each of these 10 frames and rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
+We extract a 300-dimensional word2vec \cite{word2vec} embeddings, for each of the 50 concepts and take an average to obtain a single vector per frame.
 We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
-Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
-In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors.
+Sentibank detectors also provide a 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
+In the end, we use a 300-dimensional concept vector and a 4096-dimensional feature vector.
 
-Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video to obtain arousal and valence scores \cite{affect}.
-Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video invokes positive or negative emotion, following a circumplex model of affect \cite{affect-model}.
+In addition to SentiBank concepts, we investigate other ways to capture emotional content in a video. 
+Following a circumplex model of affect (the experience of emotion) \cite{affect-model}, we define arousal as the dimension of affect that measures the excitement in the video, while valence measures whether the video invokes positive or negative emotion.
+We resort to an audio-visual analysis of the video to obtain its arousal and valence scores using the method described in \cite{affect}.
 For each frame in the video, we compute the arousal and valence scores using the method proposed in \cite{affect}.
@@
 \label{sem-feat}
-Visual semantics are known to play an important role in memorability \cite{sem-mem}.
-We utilize the state-of-the-art research in image captioning to capture the semantics \cite{caption-feat}.
+Visual semantics are known to play an important role in image memorability (\cite{sem-mem}, \cite{squalli_2017_deep}).
+We utilize the state-of-the-art research in image captioning to capture such high-level semantics of the video \cite{caption-feat}.
 We sample one frame for every second of the video in our dataset, resulting in 10 frames per video.
 For each of these 10 frames, we run the caption detector (code provided by the authors) and obtain a caption for the frame.
-For each non-functional word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings and taking an average of all the words to obtain a single vector per frame.
+For each non-functional word in the caption, we extract a 300-dimensional word2vec \cite{word2vec} embeddings and take an average of all the words to obtain a single vector per frame.
 We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single 300-dimensional feature vector for each video.
@@
 \label{model-eval}
-We used the features discussed in Section \ref{feat-extr} to train an SVR model with two different kernels (linear and RBF) for the memorability score prediction. 
-We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ = ${0.1, 1, 10, 100, 1000}$ and $\gamma$ = ${0.01, 0.1, 1, 10, 100}$.
-The choice of SVR is guided by the small size of the dataset and we stick to the two standard kernels because our focus was mainly on identifying which features are more important for memorability prediction.
-We have chosen to go with the same classifier for all the features so as to ensure that the differences in performances is because of the feature themselves.
+We use the features discussed in Section \ref{feat-extr} to train a Support Vector Regression (SVR) model with two different kernels (linear and RBF) for the VM score prediction. 
+We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ = $\{0.1, 1, 10, 100, 1000\}$ and $\gamma$ = $\{0.01, 0.1, 1, 10, 100\}$.
+We use the standard regression metric: Mean Squared Error ($MSE$) for the optimization process.
+The choice of SVR is guided by the small size of the dataset. 
+We have chosen to go with the same regressor for all the features because our focus was mainly on identifying which features are more important for VM prediction. 
+This way we ensure that the difference in performances is because of the features themselves.
 
-Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
-Additionally, we also use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
-This metric gives us an indication of the correlation between the predicted memorability scores and the ground-truth memorability scores.
+%Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
+We use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
+We chose this particular metric as it gives us an indication of how close the predicted memorability scores are to human labelled ground-truth memorability scores.
+
+In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating them into a single feature vector.
+While performing such a concatenation, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set.
+%Each of the videos in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
+%In this way, we eliminate movies that were not seen by a majority of the users who participated in our experiment.
+%Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
+We pick the SVR model parameters: kernel (linear or RBF), $C$ and $\gamma$ that give the best performance on the validation set.
+We retrain the model with these parameters using the training set and evaluate on the final test set.
+In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
 
@@
 \label{pred-res}
-%In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new videos.
-In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating all the feature vectors into a single feature vector.
-While concatenating the feature vectors, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set of features.
-%Each of the videos in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
-%In this way, we eliminate movies that were not seen by a majority of the users who participated in our experiment.
-%Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
-We pick the SVR model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $MSE$.
-We retrain the model with these parameters on the training set and evaluate on the final test set.
-In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
+In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new videos.
+We report the results of the prediction on both the validation and test sets.
+We also report only correlation scores: $SpCorr$ and not $MSE$ keeping in mind the space constraints.
+Additionally, $SpCorr$, which provides a rank correlation between the predicted and the ground-truth memorability scores, is more informative.
+We compute the average number of annotations per video in the train and test set for both the cases ($geq$ 4 and $geq$ 8 annotations).
+We observe that the number of annotations in the train and test sets are balanced and there is no mismatch.
+For example, each sequence in the test set has around 10 annotations while there are 9 annotations for each sequence in the training on an average.
 
-We report the results of the prediction on both the validation and test sets, so as to investigate whether there is overfitting of the model.
-We also report only correlation scores: $SpCorr$ as there is not much change in the $MSE$ scores across different features.
-On an average, the $MSE$ score varies between 0.05 and 0.08 across different features.
 Table \ref{res-4-10-ann} reports the the results of prediction capability of the model trained on different features.
-There are two sets of results reported in the table: $SpCorr$ (4 annotations) and $SpCorr$ (8 annotations).
+There are two sets of results reported in the table: $SpCorr$ ($\geq$ 4 annotations) and $SpCorr$ ($\geq$ 8 annotations).
 The first set of results correspond to the prediction capability of the model when trained on videos with at least 4 annotations and the second set corresponds to the results when trained on videos with at least 8 annotations.
 We will further discuss why we report the results with the model trained on videos with at least 8 annotations.
+
 Observing the first set of results in Table \ref{res-4-10-ann}, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
@@
 Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
+
 Other features like AudioSet come close to the performance of the above features (third best feature), but only audio signal does not seem to be enough for predicting video memorability.
@@
 
-One of our initial hypotheses was that emotion would play a important role in memorability of a video, supported by literature from psychology.
-We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
+One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
+Please recall that we used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
 Observing the scores in Table \ref{res-4-10-ann}, we can say that the SentiBank features (fourth row from bottom) perform slightly better than the affect features (sixth row from bottom).
-But the models trained on image caption features clearly out-perform those trained on emotion features.
-This could be because of the following two reasons: our choice of features to capture emotion related information is not suitable for our task, or the performance of the emotion prediction models is not good enough for memorability prediction.
-From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
-In the future, we could investigate more robust emotion/affect prediction models to improve the performance of memorability prediction.
+But the models trained on image captioning features clearly out-perform those trained on emotion features.
+This could be because of the following reasons: our choice of features to capture emotion related information is not suitable for our task, or the performance of the emotion prediction models is not good enough for memorability prediction, or we are not able able to establish the correlation between emotion and memorability because of the limited size of our dataset.
+%From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
+%In the future, we could investigate more robust emotion/affect prediction models to confirm that both emotion and memorability are correlated as suggested in the literature \cite{emo-mem}.
 
-We further investigate the effect of number of annotations, for videos in the training set, on the performance of the model.
-Based on our experimental results, we observe that the image caption feature performs the best while using training videos that have at least 4 annotations.
-Based on this result, we train different SVR models with varying number of annotations on the training set using image caption features.
-For example, we train a model, using image caption features, on training sequences with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
-We repeat this process for different number of annotations: from 4 to 15 annotations increasing the number of annotations by 1 every time.
+We further investigate the effect of the number of annotations, for videos in the training set, on the performance of the model.
+%Based on our experimental results, we observe that the image captioning feature performs the best while using training videos that have at least 4 annotations.
+%For this purpose, we train different SVR models with varying number of annotations on the training set using image captioning features only.
+We train a model, using image captioning features, on training sequences with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
+We repeat this process for different number of annotations per video in the training set: from 4 to 15 annotations.
 Please note that the validation set in each of the repetitions is fixed and only the training set changes.
-We observed that the $MSE$ does not change significantly in each of the steps and hence we provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set.
-Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $SpCorr$ keeps going down.
+We provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set in Figure \ref{num-ann}.
+We observe that $SpCorr$ first increases up to 5 annotations and then remains constant before decreasing (beyond 8 annotations).
 In the wake of this observation, we investigate the performance of all the features when we train the regression model with videos that have at least 8 annotations.
@@
     \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
-		\caption{Spearmann correlation on the validation set with models trained using image caption features for varying number of annotations.}
+		\caption{Spearmann correlation on the validation set with models trained using image captioning features for varying number of annotations.}
     \label{num-ann}
@@
     \hline
-    \textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$(4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ (8 annotations)}}\\
+    \textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$($geq$ 4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ ($geq$ 8 annotations)}}\\
     % \hline
@@
     & & & \emph{validation set}& \emph{test set} & \emph{validation set}& \emph{test set}\\ \hline
-    C3D& visual-spatio-temporal& 4096& 0.20& 0.26& 0.31& 0.34\\ \hline
-		C3D (PCA)& visual-spatio-temporal (PCA)& 225& 0.24& 0.21& 0.18& 0.17\\ \hline
-		AudioSet& audio-related& 128& 0.23& 0.22& 0.21& 0.24\\ \hline
-		Affect& affect-related& 400& 0.19& 0.17& 0.26& 0.23\\ \hline
-		SentiBank concepts& emotion-related& 300& 0.16& 0.13& 0.15& 0.17\\ \hline
-		SentiBank features& emotion-related& 4096& 0.25& 0.21& 0.27& 0.26\\ \hline
-		SentiBank features (PCA)& emotion-related (PCA)& 225& 0.22& 0.21& 0.22& 0.23\\ \hline
+    C3D& visual spatio-temporal& 4096& 0.20& 0.26& 0.31& 0.34\\ \hline
+		C3D (PCA)& visual spatio-temporal (PCA)& 225& 0.24& 0.21& 0.18& 0.17\\ \hline
+		AudioSet& audio related& 128& 0.23& 0.22& 0.21& 0.24\\ \hline
+		Affect& affect related& 400& 0.19& 0.17& 0.26& 0.23\\ \hline
+		SentiBank concepts& emotion related& 300& 0.16& 0.13& 0.15& 0.17\\ \hline
+		SentiBank features& emotion related& 4096& 0.25& 0.21& 0.27& 0.26\\ \hline
+		SentiBank features (PCA)& emotion related (PCA)& 225& 0.22& 0.21& 0.22& 0.23\\ \hline
 		Captions& visual-semantics & 300& 0.29& 0.31& 0.39& 0.38 \\ \hline
@@
   \end{tabular}
-  \caption{Prediction results (Spearmann correlation) on validation and test data for different features with models trained on sequences that have at least 4 (columns 4-5) or 10 (columns 6-7) annotations.}
+  \caption{Prediction results (Spearmann correlation) on validation and test data for different features with models trained on videos that have at least 4 (columns 4-5) or 8 (columns 6-7) annotations.}
 	\label{res-4-10-ann}	
@@
 
-These results are reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ (8 annotations).
-Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the model trained on sequences with at least 8 annotations performs better than the model trained on sequences with at least 4 annotations.
+These results are reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ ( $\geq$ 8 annotations).
+Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the models trained on sequences with at least 8 annotations perform better than the models trained on sequences with at least 4 annotations.
 The only exception being the C3D (PCA) feature, where the model trained on sequences with at least 4 annotations performs marginally better than the model trained on sequences with at least 8 annotations.
-
-Additionally, we also observe that the scores for validation set and test set are very similar across the two results and hence we can say that there is no overfitting.
-We also compute the average number of annotations per sequence in the training and test set for both the cases (4 and 8 annotations) in order to check if there is a mismatch.
-On an average, each sequence in the test set has around 10 annotations and the test set remains the same across different experiments.
-Similarly, each sequence in the training set has 9 annotations on an average in the case where we train the models with sequences having at least 4 annotations and 11 annotations in the other case.
-Looking at these numbers, we can see that the average number of annotations per sequence is quite balanced across the training and test sets and there is indeed no mismatch.
+Additionally, we also observe that the scores for validation set and test set are very similar across the two sets of results.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{CONCLUSIONS AND FUTURE WORKS}
+\section{Conclusions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%avantages de notre dataset: long-term memorabity
-%Inconvénient: mesure avec une part de subjectivité
+%protocol
+We proposed a new protocol to collect long-term memorability annotations for videos.
+It enabled us to measure memory performance after weeks to years.
+It appears from the analysis of the data that memory of videos continue to decrease for years, which justify a measurement of memory performance after a significant retention duration, longer than proposed in previous work.
+The principal weakness of our protocol is the part of subjectivity susceptible to enter in our measure of memorability, that one should counteract by appropriate controls.
+Our current work focus on collecting VM annotations at a large scale using crowdsourcing, measuring VM in a more objective manner.
+The implemented protocol measure memory performance after two different retention durations, which will enable us to understand what makes a video lastingly memorable.
+
+%modelling Karthik
 After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground-truth memorability scores.
-At this moment, we cannot claim that this would hold true for any other type of videos because we have only investigated one specific type of videos i.e., movies.
+It might be very interesting to investigate if these findings hold for generic video types other than movies.
 Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
@@
 
-%we propose to turn to a "totally objective" measure of memory to constitute a dataset for video memorability prediction
-
-% Nous construisons actuellement une base de données de grande ampleur avec une mesure purement objective de la mémoire.
-%Nulle part les auteurs ne parlent de mémoire à long terme, à court terme… => Ils appréhendent un problème par nature interdisciplinaire d’un point de vue purement computer science => Or la qualité des données est ce qui détermine ce que les modèles vont finalement prédire.
-// %D'abord, le test de mémoire interroge une mémoire de vidéos vues longtemps (en moyenne, probablement plusieurs mois ou années) auparavant.
-%cite QooMEX for personlization of memorability model
-
-% A conclusion section is not required.
-Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion.
-A conclusion might elaborate on the importance of the work or suggest applications and extensions.
-
-%%%
- %Liste des features (comme/ceux de MediaEval 2015 à donner au téléchargement avec le début en sec de nos sequences)
- %Release text file for movie title, start/end time + features
-
-%file to make the data + movies names available.
-
-%In particular, our protocol try to correct to integrate the concept of "very long-term memory", that escaped to previous studie in image and videos, which measured the memory of items some minutes after the memory encoding.
-
-%anntation des séquences pas à même vitesse, en cas de généralisation : changer les séquences annoté plus de 15 fois par exemple par de nouvelles.
-
-% Conclusion => Nos travaux futurs
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
@@
                                   % sure that you do not shorten the textheight too much.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{APPENDIX}
+%\section*{Appendix}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Appendixes should appear before the acknowledgment.
+ %Liste des features (comme/ceux de MediaEval 2015 à donner au téléchargement avec le début en sec de nos sequences)
+ %Release text file for movie title, start/end time + features
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{ACKNOWLEDGMENT}
+%\section*{Acknowledgment}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%This section is optional; it is a location for you
-%to acknowledge grants, funding, editing assistance and
-%what have you.
+%optional (%to acknowledge grants, funding, editing assistance and what have you.)
 

### 2018-02-01 ARTICLE.tex
Final version of article, submitted to ICMR

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 
@@
 \acmYear{2018}
-\copyrightyear{2018}
-
-\acmArticle{4}%??
-\acmPrice{15.00}%??
+%\copyrightyear{2018}
+%\acmArticle{4}%??
+%\acmPrice{15.00}%??
 
@@
 %author 1
-\author{Removed for review}
+%\author{Removed for review}
 %author 2
-\author{Removed for review}
-%author 3
-\author{Removed for review}
-%author 4
-\author{Removed for review}
+%\author{Removed for review}
 
 % The default list of authors is too long for headers.
-\renewcommand{\shortauthors}{R. Cohendet et al.}
+\renewcommand{\shortauthors}{Removed for blind review}%{R. Cohendet et al.}%Removed for blind review
 
@@
 \begin{abstract}
-The growing number of video contents on the Internet encourages us to find new ways to make more relevant their occurrences in our everyday life.
-In this context, memorability can be regarded as a useful metric of video importance to help us make a choice between competing videos.
-However, the research on computational understanding of video memorability is in its early stages.
-There is no available dataset for modelling purposes, and the few previous attempts provided protocols not generalizable to collect data at a large scale.
-Furthermore, the video computational features valuable to build a robust memorability predictor remain largely undiscovered.
-In this article, we propose a new protocol to collect long-term memorability annotations, that we use to measure memory performances of 104 participants from weeks to years after memorization.
-We then analyze the data collected for 660 videos by focusing on its quality, and test it against different characteristics such as response time, duration of memory retention and repetition of visualization.
-We finally conduct an extensive feature analysis, comparing several methods and classes of features, to propose a computational model for the prediction of video memorability.
+%The growing number of video contents on the Internet encourages us to find new ways to make more relevant their occurrences in our everyday life.
+Memorability can be regarded as a useful metric of video importance to help make a choice between competing videos.
+Research on computational understanding of video memorability is however in its early stages.
+There is no available dataset for modelling purposes, and the few previous attempts provided protocols to collect video memorability data that would be difficult to generalize.
+Furthermore, the computational features needed to build a robust memorability predictor remain largely undiscovered.
+In this article, we propose a new protocol to collect long-term video memorability annotations.
+We measure the memory performances of 104 participants from weeks to years after memorization to build a dataset of 660 videos for video memorability prediction.
+This dataset is made available for the research community.
+We then analyze the collected data in order to better understand video memorability, in particular the effects of response time, duration of memory retention and repetition of visualization on video memorability.
+We finally investigate the use of various types of audio and visual features and build a computational model for video memorability prediction.
+We conclude that high level visual semantics help better predict the memorability of videos.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\keywords{Video memorability, Long-term memory, Measurement protocol, Memorability scores, Deep learning, Multimedia information retrieval}
+\keywords{Video memorability, Long-term memory, Measurement protocol, Deep learning, Multimedia information retrieval}
 
@@
 %opening
-Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- digital contents.
-Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as useful to help us make a choice between competing videos.
+Enhancing the relevance of multimedia occurrences in our everyday life requires new ways to organize -- in particular, to retrieve -- digital content.
+Like other metrics of video "importance", such as aesthetics \cite{dhar_2011_high} or interestingness \cite{demarty_2016_mediaeval}, memorability can be regarded as useful to help make a choice between competing videos.
 In addition, memorability has the advantage of being clearly definable and objectively measurable (i.e., using a measure that is not influenced by the observer's personal judgment).
-Thus, memorability has initially be defined as the probability for an image to be recognized \textbf{a few minutes} after a single view, when presented amidst a stream
+Image memorability has initially been defined as the probability for an image to be recognized \textbf{a few minutes} after a single view, when presented amidst a stream
 of images \cite{isola_2011_makes}.
@@
 %from images to VM
-The computational understanding of video memorability (VM) follows on from the study of image memorability prediction which has attracted increasing attention from the 2001 seminal work of Isola \textit{et al.} \cite{isola_2011_makes}.
-With the recent introduction of deep learning to address the challenge of image memorability prediction, models also achieved very good results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
-As a result of this success, researchers have recently extended this challenge to the videos.
-However, to the best of our knowledge, only two available studies focused on VM prediction \cite{han_2015_learning,shekhar_2017_show}.
+The computational understanding of video memorability (VM) follows on from the study of image memorability (IM) prediction which has attracted increasing attention since the seminal work of Isola \textit{et al.} \cite{isola_2011_makes}.
+With the recent introduction of deep learning to address the challenge of IM prediction, models also achieved very good results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
+As a result of this success, researchers have recently extended this challenge to videos.
+However, to the best of our knowledge, only two available studies focus on VM prediction \cite{han_2015_learning,shekhar_2017_show}.
 
 %explain the scarcity of studies on VM
-Several problems could explained this scarcity of studies.
+Several problems could explain this scarcity of studies on VM.
 Firstly, there is no publicly available dataset to train and test models.
-This is probably the most serious obstacle to the rapid expansion of the VM prediction's search field.
-Accordingly, to provide researchers with ground truth data should be our very first objective, similarly with the work of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} which have enabled research on image memorability to flourish.
-The second point is closely related to the first one: there is no widely accepted definition of VM.
+This is probably the most serious obstacle to the rapid expansion of research in VM prediction.
+Following the foot steps of researchers in IM \cite{isola_2011_makes,khosla_2015_understanding}, providing data and ground truth for VM should be one of our first objectives.
+The second point, closely related to the previous one, is the lack of a common definition for VM.
 The previous attempts to predict VM \cite{han_2015_learning,shekhar_2017_show} were based on different measures of memorability.
-Furthermore, contrary to images, videos do not constitute clearly defined units.
-They have supplementary dimensions -- sound and movement -- that critically contribute to the semantic and emotional information conveyed.
-If harmonized, the videos used and the way memorability is measured will have a critical impact on what we will understand by VM.
-The definition of image memorability by \cite{isola_2011_makes} had a great impact on subsequent work as aforementioned.
-But it also inevitably limited researchers.
-In particular, in previous research image memorability corresponds to memory performances measured only a few minutes after memorization.
-But these might be poor predictors of longer term memory performances (at least in some instances; e.g., for emotional images \cite{cohendet_2016_prediction}).
-Thus, VM data would benefit from a protocol that would measure lasting long-term memory performance.
+Furthermore, in comparison to images, videos have supplementary dimensions -- sound and visual movement -- that critically contribute to the semantic and emotional information conveyed and this makes it difficult to come up with a common definition of VM.
+But the videos used and the way memorability is measured have a critical impact on what we understand by VM.
+Similarly, the definition of IM \cite{isola_2011_makes} inevitably limited researchers.
+In particular, previous research on IM focused on the measurement of memory performances only a few minutes after memorization.
+But these short-term memory performances might be poor predictors of longer term memory performances.
+For example, the memorability of emotional images would decrease in a non-harmonious way between two measures of memory performance (a few minutes and one day after memorization), inducing in a dataset a change in the ranking of memorability values over time \cite{cohendet_2016_prediction}.
+To this end, VM data would be expected to benefit from a protocol that would measure lasting long-term memory performance.
 
 %modelling
-Regarding modelling, our capacity to propose efficient computational models is also important to meet the challenge of VM prediction. 
-The previous attempts at predicting VM \cite{han_2015_learning,shekhar_2017_show} shed light on several features which have a predictive power of VM.
-However, videos contain a great volume of data which can be used for feature extraction purposes, and the work is far for complete.
+Regarding modelling, the previous attempts at predicting VM \cite{han_2015_learning,shekhar_2017_show} highlighted several features which contribute to the prediction of VM, such as semantic, saliency and color features.
+But the work is far from complete and our capacity to propose efficient computational models will participate to answer the challenge of VM prediction.
 
 %objectives of the present study
-The aim of this article is to participate in the expansion of the VM prediction emerging search field.
-In particular, we:
+While participating in the expansion of research on VM, the contributions of this paper are threefold:
 \begin{itemize}
-	\item propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect quality ground truth data (section 2);
-	\item assess and analyze the collected data to better understand VM (section 3);
-	\item model VM in an attempt to understand which computation features are better for this task (section 4).
+	\item We propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect ground truth data of good quality and make the corresponding dataset available for the research community (section \ref{section_dataset_construction}).
+	\item We assess and analyze the collected data and provide useful insights on the understanding of VM (section \ref{study_memorability}).
+	\item We build a computational model based on machine learning techniques which allows to predict VM score of a given video.
+For this purpose, we investigate the use of various types of audio and visual features, ranging from low-level characteristics to emotional and (highly) semantic features (section \ref{mem-pred}).
 \end{itemize}
-In the next subsection, we review previous work on image and video memorability prediction, focusing on protocol and modelling questions.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Previous work}
+\section{Previous work}
+In this section, we review previous work on image and video memorability prediction, focusing on annotation protocols and modelling aspects.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Measurement of memorability}
+%work in psychology
+Long-term memory has been studied for over a century in psychology, from multiple perspectives, since the seminal experimental studies of Ebbinghaus \cite{ebbinghaus_1913_memory} until the more recent neuro-imaging studies \cite{blumenfeld_2007_prefrontal}.
+This work provided researchers, interested in computational understanding of IM and VM, with several memory tests (see \cite{richardson_1988_measures} for an extensive overview), such as the recognition test \cite{isola_2011_makes,khosla_2015_understanding,han_2015_learning} or the textual questions-based recall survey \cite{shekhar_2017_show}.
+It also demonstrated that humans have an extensive long-term visual memory, which enables them to recall a great amount of images \cite{standing_1973_learning} and image details \cite{brady_2008_visual}, as well as of videos \cite{furman_2007_they}.
+This is one of the observations at the origin of the work on memorability in computer science \cite{isola_2011_makes}.
+Several factors have also been highlighted for their critical influence on long-term memory, such as emotion \cite{kensinger_2008_memory}, semantics \cite{quillan_1966_semantic}, several demographic factors \cite{cohendet_2016_using}, memory re-evocation \cite{nadel_1997_memory}, or passage of time \cite{mcgaugh_2000_memory}.
+These different factors are all important to better understand memorability, and are found valuable computational features for image and video memorability prediction.
+
 %Images work -- towards a long-term memorability
-Almost all studies on image memorability prediction made use of one or the other of the two available large datasets designed specifically to meet this challenge \cite{isola_2011_makes,khosla_2015_understanding}.
-To build these datasets, the authors, possibly constrained by the difficulty in conducting long crowdsourcing studies, measured memory performance a few minutes after memorization to obtain their memorability annotations.
-As proposed in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
-Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation, which lasts weeks to years \cite{mcgaugh_2000_memory}.
+Focusing on the work in computer vision, most studies on IM prediction made use of one of the two available large datasets designed specifically to meet this challenge \cite{isola_2011_makes,khosla_2015_understanding}.
+To address more specific problems, several other datasets have also been publicly released concerning memorability of face photographs \cite{bainbridge_2013_intrinsic}, visualization pictures \cite{borkin_2013_makes}, emotional images \cite{cohendet_2016_using} and scene categories \cite{bylinskii_2015_intrinsic}.
+To build the most used datasets presented in \cite{isola_2011_makes,khosla_2015_understanding}, the authors, possibly constrained by the difficulty in conducting long crowdsourcing studies, measured memory performance a few minutes after memorization to obtain their memorability annotations.
+This could be a problem if we conceive that memorability reflects a lasting memory performance \cite{cohendet_2016_prediction}.
+Indeed, it has been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation, which lasts weeks to years \cite{mcgaugh_2000_memory}.
 %In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall follow an exponential decay, and is particularly strong in the few days immediately following the memorization \cite{ebbinghaus_1913_memory}.
-Because several factors influence the consolidation process (e.g., emotions, sleep, re-evocation), which does not equally affect all memories, the order of memorability ratings measured for videos is susceptible to change over time \cite{cohendet_2016_prediction}.
+Because several factors influence the consolidation process (e.g., emotions, sleep, re-evocation), which does not equally affect all memories, the order of memorability ratings measured for content, and especially videos, is susceptible to change over time \cite{cohendet_2016_prediction}.
 A protocol to collect memorability annotations would benefit from the capacity to capture long-lasting memory performances, averaged to obtain what we will later refer to as "long-term memorability".
@@
 %Han et al.
-To our knowledge, the first attempt at predicting VM had been proposed by \cite{han_2015_learning}. 
-The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
-In contrast to the "memory game" proposed by Isola \textit{et al.} to collect memorability data, their protocol is however much heavier.
-They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
+To our knowledge, the first attempt at measuring VM \cite{han_2015_learning} partially adapted the protocol proposed to measure IM \cite{isola_2011_makes} to videos.
+The resulting protocol is however much heavier than the memory game protocol \cite{isola_2011_makes}.
+They followed a classical recognition process, which consists of two steps: a free viewing task, followed two days later by a recall task.
 The task duration, for each of the 20 participants, was about 24 hours, spread over 10 sessions (five free viewing tasks and five recall tasks) of about two hours each.
-The authors used the same proportion of fillers (i.e., non repeated videos) in the free viewing and recall tasks (i.e., 4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
-If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in the free viewing task would have very little impact on the memorability scores (often authors even use only material interrogated later in the learning/free viewing task).
-Furthermore, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
-Furthermore, authors measure memory after two days, but the consolidation process lasts weeks to years: it would be interesting to collect even longer term memory performance measures.
+The authors used the same proportion of fillers (i.e., non repeated videos) in the free viewing and recall tasks (i.e., 4/5 of fillers and 1/5 of repeated videos named targets) to guarantee that viewers were unaware of targets.
+If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is a way here to alleviate the task without impacting its quality; indeed, reducing the number of fillers in the free viewing task would have very little impact on the memorability scores (often authors even use only material interrogated later in the learning/free viewing task).
+Furthermore, the long time span of the experiment makes the generalization of this protocol difficult, in particular if one targets the construction of an extensive dataset.
+Moreover, authors measured memory after two days, but, as aforementioned, it is known that the consolidation process lasts weeks to years: it would be interesting to collect even longer term memory performances.
 
 %Shekhar et al.
-Another earlier approach was the one of \cite{shekhar_2017_show}.
-The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequence a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on a road?").
+In another earlier approach for VM measurement, the participants performed a crowdsourcing experiment that consisted of a free viewing task during which they saw a series of videos, followed by a recall task in which they had to answer textual questions (such as: "Did you see a man juggling?", "Did you see a car on road?") \cite{shekhar_2017_show}.
 The major drawback of the study comes from the use of questions instead of a classic visual recognition task.
-Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty
-The authors have tested complexity of textual questions, but their measure is not sufficient to represent the real difficulty of the question in its complexity (ease of understanding, in particular for the non-anglophone people that works with Amazon Mechanical Turk, ease of imaging, ease to retrieve a scene by some words more the by others...).
-Especially since the authors took into account the response time of the participants to calculate the memorability scores of the videos.
+Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty.
+The authors tested the complexity of the chosen textual questions using the Flesch-Kincaid Grade Level readability metric, which is designed to quantify how difficult to understand an English text is \cite{kincaid_1975_derivation}.
+However, this measure might not be sufficient to represent the questions in all their complexities, such as the ease for a person to draw images from the words, to connect the words to the associated concepts in memory, to associate the words to the corresponding scene, etc.
+Moreover, there is an unequal distribution of the reading comprehension of English among Amazon Mechanical Turk workers, who are not necessary anglophone people.
+This problem of a potential unequal complexity of the question becomes even more important in view of the use by authors of the response time to calculate memorability scores, which might also critically depend on the complexity of the questions.
 One will note that this potential bias could have also affected the measure of inter-human consistency.
-Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limited (e.g., the question "Did you see a car on a road?" implies that in a whole experimental session one can have just one car on a road).
+Furthermore, the questions were handcrafted, and the choices for types of questions and videos were very limited (e.g., the question "Did you see a car on road?" implies that in a whole experimental session only one car on a road should appear to connect the memory performance to one particular video).
 The authors also manually ensured that no textual questions nor videos in a session were similar in content.
@@
 %intro on focus on features to discover
-Previous attempts at predicting image and video memorability highlighted quite a few features that correlates with memorability.
+Previous attempts at predicting image and video memorability highlighted quite a few features that correlate with memorability.
 
-%image memorability prediction
-The pioneering work of Isola \textit{et al.} focused primarily on building computational models to predict image memorability from low-level visual features \cite{isola_2011_makes}.
-It appeared from this first attempt that it is possible to predict to some extent the degree of picture's memorability.
-Several characteristics have been found to be relevant for predicting memorability in subsequent work, for example saliency features \cite{mancas_2013_memorability}, interestingness and aesthetic \cite{isola_2014_makes}, or emotions \cite{khosla_2015_understanding}.
-The best results was finally obtained by using fine-tuned deep features, which outperformed all other features in \cite{khosla_2015_understanding}, reaching a rank correlation of $.64$ which is near human consistency ($.68$) measured for ground truth collected in the study.
-This result was confirmed in \cite{baveye_2016_deep,squalli_2017_deep}.
+%IM prediction
+The pioneering work of Isola \textit{et al.} focused primarily on building computational models to predict IM from low-level visual features \cite{isola_2011_makes}.
+From their work, it appeared that the degree of an image's memorability can be predicted to a certain extent.
+Several characteristics have also been found to be relevant for predicting memorability in subsequent work, for example saliency \cite{mancas_2013_memorability}, interestingness and aesthetics \cite{isola_2014_makes}, or emotions \cite{khosla_2015_understanding}.
+The best results were finally obtained by using fine-tuned deep features, which outperformed all other features in \cite{khosla_2015_understanding}, reaching a rank correlation of $.64$ which is near human consistency ($.68$) when measured for the ground truth collected in the study.
+This result was later confirmed by other researchers \cite{baveye_2016_deep,squalli_2017_deep}.
 
 %VM prediction
-Regarding the VM prediction, Han \textit{et al.} propose a method which combines the power of audiovisual and fMRI-derived features \cite{shekhar_2017_show}.
-They preliminary built a computational model learned with fMRI features, which supposedly convey the brain activity of memorizing videos, which enable them to finally predict VM without fMRI scans.
-However, the method would be difficult to generalize
-
+Regarding VM prediction, Han \textit{et al.} proposed a method which combines the power of audio-visual and fMRI-derived features \cite{han_2015_learning}.
+They preliminary built a computational model learned with fMRI features, which supposedly conveys the brain activity of memorizing videos.
+This enabled them to finally predict VM without the use of fMRI scans in a second step.
+However, the method would be difficult to generalize.
 Shekhar \textit{et al.} conducted a performance analysis of several computationally extracted features before building their memorability predictor \cite{shekhar_2017_show}.
-The analysis encompassed C3D deep learning feature for video classification, video semantics obtained thanks to video captioning method, saliency features, dense trajectories, and color features.
-They found that the most predictive feature combination used video semantics, spatio-temporal, saliency and color features.
-The feature that performed the best when tested alone was video semantics.
-The authors used a captioning method to generate a semantic description of the videos.
-The text generated by this method was then processed by a recursive auto-encoder network that outputted a 100-dimensional representation of the videos.
-Due to the particularity of the dataset of Shekhar \textit{et al.} -- that is, their aforementioned use of questions to measure memorability --, it would be interesting to confirm if this combination works equally well on another dataset, and in particular if images captioning features also perform the best.
+The analysis encompassed C3D deep learning features, semantic features obtained from some video captioning process, saliency features, dense trajectories, and color features.
+They found that the most predictive feature combination used captioning features, dense trajectories, saliency and color features.
+The features that performed the best when used alone were image captioning features, i.e., those conveying more semantics.
+Due to the particularity of the dataset of Shekhar \textit{et al.} -- that is, their aforementioned use of questions to measure memorability --, it would be interesting to confirm if this combination works equally well on another dataset, and in particular if image captioning features also perform the best.
 
@@
 \section{Memorability dataset construction}
+\label{section_dataset_construction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Video collection}
-%why did we choose movie
-We wanted our protocol to measure memory performance after a significant retention period.
+%why did we choose movies
+We want our protocol to measure memory performance after a significant retention period.
 This can be achieved either with a longitudinal study, or by measuring a memory created prior to the experiment.
 We chose the latter because it enabled us to immediately measure very long-term memory.
-Thus, the main characteristic of the proposed protocol, in contrast with previous work, is the absence of learning (often, free viewing) task, replaced by a questionnaire designed to collect information about the participants' prior memory.
+Thus, the main characteristic of the proposed protocol, in contrast with previous work, is the absence of learning (often, free viewing) task.
+It is replaced with a questionnaire designed to collect information about the participants' prior memory.
+In order to repeat measures of memory performances for different persons on the same material, to obtain average performance, we worked with movies famous enough to have been seen by several of our participants.
 
@@
  %with HD or dvdrip quality, in orginal (i.e.,english) version, without without subtitles [add if results HD vs. DVDrip]
-We first established a list of 100 movies, taking care to mix popularity and genres.
+ %sauf slumdog millionaire
+We first established a list of 100 occidental movies, taking care of mixing popularity and genres.
 We then manually selected seven videos of 10 seconds from each movie.
@@
 We also gave preference to the videos we called "neutral", by contrast to the "typical" ones.
-According to our definition, a neutral video is a part of a movie which contains no element that would enable someone to easily guess it belongs to a particular movie.
+According to our definition, a neutral video is a part of a movie which contains no element that would enable someone to easily guess this video belongs to a particular movie.
 The list of undesirable elements includes but is not limited to: recognizable famous actors, typical music, style, etc.
 Typical videos are simply defined as non-neutral videos.
-In most movies, just a few or no 10-sec neutral videos exist.
-That explains why we obtained only 127 neutral videos for 573 typical ones (while we expected two neutral and five typical videos per movie).
+In most movies, just a few or no 10-second neutral videos exist.
+That explains why we obtained only 127 neutral videos for 573 typical ones (while we expected two neutral and five typical videos per movie, i.e., 200 neutral and 500 typical videos in total).
 
@@
 The protocol is composed of two tasks.
-Firstly, participants had to fill a questionnaire intended to collect data about whether they know the 100 selected movies.
-Secondly, participants performed a recognition task on videos selected from their responses to the questionnaire.
+Firstly, participants had to fill in a questionnaire intended to collect data about whether they have seen the 100 selected movies.
+Secondly, participants performed a recognition task on videos selected based on their responses to the questionnaire.
 
 %participants, facilities and apparatus
-104 participants ($22-58$ years of age; $\mu_{age}=37.1 $; $\sigma_{age}=10.4$; $26\%$ of them female), mostly educated persons (engineers or researchers mainly), participated in the experiment on a volunteer basis.
-The experiment was taking place in a room insulated from noise and equipped with subdued lights.
-The videos, of HD or DVD quality, were displayed on a 60 inch monitor (Sony Bravia).
-The participants were seated at a distance of about 220 centimeters from the screen (that is three times the screen height).
-
+104 participants ($22-58$ years of age; age average = 37.1; stdev = 10.4; $26\%$ females; mostly educated persons -- engineers or researchers mainly), participated in the experiment on a voluntary basis.
+The experiment was taking place in a well-controlled environment: a room insulated from noise and equipped with subdued lights.
+The videos, of HD or DVD quality, were displayed on a 60 inch monitor.
+The participants were seated at a distance of about 220 centimeters from the screen (three times the screen height).
 %questionnaire
-Having provided basic demographics, participants filled out a questionnaire on the selected movies.
-For each of them, they were asked whether they remembered watching fully the movie.%Each item included a movie poster with its name below, followed by the question.
-In case of a positive answer, three additional questions followed on: 1/ their confidence of watching the movie (Not confident / slightly confident / 50\% confident / considerably confident / 100\% confident), 2/ the duration since they saw the movie (less than month / 1 year / 5 years / 10 years / more than 10 years), and 3/ the number of times they saw the movie (once / 2-4 times / 5-9 times / 10-19 times / more than 20 times).
-In case of a negative answer, only one question followed on, relating to their confidence of not having seen the movie.
+ %Each item included a movie poster with its name below, followed by the question.
+Having provided basic demographics, participants answered the questionnaire.
+For each of the 100 movies, they were asked whether they remembered watching fully the movie.
+In case of a positive answer, three additional questions followed: 1/ their confidence of watching the movie (\textit{not confident / slightly confident / 50\% confident / considerably confident / 100\% confident}), 2/ the last time they saw the movie (\textit{less than month / 1 year / 5 years / 10 years / more than 10 years}), and 3/ the number of times they saw the movie (\textit{once / 2-4 times / 5-9 times / 10-19 times / more than 20 times}).
+In case of a negative answer, only one question was asked, related to their confidence of not having seen the movie.
 The questionnaire required about 20 minutes to complete.
@@
 %Algorithm selection
-Based on the answers to the questionnaire, an algorithm selected 80 targets and 40 fillers (i.e.
-videos from never seen movies) among the movies associated with the highest degree of certitude, with a maximum of two videos from the same movie.
-Another selection criterion was the videos' number of annotations, so that it was balanced.
-
+Based on the answers to the questionnaire, an algorithm automatically selected 80 targets (i.e., videos from seen movies) and 40 fillers (i.e., videos from never seen movies) among the movies associated with the highest degree of certitude, with a maximum of two videos from the same movie.
+The fillers enable to quantify how much lucky confusions account for the correct recognitions.
+During the video selection, the current number of annotations was also taken into account to equally balance the annotations among all the videos.
 %Recognition task
-The questionnaire was followed by the recognition task itself, in which participants saw the 120 randomly chosen 10-seconds videos, separated by an inter-stimuli interval of 2 seconds.
-They had to press the space bar when they recognized a video in particular, and not when they guessed that the particular video came from a movie they had seen (which was possible only for the typical videos).
-In case a participant pressed the space bar, the video continued until its end.
+Given such 120 videos selected, participants performed a recognition task where they saw the videos separated by an inter-stimuli interval of 2 seconds.
+They had to press the space bar when they recognized a video in particular, and not when they were guessing that a particular video came from a movie they had seen (which was possible only for the typical videos).
+%In case a participant pressed the space bar, the video continued until its end.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Memorability scores calculation}
-%first discuss why we filtered the data from 700 to 660 and give other statistics.
+\subsection{Memorability score calculation}
+%Each video has also been viewed as a filler by $10.5$ participants.
 After collecting the data, we kept only the 660 videos that had been seen at least 4 times as targets (from the initial set of 700 videos).
-On average, each video of our dataset has been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enter in the calculation of a memorability score.
-Each video has also been viewed as a filler by $10.5$ participants.
-
-%Then explain how the scores are computed
-We then assigned a memorability score to each video, defined as the correct recognition rate of the video.
-The average percentage of correct detections for all participants was $46.71\%$ ($\sigma=14.65\%$), and the average false alarm (i.e., answer on a filler) rate was $4.16\%$ ($\sigma=5.27\%$).
+On average, each video of our dataset has been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enters in the calculation of a memorability score.
+We then assigned a memorability score to each video, defined as the correct recognition rate of the video when viewed as target.
+The average percentage of correct detections for all participants was $46.71\%$ (stdev = 14.65\%), and the average false alarm rate (i.e., the percentage of answers on fillers) was $4.16\%$ (stdev = 5.27\%).
 Figure \ref{fig:response_time_vs_memorability}(a) provides a distribution of the videos according to their degree of memorability.
+The dataset is publicly released here\footnote{\textit{www.remove-for-blind-review}.
+The original videos are not released because of copyright issues, but instead we provide the process we used to extract them from the movies and the features used in section \ref{mem-pred}.}.
 
@@
 \section{Study of the memorability annotations}
+\label{study_memorability}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this section, we conduct an analysis of the ground truth data collected thanks to the protocol described in the previous section.
-We firstly focus on quality of data through a human consistency on memorability analysis, and by comparing neutral videos with typical videos.
-We then test the data against different characteristics such as response time, duration of memory retention and repetition of visualization.
-In what follows, error bars in the graphs correspond to standard error of the mean, $\mu$ to the mean, $\sigma$ to the standard deviation ans $N$ to the number of observations in the statistics.
+In this section, we conduct an analysis of the ground truth data collected through the protocol described above.
+We firstly perform a human consistency analysis.
+Then we compare neutral videos with typical videos.
+We finally study which factors affect the memorability among response time, duration of memory retention and repetition of visualization.
+In what follows, error bars in the graphs correspond to standard error of the mean, $\mu$ to the mean, $Mdn$ to the median and $N$ to the number of observations in the statistics.
 
@@
 \subsection{Consistency analysis}
+\label{ssec:human}
 %global consistency calculation
-We implemented the method proposed in \cite{isola_2014_makes} to measure the human consistency.
-We randomly split our 104 participants into two independent halves, and calculate how well video memorability scores from the first half of the participants match with video memorability scores from the second half of the participants.
-Averaging over 25 random half-split trials, we calculated a Spearman's rank correlation $\rho$ of $0.57$ between these two sets of scores.
+We follow the method proposed in \cite{isola_2014_makes} to measure human consistency when assessing memorability of videos.
+We randomly split our 104 participants into two independent groups of equal size, and calculate how well video memorability scores from the first group of participants match with video memorability scores from the second group.
+Averaging over 25 random half-split trials, an average Spearman's rank correlation (i.e., a global human consistency) of $0.57$ is observed between these pairs of scores.
 
@@
  %Histogram of number of nb of videos per possibble nb of annotations
- %Curve of human consistency
+ %Graph of human consistency
 \begin{figure}[!htbp]
@@
 	\quad
-	\caption{\label{fig:human_consistency}(a) Number of videos for each possible number of annotations per video.
-(b) Human consistency averaged over 25 random splits obtained using the method proposed by \cite{isola_2014_makes} (with linear trendline).}
+	\caption{\label{fig:human_consistency}(a) Distribution of the number of videos according to the number of annotations per video. (b) Human consistency (with linear trendline).}
 \end{figure}
 
-%per nb of annotations consistency %comparison with image memorability
-We reproduced this calculation to obtain 25 Spearman's correlation coefficients as a function of the mean number of annotations per video, presented in figure \ref{fig:human_consistency}(b).
-This curve is to be compared with the histogram presented in figure \ref{fig:human_consistency}(a), which shows that the number of videos for each number of annotations was unequal.
-According to the curve, we achieved a mean consistency of $.70$ from about 18 annotations, which is consistent with the previous attempt of \cite{han_2015_learning}.
-$.70$ corresponds also to the maximum consistency obtained when collecting image memorability scores \cite{isola_2011_makes,khosla_2015_understanding}, but for a much bigger number of annotations (80) per image.
-It must be noted that the protocols are different between the image memorability experiments conducted in \cite{isola_2011_makes,khosla_2015_understanding} and ours or the work in \cite{han_2015_learning}.
+%per nb of annotations consistency %comparison with IM
+We reproduced this calculation to obtain human consistency as a function of the number of annotations per video, presented in Figure \ref{fig:human_consistency}(b).
+This graph is to be compared with the histogram presented in Figure \ref{fig:human_consistency}(a), which shows that the number of videos for each number of annotations was unequal.
+According to the graph, we achieved a consistency of $.70$ from about 18 annotations, which is consistent with the previous finding \cite{han_2015_learning}.
+This number also corresponds to the maximal consistency obtained when collecting IM scores \cite{isola_2011_makes,khosla_2015_understanding}, but for a much bigger number of annotations (80) per image.
+It must be noted that the protocols are different between the IM experiments conducted in \cite{isola_2011_makes,khosla_2015_understanding} and ours or the work in \cite{han_2015_learning}.
 We conducted a measure of long-term memory performance after at least two days of memorization, whereas in \cite{isola_2011_makes,khosla_2015_understanding} it is measured after a dozen of seconds to a few minutes.
-In addition, VM annotations were collected through in-lab experiments, and images annotations through crowdsourcing experiments.
-These factors might have contributed to the shortest number of annotations necessary to reach a high human consistency for videos.
-However, it would be interesting in the future to confirm if an important difference exists between images and videos regarding the number of annotations necessary to achieve a high human response consistency.
-Apart from the conclusions we could draw about the universality of the intrinsic memorability of videos compared to images, this would mean that the magnitude of the work to carry out to build an extensive database for VM prediction is substantially smaller than one could expect from work on image memorability prediction.
+In addition, VM annotations were collected through in-lab experiments, and IM annotations through crowdsourcing experiments.
+%These factors might have contributed to the smaller number of annotations necessary to reach a high human consistency for videos.
+However, it would be interesting in the future to confirm if an important difference exists between images and videos regarding the number of annotations necessary to achieve a high human consistency.
+Apart from the conclusions we could draw about the universality of the intrinsic memorability of videos compared to images, this would mean that the magnitude of the work to carry out to build an extensive database for VM prediction is substantially smaller than one could expect from work on IM prediction.
 
@@
 \subsection{Neutral and typical videos}
-In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie which title was proposed in the questionnaire.
-We perform an analysis to compare neutral videos, which contain no element that would enable participants to guess that a video belongs to a particular movie, and typical videos.
+In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie whose title was proposed in the questionnaire.
+In this section, we perform an analysis to compare neutral videos, which contain no element that would enable participants to guess that a video belongs to a particular movie, and typical videos.
 Indeed, if neutral videos received objective answers from participants, it might be more subjective for typical videos, that could be more or less easily related to the movie they belong to.
 
-A Wilcoxon rank-sum test indicated that the memorability was greater for neutral ($Mdn=.2, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, $Z=10.22, p<.00001$.
+A Wilcoxon rank-sum test indicated that the memorability was greater for neutral ($Mdn=.20, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, with $Z=10.22, p<.00001$.
 Apart from the subjectivity aspect, we expected such a result because neutral videos contain less contextual elements, useful for recognition.
-Thus, this result does not necessarily mean that participants tended to guess -- rather to purely recognize -- videos drawn from movies they have seen.
+Thus, this result does not necessarily mean that participants tended to guess -- rather than simply recognize -- videos selected from movies they have seen.
 
-A Wilcoxon rank-sum test indicated that the human consistency on memorability was slightly greater for neutral ($Mdn=.44, \mu=.45$) than for typical ($Mdn=.40, \mu=.41$) videos, $Z=2.75, p<.01$.
+As for the human consistency on memorability, a Wilcoxon rank-sum test indicated that it was slightly greater for neutral ($Mdn=.44, \mu=.45$) than for typical ($Mdn=.40, \mu=.41$) videos, with $Z=2.75, p<.01$.
 Along with the comments collected from the participants, who have as a majority reported difficulty to know if they were guessing or really recognizing the videos, this result suggests that human congruency is higher for more 'objectively' recognized segments than for ones with subjectivity as part of the equation.
-One could note there is probably not just pure subjectivity here, but also a bias: some participants could have helped themselves with the context.
+One could note there is probably not just pure subjectivity here.
+Context might have biased the results as it might have helped some participants in their recognition task.
 If confirmed, this would constitute a weakness of our protocol to collect extensive data, that one should in that case counteract by adapted measures of control.
 
-We can also note that the false alarm rate was low for neutral videos ($\mu=.05$) as well as for typical videos ($\mu=.03$).
+We can also note that the average false alarm rate (that is, the percentage of wrongly recognized filler videos) was low for neutral videos ($.05\%$) as well as for typical videos ($.03\%$).
 Specifically, we expect lucky confusions to account for little of correct detections on average for the two sorts of videos.
@@
 	\quad
-	\caption{\label{fig:response_time_vs_memorability}(a) Distribution of the memorability scores. (b) Mean response time for correct recognitions against videos' memorability scores.}
+	\caption{\label{fig:response_time_vs_memorability}(a) Distribution of the number of videos depending on memorability score ranges.
+(b) Mean response time for correct recognitions against memorability scores.}
 \end{figure}
@@
 We also observed a Pearson's correlation of $-0.36$ ($p<.0001$) between the response time on the targets and their memorability scores.
-These two results indicate that participants tended to answer quicker when the videos were more memorable, even though the participants did not receive any instruction to answer quickly.
+These two results indicate that participants tended to answer quicker when the videos were more memorable, even though the participants did not receive any instruction to do so.
 This suggests that people tend to naturally answer rapidly after having recognized the video.
@@
 The authors chose this measure to avoid a long gap between viewing and recall stage.
-Our results validate -- to some extent -- their modus operandi: the fact that response time decreases linearly when the memorability increases suggest that the response time is a good indicator of the memorability of the videos (at least, in a recognition task).
+Our results validate -- to some extent -- their modus operandi: the fact that the response time decreases linearly when the memorability increases suggest that the response time is a good indicator of the memorability of the videos (at least, in a recognition task).
  %However, the linear continuous decrease in response time when memorability score values increase supports the first hypothesis: one can hardly imagine that the early recognizable elements (e.g., the face of a person, a particular place...) decreases linearly with the memorability of our 10-sec semantically consistent videos.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{User context and memorability}
-To provide us with an estimation context-related factors collected through our questionnaire, we processed to a logistic regression, using demographics and answers to the questionnaire as regressors, and the detection of a target video (with two possible discrete outcomes, detected or not) as observations to fit.
+\subsection{User context and memorability}%\cite{walker_1967_estimation}
+To provide us with insights on which context-related factors collected through our questionnaire were linked to memorability, we processed to a logistic regression, using demographics and answers to the questionnaire as regressors, and the detection of a target video (with two possible discrete outcomes, detected or not) as observations to fit.
 Regarding the participants' nationality, we grouped them into occidental (69 pers) and non-occidental (35 pers) categories, motivated by our use of occidental movies, which could have more meaning for occidental than for non-occidental people.
-We also tested age and gender to reveal a potential bias in our movies' choice, maybe more interesting for people of a certain age and gender.
-The results of the logistic regression are shown in Figure \ref{fig:logistic_regression} (a).
-The method also provides a measure of the statistical significance of each feature in the model through their \textit{p}-value.
-
-The model, in case of a single observation, can be written as:
+We also tested age and gender to reveal a potential bias in our movies' choice, that may be more memorable for people of a certain age and gender.
+%The method provides a measure of the statistical significance of each feature in the model through their \textit{p}-value.
+The model, in case of a single observation $n$, can be written as:
 \begin{equation}
-	y_n=\beta_0 + \sum\limits_{k=1}^K x_{nk} \beta_k+\epsilon_n
+	y_{n}=
+	\begin{cases}
+	1 & \textit{if } \beta_0 + \sum\limits_{k=1}^K x_{n,k} \beta_k+\epsilon_n>0 \\
+	0 & else
+	\end{cases}
 \end{equation}
-where $y$ is the dependent variable -- the probability to correctly recognize a video --, $x$ are our predictor values, $\beta$ are the coefficients to be estimated, and $\epsilon$ indicates the error term.
+where $y_{n}$ denotes the dependent variable which can take two values, 1 for recognition or 0 for omission of the $n$-th target observation, $x_{n,k}$ our k-th feature value (last view, number of views, nationality, age, gender), $\beta_{k}$ are the coefficients to be estimated, and $\epsilon_{n}$ indicates the error term.
 
@@
 	\centering
-	\subfloat[]{\includegraphics[width=0.75\columnwidth]{figures/logistic_regression.png}}
-	\\
-	%\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/film_last_viewing.png}}
+	\subfloat[]{\includegraphics[width=0.45\columnwidth]{figures/film_last_viewing.png}}
 	\quad
@@
 	\quad
-	\caption{\label{fig:logistic_regression}(a) Features regression coefficients for the probability of detecting the repetition of a video (Features significance: *$p<.0001$).
-(b) Mean memorability depending on when occurred the last viewing, and (c) how many times the movie had been seen.}
+	\caption{\label{fig:logistic_regression}Correct recognition rate depending on (a) when occurred the last viewing, and (b) the number of views.}
 \end{figure}
@@
 %memorability change over time
-Firstly, the retention duration is highly negatively correlated with the probability to recognize a video.
-Figure \ref{fig:logistic_regression} (b) shows that this decrease in memory for videos over time is continuous.
-This result indicates that long-term memory of videos continue to be altered over times for years.
-It implies that a memorability score, to provide an accurate representation of an average long-term memory performance, should correspond to a memory measure carried out as late as possible after the memorization.
+According to the results of the logistic regression, the retention duration (last view) is highly negatively correlated with the probability to recognize a video ($\beta=-.37, p<.0001$).
+Figure \ref{fig:logistic_regression} (a) shows that this decrease in memory for videos over time is continuous.
+This result indicates that long-term memory of videos continues to be altered over time for years.
+In the context or experiment, it implies that, to provide an accurate representation of an average long-term memory performance, a memorability score should correspond to a memory measure carried out as late as possible after the memorization.
+%depends on the application and which type of memory is concerned.
 
 %memorability and viewving repetition
-Secondly, the number of views is highly correlated with the probability to recognize a video.
+The results of the logistic regression also show that the number of views is highly correlated with the probability to recognize a video ($\beta=.44, p<.0001$).
 As expected, the more a movie was seen, the better the videos were memorized.
-Figure \ref{fig:logistic_regression} (c) shows that this continues to be true even with more than 9 viewings.
-(However, the number of observations was very low (12) for videos which belongs to movies with 10 or more views.)
+Figure \ref{fig:logistic_regression} (b) shows that this continues to be true even with more than 9 viewings (however, the number of observations -- 12 -- was very low for videos which belong to movies with 10 or more views).
 One should note that the repetition of a viewing could not be the (only) factor involved in the above phenomenon; in particular, viewing again a movie may be the sign of a special interest which would explained a better memorization (e.g., via a greater attentional and emotional investment).
@@
 %Ajouter Age et expliquer que les movies sont plutôt jeune génération -- donner corrélation Age*Memorability => pas possible
-According to Figure \ref{fig:logistic_regression} (a), we observed no significant effects of the demographic factors (nationality, age and gender).
-This suggests that the videos were about equally susceptible to be recognized by the different participants (or, that the relations between these factors and the observations are too complex to have been captured by the model).
+We observed no significant effects of the demographic factors (nationality, age and gender).
+This suggests that the videos were equally susceptible to be recognized by the different participants (or, that the relations between these factors and the observations are too complex to have been captured by the model).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Until now, we have presented the video collection and the annotation protocol. 
+Until now, we have presented the video collection and the annotation protocol together with some insights on human VM. 
 %We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
@@
 We chose to split our dataset at the level of movies, instead of the videos, in order to avoid videos from the same movie being present in the training as well as the evaluation (validation+test) set.
+To ensure that such random split did not lead to any mismatch, we computed the average number of annotations per video in each of them.
+We observe that the numbers of annotations are balanced: each video in the test set has around 10 annotations while there are 9 annotations for each one in the train set on an average.
 
@@
 We run the SentiBank concept detector on each of these 10 frames and rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
-We extract a 300-dimensional word2vec \cite{word2vec} embeddings, for each of the 50 concepts and take an average to obtain a single vector per frame.
+We extract 300-dimensional word embeddings (Word2Vec \cite{word2vec}), for each of the 50 concepts and take an average to obtain a single vector per frame.
 We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
-Sentibank detectors also provide a 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
+Sentibank detectors also provide a 4096-dimensional feature for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
 In the end, we use a 300-dimensional concept vector and a 4096-dimensional feature vector.
@@
 For each of these 10 frames, we run the caption detector (code provided by the authors) and obtain a caption for the frame.
-For each non-functional word in the caption, we extract a 300-dimensional word2vec \cite{word2vec} embeddings and take an average of all the words to obtain a single vector per frame.
+For each non-functional word in the caption, we extract 300-dimensional word embeddings (Word2Vec \cite{word2vec}) and take an average across all the words to obtain a single vector per frame.
 We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single 300-dimensional feature vector for each video.
 
-\subsection{Modelling and evaluation}
+\subsection{Modelling and evaluation metric}
 \label{model-eval}
-We use the features discussed in Section \ref{feat-extr} to train a Support Vector Regression (SVR) model with two different kernels (linear and RBF) for the VM score prediction. 
-We use the grid search strategy to obtain the best hyper-parameters for SVR: $C$ = $\{0.1, 1, 10, 100, 1000\}$ and $\gamma$ = $\{0.01, 0.1, 1, 10, 100\}$.
-We use the standard regression metric: Mean Squared Error ($MSE$) for the optimization process.
-The choice of SVR is guided by the small size of the dataset. 
-We have chosen to go with the same regressor for all the features because our focus was mainly on identifying which features are more important for VM prediction. 
+We use the features discussed in Section \ref{feat-extr} to train a Support Vector Regression (SVR) model for the VM score prediction.
+The choice of SVR is guided by the nature of the problem as well as by the small size of the dataset.
+We have chosen to go with the same regressor for all the features because our focus is mainly on identifying which features are more important for VM prediction.
 This way we ensure that the difference in performances is because of the features themselves.
+Note that, in addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating them into a single feature vector.
+While performing such a concatenation, we use the low-dimension version of the features for C3D and SentiBank features, obtained after applying a dimension reduction method (PCA) to the original set.
+In the experiments where we use PCA for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
+
+We use the grid search strategy to obtain the best hyper-parameters for SVR in term of the Mean Squared Error ($MSE$) between the predicted scores and the ground-truth on the validation set.
+The choice of hyper-parameters in the grid search are: kernel = $\{$linear, RBF$\}$, $C$ = $\{0.1, 1, 10, 100, 1000\}$ and $\gamma$ = $\{0.01, 0.1, 1, 10, 100\}$.
 
 %Since we are posing the memorability prediction problem as a regression problem, we use the standard regression metric: Mean squared error ($MSE$) for evaluation.
-We use the spearmann correlation ($SpCorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
-We chose this particular metric as it gives us an indication of how close the predicted memorability scores are to human labelled ground-truth memorability scores.
+The prediction performance is evaluated by the Spearman correlation ($SpCorr$), which measures the rank correlation between the predicted memorability scores and the ground-truth scores.
+This metric is chosen as (1) we focus more on the relative memorability between videos rather than on  their absolute memorability scores, and (2) it gives an indication of how close the predicted memorability scores are to the human consistency when annotating memorability scores.
 
-In addition to the variety of features explained in Section \ref{feat-extr}, we also explore a combination of all the features by concatenating them into a single feature vector.
-While performing such a concatenation, we use the lower dimensional features for C3D and SentiBank features, obtained after applying PCA to the original set.
 %Each of the videos in our dataset has multiple annotations and we pick sequences with at least 4 annotations for training the models.
@@
 %Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies), before reporting the final scores on the test set (15 movies).
-We pick the SVR model parameters: kernel (linear or RBF), $C$ and $\gamma$ that give the best performance on the validation set.
-We retrain the model with these parameters using the training set and evaluate on the final test set.
-In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
-
-\subsection{Memorability prediction results}
-\label{pred-res}
-In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new videos.
-We report the results of the prediction on both the validation and test sets.
-We also report only correlation scores: $SpCorr$ and not $MSE$ keeping in mind the space constraints.
-Additionally, $SpCorr$, which provides a rank correlation between the predicted and the ground-truth memorability scores, is more informative.
-We compute the average number of annotations per video in the train and test set for both the cases ($geq$ 4 and $geq$ 8 annotations).
-We observe that the number of annotations in the train and test sets are balanced and there is no mismatch.
-For example, each sequence in the test set has around 10 annotations while there are 9 annotations for each sequence in the training on an average.
-
-Table \ref{res-4-10-ann} reports the the results of prediction capability of the model trained on different features.
-There are two sets of results reported in the table: $SpCorr$ ($\geq$ 4 annotations) and $SpCorr$ ($\geq$ 8 annotations).
-The first set of results correspond to the prediction capability of the model when trained on videos with at least 4 annotations and the second set corresponds to the results when trained on videos with at least 8 annotations.
-We will further discuss why we report the results with the model trained on videos with at least 8 annotations.
-
-Observing the first set of results in Table \ref{res-4-10-ann}, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set. 
-Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
-These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
-Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
-
-Other features like AudioSet come close to the performance of the above features (third best feature), but only audio signal does not seem to be enough for predicting video memorability.
-Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features.
-One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector.
-In future, we could look at selectively combining the features to investigate if that improves the performance.
-
-One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
-Please recall that we used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
-Observing the scores in Table \ref{res-4-10-ann}, we can say that the SentiBank features (fourth row from bottom) perform slightly better than the affect features (sixth row from bottom).
-But the models trained on image captioning features clearly out-perform those trained on emotion features.
-This could be because of the following reasons: our choice of features to capture emotion related information is not suitable for our task, or the performance of the emotion prediction models is not good enough for memorability prediction, or we are not able able to establish the correlation between emotion and memorability because of the limited size of our dataset.
-%From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
-%In the future, we could investigate more robust emotion/affect prediction models to confirm that both emotion and memorability are correlated as suggested in the literature \cite{emo-mem}.
-
-We further investigate the effect of the number of annotations, for videos in the training set, on the performance of the model.
-%Based on our experimental results, we observe that the image captioning feature performs the best while using training videos that have at least 4 annotations.
-%For this purpose, we train different SVR models with varying number of annotations on the training set using image captioning features only.
-We train a model, using image captioning features, on training sequences with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
-We repeat this process for different number of annotations per video in the training set: from 4 to 15 annotations.
-Please note that the validation set in each of the repetitions is fixed and only the training set changes.
-We provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set in Figure \ref{num-ann}.
-We observe that $SpCorr$ first increases up to 5 annotations and then remains constant before decreasing (beyond 8 annotations).
-In the wake of this observation, we investigate the performance of all the features when we train the regression model with videos that have at least 8 annotations.
+%We pick the SVR model parameters: kernel (linear or RBF), $C$ and $\gamma$ that give the best performance on the validation set.
+%We retrain the model with these parameters using the training set and evaluate on the final test set.
+%In the experiments where we use a dimension reduction method (PCA) for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
 
@@
 	\centering
-    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
-		\caption{Spearmann correlation on the validation set with models trained using image captioning features for varying number of annotations.}
+%	\includegraphics[width=0.7\columnwidth]{figures/annotations_modelling.png}
+	\includegraphics[width=0.7\columnwidth]{figures/annotations_modelling.png}
+	\caption{Spearman correlation on the validation set with models using image captioning features trained on videos with varying number of annotations.}
 	\label{num-ann}
@@
 		\hline
-    \textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$($geq$ 4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ ($geq$ 8 annotations)}}\\
+		\textbf{Feature} & \textbf{Feature type}& \textbf{Dimension}& \multicolumn{2}{c|}{\textbf{$SpCorr$($\geq$ 4 annotations)}} & \multicolumn{2}{c|}{\textbf{$SpCorr$ ($\geq$ 8 annotations)}}\\
 		% \hline
@@
 		C3D& visual spatio-temporal& 4096& 0.20& 0.26& 0.31& 0.34\\ \hline
-		C3D (PCA)& visual spatio-temporal (PCA)& 225& 0.24& 0.21& 0.18& 0.17\\ \hline
+		C3D (PCA)& visual spatio-temporal & 225& 0.24& 0.21& 0.18& 0.17\\ \hline
 		AudioSet& audio related& 128& 0.23& 0.22& 0.21& 0.24\\ \hline
@@
 		SentiBank features& emotion related& 4096& 0.25& 0.21& 0.27& 0.26\\ \hline
-		SentiBank features (PCA)& emotion related (PCA)& 225& 0.22& 0.21& 0.22& 0.23\\ \hline
-		Captions& visual-semantics & 300& 0.29& 0.31& 0.39& 0.38 \\ \hline
-		Combination& combine-all-features (PCA)& 1578& 0.24& 0.23& 0.29& 0.27\\ \hline		
-    \hline
+		SentiBank features (PCA)& emotion related & 225& 0.22& 0.21& 0.22& 0.23\\ \hline
+		Captions& visual semantics & 300& \textbf{0.29}& \textbf{0.31}& \textbf{0.39}& \textbf{0.38} \\ \hline
+		Combination (PCA)& combine all features & 1578& 0.24& 0.23& 0.29& 0.27\\ \hline		
 	\end{tabular}
-  \caption{Prediction results (Spearmann correlation) on validation and test data for different features with models trained on videos that have at least 4 (columns 4-5) or 8 (columns 6-7) annotations.}
+	\caption{Prediction results in terms of Spearman correlation scores on validation and test data for different features with models trained on videos that have at least 4 (columns 4-5) or 8 (columns 6-7) annotations.}
 	\label{res-4-10-ann}	
@@
 
-These results are reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ ( $\geq$ 8 annotations).
-Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that the models trained on sequences with at least 8 annotations perform better than the models trained on sequences with at least 4 annotations.
-The only exception being the C3D (PCA) feature, where the model trained on sequences with at least 4 annotations performs marginally better than the model trained on sequences with at least 8 annotations.
-Additionally, we also observe that the scores for validation set and test set are very similar across the two sets of results.
+\subsection{Memorability prediction results}
+\label{pred-res}
+In this section, we will discuss how the models trained on different features perform while predicting memorability scores of new videos.
+%We report the results of the prediction on both the validation and test sets.
+%We also report only correlation scores: $SpCorr$ and not $MSE$ keeping in mind the space constraints.
+%Additionally, $SpCorr$ is more informative because it provides a rank correlation between the predicted memorability scores and the ground truth.
+%We compute the average number of annotations per video in the train and test set for both the cases ($\geq$ 4 and $\geq$ 8 annotations).
+%We observe that the number of annotations in the train and test sets are balanced and there is no mismatch.
+%For example, each sequence in the test set has around 10 annotations while there are 9 annotations for each sequence in the training on an average.
+Table \ref{res-4-10-ann} reports the prediction results obtained by the SVR model when trained on different features, for the validation and test sets.
+Two sets of results are reported: $SpCorr$ ($\geq$ 4 annotations) and $SpCorr$ ($\geq$ 8 annotations). The former 
+corresponds to the prediction capability of the model when trained on videos with at least 4 annotations, and the latter corresponds to the results when trained on videos with at least 8 annotations where the ground-truth is better annotated.
+%We will further discuss why we report the results with the model trained on videos with at least 8 annotations.
+
+As it can be seen, visual semantic features derived from image captioning clearly offer better prediction results compared to all other considered features.
+This is not surprising as they capture visual attributes along the scene, which are known to play an important role in human memory.
+This is also inline with a previous study in IM \cite{squalli_2017_deep} where image captioning-based features helped better predict IM than the conventional CNN features.
+%These best prediction results (i.e., 0.31 and 0.38 for the test sets) are quite close to the human consistency shown in Figure \ref{fig:human_consistency}b (i.e.
+about 0.36 when having 4 annotations per video and about 0.47 when having 8 annotations per video).
+When predicting VM on the test set, C3D features showed to be quite effective to the task as they encode the visual spatio-temporal information of the video.
+On the contrary, audio information captured by AudioSet features does not seem to be enough for VM prediction.
+One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
+However, when observing the results in Table \ref{res-4-10-ann}, both Affect \cite{affect} and SentiBank \cite{sb-feat} performed quite poorly compared to the image captioning features.
+%Observing the first set of results in Table \ref{res-4-10-ann}, where we have at least 4 annotations per video, we can clearly see that the image captions and C3D occupy the top-2 places in terms of performance on the test set.
+%Image captions capture the visual semantics in the video, while C3D features encode the visual spatio-temporal information.
+%These results are consistent with previous results on predicting image memorability \cite{squalli_2017_deep}.
+%Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
+%Other features like AudioSet come close to the performance of the C3D, but only audio signal does not seem to be enough for predicting video memorability.
+Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features.
+One of the reasons for this could be that there is a lot of redundance when combining all the features into a single feature vector, or it could be that the size of the combined feature vector is too big when compared to the size of the dataset.
+Thus, in future work we could look at selectively combining the features to investigate if that improves the performance.
+
+%One of our initial hypotheses was that emotion would play an important role in VM, supported by literature from psychology \cite{emo-mem}.
+%Please recall that we used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
+%However, when observing the results in Table \ref{res-4-10-ann}, both Affect \cite{affect} and SentiBank \cite{sb-feat} performed quite poorly compared to the image captioning.
+%we can say that the SentiBank features (fourth row from bottom) perform slightly better than the affect features (sixth row from bottom).
+%But the models trained on image captioning features clearly out-perform those trained on emotion features.
+%This could be because of the following reasons: our choice of features to capture emotion related information is not suitable for the task, or the performance of the emotion prediction models is not good enough for memorability prediction, or we are unable to establish the correlation between emotion and memorability because of the limited size of our dataset.
+%From the analysis on this dataset and the features we used for capturing emotion related information, we \emph{cannot} conclude that emotion plays an important role in memorability.
+%In the future, we could investigate more robust emotion/affect prediction models to confirm that both emotion and memorability are correlated as suggested in the literature \cite{emo-mem}.
+
+As our dataset contains videos with different numbers of annotations, we further investigate the effect of the number of annotations on the prediction performance.
+%Based on our experimental results, we observe that the image captioning feature performs the best while using training videos that have at least 4 annotations.
+%For this purpose, we train different SVR models with varying number of annotations on the training set using image captioning features only.
+For this purpose, we train a first SVR model, using image captioning features, on videos with at least 4 annotations and use this model to predict the memorability score for videos in the validation set.
+We repeat this process for different numbers of annotations per video (from 4 to 15) in the training set.
+Please note that the validation set in each of the repetitions is fixed and only the training set changes.
+We provide a demonstration of how $SpCorr$ varies with an increasing number of annotations in the training set in Figure \ref{num-ann}.
+As can be seen, $SpCorr$ first increases up to 5 annotations and then remains more or less constant before decreasing (beyond 10 annotations).
+In the wake of this observation, we also investigated the performance of all the features when we train the regression model with videos that have at least 8 annotations as reported in the second part of Table \ref{res-4-10-ann}: $SpCorr$ ( $\geq$ 8 annotations).
+
+Comparing the two sets of results in Table \ref{res-4-10-ann}, we observe that globally the models trained on videos with at least 8 annotations perform better than the models trained on videos with at least 4 annotations.
+These results are comparable to the human consistency analysis shown in Section \ref{ssec:human}.
+%We can say that with an increase in the number of annotations per video, we can expect an increase in the performance until a certain number of annotations (8 annotations in this case).
+%From the table, we also observe that the scores for validation set and test set are very similar across the two sets of results.
+We finally performed an additional 10-fold cross-validation on videos with at least 8 annotations (train+validation set) using the image captioning features.
+From this study, we re-confirmed that there is no overfitting issue in our model and we observed the average value of $SpCorr$ across the 10 folds to be 0.33, which is close to the performance on the test set.
+%Looking at the performance in the cross-validation experiment, we can say that there is no over-fitting because of the small dataset.
 
@@
 \section{Conclusions}
+\label{conclusions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %protocol
-We proposed a new protocol to collect long-term memorability annotations for videos.
-It enabled us to measure memory performance after weeks to years.
-It appears from the analysis of the data that memory of videos continue to decrease for years, which justify a measurement of memory performance after a significant retention duration, longer than proposed in previous work.
-The principal weakness of our protocol is the part of subjectivity susceptible to enter in our measure of memorability, that one should counteract by appropriate controls.
-Our current work focus on collecting VM annotations at a large scale using crowdsourcing, measuring VM in a more objective manner.
-The implemented protocol measure memory performance after two different retention durations, which will enable us to understand what makes a video lastingly memorable.
+In this paper, we have presented a novel protocol to collect long-term memorability annotations for videos, which enabled us to build an important dataset to support research in this subject.
+We then performed a range of statistical studies on this dataset to understand important factors in the annotation process as well as how they can affect the video memorability.
+One of our key observations is that high human consistency in video memorability can be obtained by only about 18 annotations, which is significantly lower than for images according to the previous studies.
+We finally proposed computational models for video memorability prediction where we investigated the use of various audio-visual features for the task.
+For this, we observed that the visual semantic features offer the best prediction result, which re-confirms the correlation between the visual attributes and memorability.
+Our current work focuses on building a large scale video memorability dataset using crowdsourcing annotations.
+
+%and built a dataset publicly available.
+%We measured memory performance after weeks to years.
+%It appears from the analysis that memory of videos continues to decrease for years, which justifies a measurement of memory performance after a significant retention duration, longer than what is proposed in previous work.
+%The principal weakness of our protocol is the part of subjectivity susceptible to enter in our measure of memorability, that one should counteract by appropriate controls.
+%Our current work focuses on collecting VM annotations at a large scale using crowdsourcing, measuring VM in a more objective manner.
 
 %modelling Karthik
-After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground-truth memorability scores.
-It might be very interesting to investigate if these findings hold for generic video types other than movies.
-Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
-We also found that emotion features are not that well correlated with ground-truth memorability scores, as indicated in psychology literature.
-An immediate direction for future research would be to explore other approaches to encode emotion in videos and investigate whether we can improve the prediction performance using these approaches.
-We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
+%After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground truth memorability scores.
+%It might be very interesting to investigate if these findings hold for generic video types other than movies.
+%Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
+%We also found that emotion features are not that well correlated with ground truth memorability scores, as indicated in psychology literature.
 
+%An immediate direction for future research would be to explore other approaches to encode emotion in videos and investigate whether we can improve the prediction performance using these approaches.
+%We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+%\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                   % on the last page of the document manually. It shortens
@@
                                   % sure that you do not shorten the textheight too much.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \bibliography{references} 
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-04-27 ARTICLE.tex
Final version of accepted paper after reviewing

@@
 \usepackage{subfig}
-\usepackage{balance}
 \usepackage{caption}
@@
 \usepackage{amsmath}
+%\usepackage{times}
+\usepackage{hyperref}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{10.475/123_4}
-
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% CONFERENCE
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Conference
-\acmConference[ICMR 2018]{ACM Yokohama conference}{June 2018}{Yokohama, Japan} 
+\copyrightyear{2018} 
 \acmYear{2018} 
-%\copyrightyear{2018}
-%\acmArticle{4}%??
-%\acmPrice{15.00}%??
+\setcopyright{acmcopyright}
+\acmConference[ICMR '18]{2018 International Conference on Multimedia Retrieval}{June 11--14, 2018}{Yokohama, Japan}
+\acmBooktitle{ICMR '18: 2018 International Conference on Multimedia Retrieval, June 11--14, 2018, Yokohama, Japan}
+\acmPrice{15.00}
+\acmDOI{10.1145/3206025.3206056}
+\acmISBN{978-1-4503-5046-4/18/06}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Annotating, understanding, and predicting long-term video memorability}
+\title{Annotating, Understanding, and Predicting Long-term Video Memorability}
 %\titlenote{}
@@
 %author 1
-%\author{Romain Cohendet}
+\author{Romain Cohendet$^{\ast}$}{\thanks{$^{\ast}$Both authors contributed equally to this work.}}
 %\authornote{}
 %\orcid{1234-5678-9012}
-%\affiliation{%
-%	\institution{Technicolor}
-%	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
-%	\city{Rennes} 
-%	\state{France} 
-%	\postcode{}
-%}
-%\email{romain.cohendet@technicolor.com}
+\affiliation{%
+	\institution{Technicolor}
+	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
+	\city{Rennes}
+	\state{France}
+	\postcode{}
+}
+\email{Romain.Cohendet@technicolor.com}
 
 %author 2
-%\author{Karthik Yadati}
-%\affiliation{%
-%	\institution{University of Delft}
-%	\streetaddress{}
-%	\city{Delft} 
-%	\state{} 
-%	\postcode{}
-%}
-%\email{N.K.Yadati@tudelft.nl}
+\author{Karthik Yadati$^{\ast}$}
+	\affiliation{%
+	\institution{Delft University of Technology }
+	\streetaddress{}
+	\city{Delft} 
+	\state{} 
+	\postcode{}
+}
+\email{N.K.Yadati@tudelft.nl}
 
 %author 3
-%\author{Ngoc Khanh Duong}
-%\affiliation{%
-%	\institution{Technicolor}
-%	\city{Rennes} 
-%	\country{France}}
-%\email{Quang-Khanh-Ngoc.Duong@technicolor.com}
+\author{Ngoc Q. K. Duong}
+\affiliation{%
+	\institution{Technicolor}
+	\city{Rennes} 
+	\country{France}}
+\email{Quang-Khanh-Ngoc.Duong@technicolor.com}
 
 %author 4
-%\author{Claire-Helene Demarty}
-%\affiliation{%
-%	\institution{Technicolor}
-%	\city{Rennes}
-%	\country{France}
-%}
-%\email{Claire-Helene.Demarty@technicolor.com}
-
-%%%%%%%%%%%%
-%%% for blind review %%%
-%author 1
-%\author{Removed for review}
-%author 2
-%\author{Removed for review}
+\author{Claire-H\'el\`ene Demarty}
+\affiliation{%
+	\institution{Technicolor}
+	\city{Rennes}
+	\country{France}
+}
+\email{Claire-Helene.Demarty@technicolor.com}
 
 % The default list of authors is too long for headers.
-\renewcommand{\shortauthors}{Removed for blind review}%{R. Cohendet et al.}%Removed for blind review
+\renewcommand{\shortauthors}{R. Cohendet et al.}%Removed for blind review
 
@@
 
-%\ccsdesc[500]{Computer systems organization~Embedded systems}
-%\ccsdesc[300]{Computer systems organization~Redundancy}
-%\ccsdesc{Computer systems organization~Robotics}
-%\ccsdesc[100]{Networks~Network reliability}
+%\ccsdesc[500]{Computing methodologies~Computer vision}
+%\ccsdesc[500]{Computing methodologies~Scene understanding}
+%\ccsdesc[300]{Computing methodologies~Visual content-based indexing and retrieval}
+%\ccsdesc[300]{Computing methodologies~Machine learning}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\keywords{Video memorability, Long-term memory, Measurement protocol, Deep learning, Multimedia information retrieval}
+\keywords{Video memorability, Long-term memory, Scene understanding, Measurement protocol, Global video features, Attributes}
 
@@
  %\input{samplebody-conf}
-
-%% About the ICMR article on Media retrieval with VM
-	% focus on "Multimedia content-based retrieval"
-	% Length: from 6 to 8 pages + 1 additional page for references (no longer distinction between long and short papers)
-	% Submission deadline: 2018/01/20
-	% Contributions addressing the challenges of large-scale search and user behavior analysis are especially welcome
 
@@
 	\item We assess and analyze the collected data and provide useful insights on the understanding of VM (section \ref{study_memorability}).
-	\item We build a computational model based on machine learning techniques which allows to predict VM score of a given video.
+	\item We build a computational model based on machine learning techniques which allows to predict a VM score for a given video.
 For this purpose, we investigate the use of various types of audio and visual features, ranging from low-level characteristics to emotional and (highly) semantic features (section \ref{mem-pred}).
@@
 \section{Previous work}
-In this section, we review previous work on image and video memorability prediction, focusing on annotation protocols and modelling aspects.
+In this section, we review previous work on IM and VM prediction, focusing on annotation protocols and modelling aspects.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Several factors have also been highlighted for their critical influence on long-term memory, such as emotion \cite{kensinger_2008_memory}, semantics \cite{quillan_1966_semantic}, several demographic factors \cite{cohendet_2016_using}, memory re-evocation \cite{nadel_1997_memory}, or passage of time \cite{mcgaugh_2000_memory}.
-These different factors are all important to better understand memorability, and are found valuable computational features for image and video memorability prediction.
+These different factors are all important to better understand memorability, and are found valuable computational features for IM and VM prediction.
 
@@
 However, this measure might not be sufficient to represent the questions in all their complexities, such as the ease for a person to draw images from the words, to connect the words to the associated concepts in memory, to associate the words to the corresponding scene, etc.
-Moreover, there is an unequal distribution of the reading comprehension of English among Amazon Mechanical Turk workers, who are not necessary anglophone people.
+Moreover, there is an unequal distribution of the reading comprehension of English among Amazon Mechanical Turk workers, who are not necessary native English speakers.
 This problem of a potential unequal complexity of the question becomes even more important in view of the use by authors of the response time to calculate memorability scores, which might also critically depend on the complexity of the questions.
@@
 After collecting the data, we kept only the 660 videos that had been seen at least 4 times as targets (from the initial set of 700 videos).
-On average, each video of our dataset has been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enters in the calculation of a memorability score.
+On average, each video of our dataset had been viewed as a target by $10.7$ participants; which corresponds to the mean number of observations that enter into the calculation of a memorability score.
 We then assigned a memorability score to each video, defined as the correct recognition rate of the video when viewed as target.
@@
 Figure \ref{fig:response_time_vs_memorability}(a) provides a distribution of the videos according to their degree of memorability.
-The dataset is publicly released here\footnote{\textit{www.remove-for-blind-review}.
-The original videos are not released because of copyright issues, but instead we provide the process we used to extract them from the movies and the features used in section \ref{mem-pred}.}.
+The dataset is publicly released here\footnote{\url{https://www.technicolor.com/dream/research-innovation/movie-memorability-dataset}}.
+In addition to the videos and corresponding ground truth, we also provide the features used in section \ref{mem-pred}.
 
@@
 We follow the method proposed in \cite{isola_2014_makes} to measure human consistency when assessing memorability of videos.
-We randomly split our 104 participants into two independent groups of equal size, and calculate how well video memorability scores from the first group of participants match with video memorability scores from the second group.
+We randomly split our 104 participants into two independent groups of equal size, and calculate how well VM scores from the first group of participants match with VM scores from the second group.
 Averaging over 25 random half-split trials, an average Spearman's rank correlation (i.e., a global human consistency) of $0.57$ is observed between these pairs of scores.
@@
 In our experiment, participants were given clear instructions that they had to really recognize any video they were presented as already seen, and not only guess that a video was coming from a movie whose title was proposed in the questionnaire.
-In this section, we perform an analysis to compare neutral videos, which contain no element that would enable participants to guess that a video belongs to a particular movie, and typical videos.
-Indeed, if neutral videos received objective answers from participants, it might be more subjective for typical videos, that could be more or less easily related to the movie they belong to.
+However, a number of participants reported difficulty in making the distinction between guessing and recognizing: sometimes, because they knew that a video belonged to a movie they had seen, they had the false feeling of recognizing it.
+In this section, we perform an analysis to compare neutral videos, which contain no element that would enable participants to easily guess that a video belongs to a particular movie, and typical videos, that could be more or less easily related to the movie they belong to.
 
-A Wilcoxon rank-sum test indicated that the memorability was greater for neutral ($Mdn=.20, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, with $Z=10.22, p<.00001$.
-Apart from the subjectivity aspect, we expected such a result because neutral videos contain less contextual elements, useful for recognition.
-Thus, this result does not necessarily mean that participants tended to guess -- rather than simply recognize -- videos selected from movies they have seen.
-
-As for the human consistency on memorability, a Wilcoxon rank-sum test indicated that it was slightly greater for neutral ($Mdn=.44, \mu=.45$) than for typical ($Mdn=.40, \mu=.41$) videos, with $Z=2.75, p<.01$.
-Along with the comments collected from the participants, who have as a majority reported difficulty to know if they were guessing or really recognizing the videos, this result suggests that human congruency is higher for more 'objectively' recognized segments than for ones with subjectivity as part of the equation.
-One could note there is probably not just pure subjectivity here.
-Context might have biased the results as it might have helped some participants in their recognition task.
-If confirmed, this would constitute a weakness of our protocol to collect extensive data, that one should in that case counteract by adapted measures of control.
+A Wilcoxon rank-sum test indicated that the memorability was lower for neutral ($Mdn=.20, \mu=.24$) than for typical ($Mdn=.53, \mu=.53$) videos, with $Z=10.22, p<.00001$.
+This result does not necessarily mean that participants tended to guess when it was possible to do so -- rather than simply recognize -- videos selected from movies they have seen.
+Indeed, neutral videos contain far less contextual elements useful for recognition than typical ones, making the latter -- if more easily "guessable" -- at least more easily recognizable.
+However, it should draw our attention to the possibility that our measure is not purely objective, but -- to some extent -- context dependent.
+This would constitute the main weakness of our protocol, that one should counteract by adapted measures of control.
 
@@
 This also suggests either that the most memorable videos are also the most accessible in memory, and/or that the most memorable videos contain more early recognizable elements than the less memorable ones.
-In \cite{shekhar_2017_show}, the response time of the participants was taken to be the measure of video memorability.
+In \cite{shekhar_2017_show}, the response time of the participants was taken to be the measure of VM.
 The authors chose this measure to avoid a long gap between viewing and recall stage.
@@
 %Ce qui à la fin de cette section vient avant permet de conclure cette partie en disant : la meilleure façon de calculer les scores de mémorabilité, c'est comme ça.
-% The question here is: "Must we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the movie --> No' --> So the point here is just to provide an overview.
+% The question here is: "Should we correct our memorabiliy scores by taking into account the mean memorability of the participants --> No, of the movie --> No' --> So the point here is just to provide an overview.
 
@@
 In the experiments where we use PCA for C3D and SentiBank features, we retain 95\% of variance in the data while reducing the feature dimensions.
+Note that we investigated other models like LSTM, to model the sequential nature of videos.
+The results were not very encouraging and this could be because of the relatively smaller size of our dataset.
 
@@
 %Our dataset consists of videos taken from movies that have a specific story-line; semantics and spatio-temporal information seem to be playing an important role in people remembering specific scenes from movies.
-%Other features like AudioSet come close to the performance of the C3D, but only audio signal does not seem to be enough for predicting video memorability.
+%Other features like AudioSet come close to the performance of the C3D, but only audio signal does not seem to be enough for predicting VM.
 Another observation from Table \ref{res-4-10-ann} is that the combination of all the features (last row in the table) does not appear in the top-3 best performing features.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%protocol
-In this paper, we have presented a novel protocol to collect long-term memorability annotations for videos, which enabled us to build an important dataset to support research in this subject.
-We then performed a range of statistical studies on this dataset to understand important factors in the annotation process as well as how they can affect the video memorability.
-One of our key observations is that high human consistency in video memorability can be obtained by only about 18 annotations, which is significantly lower than for images according to the previous studies.
-We finally proposed computational models for video memorability prediction where we investigated the use of various audio-visual features for the task.
-For this, we observed that the visual semantic features offer the best prediction result, which re-confirms the correlation between the visual attributes and memorability.
-Our current work focuses on building a large scale video memorability dataset using crowdsourcing annotations.
-
-%and built a dataset publicly available.
-%We measured memory performance after weeks to years.
-%It appears from the analysis that memory of videos continues to decrease for years, which justifies a measurement of memory performance after a significant retention duration, longer than what is proposed in previous work.
-%The principal weakness of our protocol is the part of subjectivity susceptible to enter in our measure of memorability, that one should counteract by appropriate controls.
-%Our current work focuses on collecting VM annotations at a large scale using crowdsourcing, measuring VM in a more objective manner.
+%Protocol and M understanding
+In this paper, we have presented a novel protocol to collect long-term memorability annotations for videos, which enabled us to build a dataset to support research in this subject.
+We then performed a range of statistical studies on this dataset to understand important factors in the annotation process as well as how they can affect video memorability.
+The main strength of our protocol is that it enables to measure memory performance weeks to years after memorization without requiring a longitudinal study.
+It appears from the analysis that memory of videos continues to decrease for years, which justifies such a distant measurement, in contrast to previous work.
+The counterpart is that we cannot control the memorizing process.
+In particular, the correctness of the answers to the questionnaire designed to collect information about the participants' prior memory is not absolute.
+Even with additional information about the movies (e.g. showing an excerpt/trailer), the possibility would remain that a participant falsely answers he/she has already seen/never seen a movie.
+%We hope, however, this bias relatively limited.
+Our current work focuses on collecting VM annotations at a large scale using crowdsourcing, measuring VM in a more objective manner, by controlling the memorizing step.
+We proceed of the measurement after different periods of retention (from a few minutes to a few days), which enables us also to study long-term memory, although not as long as in the present experiment.
+Another limitation of the presented protocol is the limited choice of content.
+To design our experiment, we needed contents broadly disseminated among the population surveyed.
+In our current work, we are not limited anymore by this problem, and decided to work on neutral videos.
 
 %modelling Karthik
+We finally proposed computational models for VM prediction where we investigated the use of various audio and visual features for the task.
+For this, we observed that the visual semantic features offer the best prediction result, which re-confirms the correlation between the visual attributes and memorability.
+Although very promising, these results seem to be limited by the relative small size of the dataset.
+Again, our current work might enable us to reconfirm these results on a larger dataset and for more general videos.
+
+%Moreover, our measure captures not only pure memory performance but also a part of subjectivity.
+%Why not say also that we might try to check the response to the questionnaire by checking as proposed by one of the reviewers the answers against Netflix history (of course this can be done only for movies that are on Netflix and that were viewed since Netflix start but this is a start for cross checking the questionnaire.
+// %One improvement of our protocol could be to replace our questionnaire to by a check of a -- e.g.
+Netflix -- history, to be sure if the participants see a movie -- but problem os, if he not seen?
+
 %After exploring a variety of generic (C3D, AudioSet), perceptual (SentiBank, Affect), and semantic (Image captions) features, we can say that a model trained with  semantic features provides predictions that are most correlated with ground truth memorability scores.
@@
 %Future research could involve investigating the memorability prediction performance on other generic videos using the existing set of features as well as other important features.
-%We also found that emotion features are not that well correlated with ground truth memorability scores, as indicated in psychology literature.
+%We also found that emotion features are not that well correlated with ground truth memorability scores, as indicated in psychology literature.%no result is no effect att!!
 
@@
 %We also investigated the number of annotations required to train a model and found a sweet spot where the performance is optimal.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
                                   % sure that you do not shorten the textheight too much.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
